### 2016-05-23 ms.tex
tex added

\documentclass[aps,pre,twocolumn,letterpaper,floatfix,showpacs]{revtex4}
\usepackage{graphicx} 
\usepackage{amsmath,amssymb,amsfonts} 
\usepackage{mathtools}
\usepackage{pdfpages}
\usepackage{afterpage}
\usepackage[hidelinks]{hyperref} 
\usepackage{epstopdf}
\begin{document}
\title{Generating realistic nanopores using procedural methods}
%\author{A. Hafreager, N. Groeneboom, A. Malthe-S\oe renssen$^1$}
\email{anders.hafreager@fys.uio.no}
\author{Anders Hafreager $^{1}$} 
\author{Nicolaas Groeneboom $^{1}$} 
\author{Anders Malthe-S\o renssen $^1$}
\affiliation{$^1$Department of Physics - University of Oslo\\Sem S{\ae}lands vei 24, NO-0316, Oslo, Norway }
\date{\today}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract} 
\end{abstract}

\maketitle

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
In recent years, computational ...

\section{Method}

\subsection{Procedural noise}
\label{sec:perlin}

Procedural algorithmic methods for creating realistic-looking
fractal-like patterns have been around for more than 30 years. The most famous of these
algorithms is Perlin noise, first developed by Ken Perlin in 1982 for
use in the movie ``Tron''.  Perlin noise is a type of pseudo-random
gradient noise that mimics a Gaussian field, but at low computational cost. 
An improved version of Perlin noise is called Simplex
noise, which has several advantages over standard Perlin noise in
terms of scalability and speed.

\subsubsection{Perlin noise - an overview}
An overview of the current state of procedural noise functions can be
found in \cite{lagae:2010}.  Perlin noise \citep{perlin:2002} 
is a band limited repeatable pseudo-random function
from $\mathbb R^n \to \mathbb R$ that is computationally fast and has
properties which make it suitable for creating patterns that mimic
nature. Perlin noise is an approximation to Gaussian filtered noise
implemented as a pseudo-random spline. The algorithm uses a predefined
grid of gradients pointing in a random direction, where for any point
within a grid, we interpolate the gradients from the four corners. The
algorithm is as follows:
\begin{itemize}
  \item[1.] For a point, obtain the four closest gradients in the grid.
   \item [2.] For each of the gradients, calculate the dot product
     between the gradient and a vector defined by the  the distance
     between the the point and the current grid corner. 
   \item [3.] Instead of linear interpolation $F(t) = t$, use a fade function to produce smoother interpolated values. Perlin noise typically uses the following fade function: $F(t) = t ^3 (t  (6 - 15) + 10)$.
    \item [4.] Obtain the final value by linearly interpolating between the x-axis and
      use this result to linearly interpolate the y-axis. 
\end{itemize} 
Simplex noise works in a similar manner, but requires fewer
computational steps, scales better in higher ($\ge 3$) dimensions and has fewer 
anisotropic properties.

\subsubsection{Properties of procedural noise}
Standard Perlin noise is known for having several anisotropic features
\citep{lagae:2010}. Perlin noise also gives rise to a ``textile''
pattern in the $x$ and $y$-directions, which is evident from the upper
left part of figure \ref{fig:noise}, depicting 2D Perlin noise in
a single octave. The upper-right part shows the same node for Simplex noise,
with less obvious patterns. However, when combining several octaves,
as depicted in the two lower parts, these anisotropic patterns tend to cancel
out. In any case, for observations of weak gravitational lensing, these anisotropies would occur on much smaller 
scales than the typical resolution. 
In addition, we have calculated the statistical distributions for Perlin and simplex noise
with $\sigma = 0.2$, and have plotted the results in figure \ref{fig:properties}.
Each procedural noise method has distributions that are close to normal, 
but with some intrinsically shifted sigma and mean.

\begin{figure}
\includegraphics[width=.5\textwidth]{noise.eps}
\caption{Perlin noise (left) vs simplex noise (right). Note the
  straight x-y pattern in the single upper Perlin mode and the less
  anisotropic simplex mode. Lower part: combining octaves tend
  to remove these patterns.}
\label{fig:noise}
\end{figure}

\begin{figure}
\includegraphics[width=.5\textwidth]{perlinstats.eps}
%\mbox{\epsfig{file=perlinstats.eps, width=\linewidth}}
\caption{Normalized distributions of Perlin noise and simplex noise compared with a
  Gaussian distribution. Note how both noise functions produce
  near-Gaussian, but slightly offset distributions.}
\label{fig:properties}
\end{figure}

\subsection{Multiple octaves: creating patterns}
\label{sec:octaves}
The different octaves from procedural noise functions can be combined
to produce a wide range of nature-like patterns. A generic
N-dimensional procedural function $\Phi(\vec x): \mathbb R^N \to
\mathbb R$ could be expressed as
\begin{equation}
\label{eq:procedural}
 \Phi(\vec x) = \Theta \Big(\sum_k \kappa (k) P(  f_s (\vec x,k) )) \Big),
 \end{equation}
where $f_s(\vec x,k)$ describes the frequency scale modifier (e.g. $f_s(\vec x,k) =
k\vec x$), $\kappa(k)$ is the octave amplitude (eg $\kappa(k) = 1/k$)
and $\Theta(x)$ an overall modifier (e.g. $\Theta(x) = x$ or
$\Theta(x) = \frac{1}{x}$). In a sense, $\kappa(k)$ is similar to Fourier coefficients. 
In its simplest case, summing directly
over the octaves with amplitude $1/k$ results in 
\begin{equation}
\label{eq:perlinlinear}
 \Phi(\vec x) = \sum_k \frac{1}{k} P( k\vec x), 
\end{equation}
and produces an 2D procedural noise pattern depicted in the lower
part of figure \ref{fig:noise}.

\ref{burganos1998gas}

The generic noise model presented in \ref{eq:procedural} can easily be extrapolated to 3 dimensions. In Lammps, we
simulate a uniform box of N $SiO_2$ particles without constraints, and proceed by cutting away particles above a given threshold $T$.
\begin{figure}
\includegraphics[width=.45\textwidth]{pores2.png}
%#\includegraphics[width=.5\textwidth]{pores2.eps}
%\includepdf{ pores1.pdf }
\caption{Model with best-fit parameters}
We construct a generic sum over procedural noise octaves with the expression: 
\label{fig:model12}
\end{figure}
\begin{equation}
	M_1(\vec x) = \sum_{\omega=0}^{\Omega} A \psi^{(\frac{1}{2})^{\omega +1}}   N\big((\vec x + \sigma \cdot \omega)\nu \cdot 2^\omega \big),
\label{eq:noisemodel1}
\end{equation}
where $\Omega$ is the total amount of octaves, $\psi$ the persistence (scale falloff), $N$ a generic $N$-dimensional noise function, $\sigma$ a predefined random vector that removes anisotropic effects and $\nu$ the frequency modifier. For each mode, the frequency is thus multiplied, while the overall amplitude is damped by a given persistence. In the end, we simulate a large set of bulk particles without any walls, and use the noise function to remove particles that are not in a procedural void.

\subsection{Simulations}
LAMMPS $SiO_2$ simulation heated -> expand box -> cool down
\begin{figure}
\includegraphics[width=.45\textwidth]{pores1.png}
%#\includegraphics[width=.5\textwidth]{pores2.eps}
%\includepdf{ pores1.pdf }
\caption{LAMMPS $SiO_2$ simulation.}
\label{fig:model12}
\end{figure}

\subsection{Implementation}
InfiniteDiffusion - nanopores. 
Octree

\subsection{Model Measure}
In order to compare our procedural noise model with actual simulations, we define various measures that quantify the internal structure of the 
\subsubsection{Distance-to-atom (DTA)}
\subsubsection{Fractal Dimension}
\subsubsection{Octree Counting Measure (OCM)}

\subsection{Diffusion Measure}

\begin{figure}
\includegraphics[width=.5\textwidth]{model1.eps}
\caption{blahblah}
\label{fig:model12}
\end{figure}

\section{Results}

\subsection{Model measure}

\subsection{Diffusion $R^2$}
\begin{figure}
\includegraphics[width=.5\textwidth]{compare.png}
\caption{Results. }
\label{fig:model12}
\end{figure}

\section{Conclusion}
It works! Hopefully.
\subsection{Future work}
Stress / shear test both models. Etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{bibliography}

%\begin{thebibliography}{9}

%\bibitem[Groeneboom2014]{groeneboom2014} Groeneboom, N.~E., \& Dahle, H.\ 2014, \apj, 783, 138

%\end{thebibliography}

\end{document}

### 2017-01-13 ms.tex
Finally working on the paper again..

@@
 simulate a uniform box of N $SiO_2$ particles without constraints, and proceed by cutting away particles above a given threshold $T$.
-\begin{figure}
-\includegraphics[width=.45\textwidth]{pores2.png}
+%\begin{figure}
+%\includegraphics[width=.45\textwidth]{pores2.png}
 %#\includegraphics[width=.5\textwidth]{pores2.eps}
 %\includepdf{ pores1.pdf }
-\caption{Model with best-fit parameters}
+%\caption{Model with best-fit parameters}
+%\label{fig:model12}
+%\end{figure}
+
+%\subsection{Simulations}
+%LAMMPS $SiO_2$ simulation heated -> expand box -> cool down
+%\begin{figure}
+%\includegraphics[width=.45\textwidth]{pores1.png}
+%%#\includegraphics[width=.5\textwidth]{pores2.eps}
+%%\includepdf{ pores1.pdf }
+%\caption{LAMMPS $SiO_2$ simulation.}
+%\label{fig:model12}
+%\end{figure}
+
+\section{Model}
 We construct a generic sum over procedural noise octaves with the expression: 
-\label{fig:model12}
-\end{figure}
 \begin{equation}
-	M_1(\vec x) = \sum_{\omega=0}^{\Omega} A \psi^{(\frac{1}{2})^{\omega +1}}   N\big((\vec x + \sigma \cdot \omega)\nu \cdot 2^\omega \big),
+  M_1(\vec x) = \sum_{\omega=0}^{\Omega} \psi^{(\frac{1}{2})^{\omega +1}}   N\big((\vec x + \sigma \cdot \omega)\nu \cdot 2^\omega \big),
 \label{eq:noisemodel1}
 \end{equation}
-where $\Omega$ is the total amount of octaves, $\psi$ the persistence (scale falloff), $N$ a generic $N$-dimensional noise function, $\sigma$ a predefined random vector that removes anisotropic effects and $\nu$ the frequency modifier.
+where $\Omega$ is the total amount of octaves, $\psi$ the persistence (scale falloff), $N$ a generic $N$-dimensional noise function, $\sigma$ a predefined random vector that removes anisotropic artifacts and $\nu$ the frequency modifier.
 For each mode, the frequency is thus multiplied, while the overall amplitude is damped by a given persistence.
-In the end, we simulate a large set of bulk particles without any walls, and use the noise function to remove particles that are not in a procedural void.
+In the end, we simulate a large set of bulk particles without any walls, and use the noise function to remove particles that are not in a procedural void given by a threshold $\tau$.
+Three examples of this model are depicted in figure \ref{fig:model_example}, where we have used $\Omega=6$ ocataves and the scale is $\nu=0.01$ and cutoff $\tau=0$.
+The left image depicts low persistence ($\psi = 0$), where only the first octave dominates (corresponding to large scales).
+The middle picture shows the same parameters with a higher persistence $(\psi = 0.5)$, where smaller scales become more dominating.
+In the right picture, the cutoff threshold $\tau = -0.3$.
 
-\subsection{Simulations}
-LAMMPS $SiO_2$ simulation heated -> expand box -> cool down
+\begin{figure*}
+\includegraphics[width=.95\textwidth]{model_examples.png}
+\caption{Examples of 3D simplex noise structures in nanoporous media.
+Left image: low persistence (large scale dominates), middle image: medium persistence (small scales starts dominating) and right image: high threshold (more atoms are discarded).
+}
+\label{fig:model_example}
+\end{figure*}
+
+\subsection{Model Measure}
+In order to perform data analysis on the model parameters, we developed a code that uses Monte Carlo Markov chains in order to produce samples for building the likelihood.
+Before doing so, we need a model measure - a quantitative method for uniquely identifying the interrnal structure of the porous media.
+Say something about pore distribution etc.
+
+\subsubsection{Distance-to-atom (DTA)}
+ The Distance-To-Atom (DTA) measure is a 1D histogram that measures the average distance between any point and the nearest wall in a 3D medium. HAFREAGER SKRIV
+
+%\subsubsection{Fractal Dimension}
+%The fractal dimension is a ratio providing a statistical index of complexity comparing how detail in a pattern changes with the scale at which it is measured. We provide a fracal
+
+%\subsubsection{Octree Counting Measure (OCM)}
+
+%\subsection{Diffusion Measure}
+
+\subsection{Likelihood}
+The likelihood $\mathcal{L}$ of paremeters $\theta$ of a model $M(\theta)$ given outcome $x$ is defined as $\mathcal{L} (M(\theta) | x) = P( x | M(\theta))$, where P is the probability distribution.
+In our model, we use the $\chi^2$ test to fit our one-dimensional measure from the data $\bf d$ with our model $\bf M(\theta)$.
+The $\chi^2$ is defined as
+\begin{equation}
+  \chi^2 = \sum_i \frac{ \Big(d_i - M(\theta)_i \Big)^2}{\sigma_i^2}
+\end{equation}    
+where $\sigma$ is the standard deviation. Our likelihood then becomes
+\begin{equation}
+\mathcal L \propto e^{-\chi^2}.
+\end{equation}
+For a model with $N$ free parameters, we build the full $N$-dimensional likelihood by letting random walkers map out the parameter space through the Monte Carlo Markov chain method (MCMC).
+For each step $M(\theta)_n$ in parameter space, we calculate the likelihood for this parameter configuration.
+Then, we create a new set of parameters for the next step $M(\theta)_{n+1}$ by choosing new parameters randomly from a $N$-dimensional sphere around the previous parameters.
+We calculate the likelihood at the new step and perform the MCMC test: Accept the new step if $\frac{\mathcal L_n}{\mathcal L_n+1} \ge 1$, or reject if the value is lower than a uniform random value $\in [0,1]$.
+
+\subsection{Model verification}
+In order to verify our likelihood code, we perform a model verification test on a mock data set with known input values.
+In this example, we only use standard simplex noise with two free parameters: the persistence $\psi$ and threshold $\tau$.
+We simulate a data set of $??
+^3 Å$ of $1 000 000$ atoms, and perform the simplex cutoff given input parameters $\Omega=6$, $\Psi = 0.6$, $\tau=-0.05$, $\nu=0.01$.
+The resulting 3D structure is depicted in figure \ref{fig:mockdata}.
 \begin{figure}
-\includegraphics[width=.45\textwidth]{pores1.png}
-%#\includegraphics[width=.5\textwidth]{pores2.eps}
-%\includepdf{ pores1.pdf }
-\caption{LAMMPS $SiO_2$ simulation.}
-\label{fig:model12}
+\includegraphics[width=.45\textwidth]{model_test.png}
+\caption{Simulated mock data. }
+\label{fig:mockdata}
 \end{figure}
 
-\subsection{Implementation}
-InfiniteDiffusion - nanopores. 
-Octree
+We continue by performing a full MCMC analysis of the mock data with using a parallelised code running on a 48 core cluster using 10000 random vectors for the DTA measure function.
+In order to produce 300 000 samples, about 4 hours was needed.
+The results from the analysis can be seen in figure \ref{fig:mockdataresults}, where the right-hand image depicts the correlation between $\tau$ and $\psi$ at $r=-0.7$.
+The center and left image shows that the input parameters of $\tau=-0.05$(threshold) and $\psi=0.6$(persistence) are reproduced well within $1 \sigma$.
 
-\subsection{Model Measure}
-In order to compare our procedural noise model with actual simulations, we define various measures that quantify the internal structure of the 
-\subsubsection{Distance-to-atom (DTA)}
-\subsubsection{Fractal Dimension}
-\subsubsection{Octree Counting Measure (OCM)}
-
-\subsection{Diffusion Measure}
+\begin{figure*}
+\includegraphics[width=.99\textwidth]{mock_data_results.png}
+\caption{Results from the model validation simulation.
+The left-hand image displays the 2D likelihood function of $\tau$ and $\psi$, with $1$ and $2\sigma$ contours.
+The correlation between these parameters are $r=-0.7$.
+The input parameters of $\tau=-0.05$ and $\psi=0.6$ are reproduced accurately well within $1\sigma$.
+}
+\label{fig:mockdataresults}
+\end{figure*}
 
 \begin{figure}
-\includegraphics[width=.5\textwidth]{model1.eps}
+%\includegraphics[width=.5\textwidth]{model1.eps}
 \caption{blahblah}
@@
 
-\subsection{Model measure}
-
-\subsection{Diffusion $R^2$}
-\begin{figure}
-\includegraphics[width=.5\textwidth]{compare.png}
-\caption{Results. }
-\label{fig:model12}
-\end{figure}
+\subsection{Simulated data}
 

### 2017-01-18 ms.tex
New stuff

@@
 \section{Introduction}
-In recent years, computational ...
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Duis consequat turpis vel mollis accumsan.
+Etiam eleifend magna vel quam porta, sed porta nibh placerat.
+Sed in faucibus risus.
+Duis suscipit tincidunt leo, nec laoreet leo elementum elementum.
+Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
+Sed eros risus, convallis sit amet turpis a, dapibus accumsan quam.
+Donec fringilla rhoncus turpis, quis ultrices odio vulputate elementum.
+Cras nec viverra urna.
+Phasellus laoreet tortor sed est sagittis laoreet.
+Nullam a volutpat ex.
+
+Mauris vestibulum sit amet libero eget facilisis.
+Vivamus a nunc sed felis ultrices convallis luctus et mauris.
+Sed tincidunt mauris nunc, vehicula placerat eros faucibus quis.
+Morbi egestas, nisi quis luctus dignissim, mauris odio pulvinar tellus, non efficitur risus erat vel nunc.
+Cras vel felis magna.
+Mauris rutrum magna a interdum sodales.
+Nam tempor, diam non sodales tristique, nulla ipsum eleifend dui, non aliquam felis tellus id enim.
+Donec eget nulla leo.
+
+Ut hendrerit cursus libero, in scelerisque tortor sollicitudin accumsan.
+Nam eleifend velit metus, quis volutpat justo faucibus ut.
+Cras ut sem in nunc fringilla egestas vitae sit amet justo.
+Ut vel condimentum tortor.
+Morbi in massa in ipsum ultrices tincidunt.
+Duis feugiat dignissim nisi ac ultricies.
+Quisque eget risus fermentum, condimentum massa ullamcorper, posuere neque.
+Nam congue consequat mi, vel commodo mi tempus eget.
+
+Nullam ultrices, velit sed venenatis ultrices, risus urna ultricies magna, at pretium lorem nibh et nunc.
+Suspendisse sed mattis leo.
+Nulla facilisi.
+Praesent finibus lobortis leo, ut porttitor massa tincidunt ut.
+Vivamus quis orci eu sem sodales ornare.
+Cras sed faucibus velit.
+Phasellus scelerisque mauris quis lorem dignissim vehicula.
+Sed tempor ut urna sed bibendum.
+Donec et nibh pulvinar, commodo magna ac, interdum nibh.
+Sed laoreet ipsum quis quam lacinia, in dictum erat laoreet.
+Etiam rhoncus mauris arcu, ac egestas lacus facilisis nec.
+Pellentesque a orci vitae felis faucibus ultrices sed non massa.
+In posuere vel ex in elementum.
+Aenean ac neque a enim gravida tristique eu ac lacus.
+Ut tincidunt venenatis lectus, et maximus mauris.
+
+Proin ornare nulla id tristique placerat.
+Integer nec elementum tellus.
+Cras at tempus lectus.
+Cras lorem est, ornare id condimentum vitae, scelerisque eu lectus.
+Aenean libero arcu, consequat vel sagittis eget, sagittis ut neque.
+Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
+Aenean nec quam sodales, mollis est sit amet, vulputate turpis.
+Praesent tempus diam eu sem laoreet finibus.
+Nulla finibus, sapien vel lobortis ultricies, mauris ipsum aliquet urna, quis rutrum sapien dui id ipsum.
+Quisque sodales ex sapien, ac egestas lacus suscipit et.
+Suspendisse accumsan rhoncus tristique.
+Pellentesque ut eros lectus.
+Praesent viverra sit amet odio nec vulputate.
+Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
+Nulla molestie metus vitae libero vulputate, at interdum enim tempus.
 
@@
 
+\begin{figure}[htb!]
+\includegraphics[width=.45\textwidth]{DTA.png}
+\caption{Distance-to-atom function for two data sets with varying persistence.
+The green line corresponds to a mock data set with low persistence (dominated by large-scale modes), while the blue line corresponds to high persistence (dominated by small scale modes).
+}
+\label{fig:dta_example}
+\end{figure}
+
 %\subsubsection{Fractal Dimension}
@@
 \subsection{Model verification}
+\begin{figure}[htb!]
+\includegraphics[width=.45\textwidth]{model_test.png}
+\caption{Simulated mock data. }
+\label{fig:mockdata}
+\end{figure}
 In order to verify our likelihood code, we perform a model verification test on a mock data set with known input values.
@@
 The resulting 3D structure is depicted in figure \ref{fig:mockdata}.
-\begin{figure}
-\includegraphics[width=.45\textwidth]{model_test.png}
-\caption{Simulated mock data. }
-\label{fig:mockdata}
-\end{figure}
-
-We continue by performing a full MCMC analysis of the mock data with using a parallelised code running on a 48 core cluster using 10000 random vectors for the DTA measure function.
-In order to produce 300 000 samples, about 4 hours was needed.
-The results from the analysis can be seen in figure \ref{fig:mockdataresults}, where the right-hand image depicts the correlation between $\tau$ and $\psi$ at $r=-0.7$.
-The center and left image shows that the input parameters of $\tau=-0.05$(threshold) and $\psi=0.6$(persistence) are reproduced well within $1 \sigma$.
 
@@
 
-\begin{figure}
-%\includegraphics[width=.5\textwidth]{model1.eps}
-\caption{blahblah}
-\label{fig:model12}
-\end{figure}
+We continue by performing a full MCMC analysis of the mock data with using a parallelised code running on a 48 core cluster using 10000 random vectors for the DTA measure function.
+In order to produce 300 000 samples, about 4 hours was needed.
+The results from the analysis can be seen in figure \ref{fig:mockdataresults}, where the right-hand image depicts the correlation between $\tau$ and $\psi$ at $r=-0.7$.
+The center and left image shows that the input parameters of $\tau=-0.05$(threshold) and $\psi=0.6$(persistence) are reproduced well within $1 \sigma$.
 
 \section{Results}
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Duis consequat turpis vel mollis accumsan.
+Etiam eleifend magna vel quam porta, sed porta nibh placerat.
+Sed in faucibus risus.
+Duis suscipit tincidunt leo, nec laoreet leo elementum elementum.
+Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
+Sed eros risus, convallis sit amet turpis a, dapibus accumsan quam.
+Donec fringilla rhoncus turpis, quis ultrices odio vulputate elementum.
+Cras nec viverra urna.
+Phasellus laoreet tortor sed est sagittis laoreet.
+Nullam a volutpat ex.
 
+Mauris vestibulum sit amet libero eget facilisis.
+Vivamus a nunc sed felis ultrices convallis luctus et mauris.
+Sed tincidunt mauris nunc, vehicula placerat eros faucibus quis.
+Morbi egestas, nisi quis luctus dignissim, mauris odio pulvinar tellus, non efficitur risus erat vel nunc.
+Cras vel felis magna.
+Mauris rutrum magna a interdum sodales.
+Nam tempor, diam non sodales tristique, nulla ipsum eleifend dui, non aliquam felis tellus id enim.
+Donec eget nulla leo.
+
+Ut hendrerit cursus libero, in scelerisque tortor sollicitudin accumsan.
+Nam eleifend velit metus, quis volutpat justo faucibus ut.
+Cras ut sem in nunc fringilla egestas vitae sit amet justo.
+Ut vel condimentum tortor.
+Morbi in massa in ipsum ultrices tincidunt.
+Duis feugiat dignissim nisi ac ultricies.
+Quisque eget risus fermentum, condimentum massa ullamcorper, posuere neque.
+Nam congue consequat mi, vel commodo mi tempus eget.
+
+Nullam ultrices, velit sed venenatis ultrices, risus urna ultricies magna, at pretium lorem nibh et nunc.
+Suspendisse sed mattis leo.
+Nulla facilisi.
+Praesent finibus lobortis leo, ut porttitor massa tincidunt ut.
+Vivamus quis orci eu sem sodales ornare.
+Cras sed faucibus velit.
+Phasellus scelerisque mauris quis lorem dignissim vehicula.
+Sed tempor ut urna sed bibendum.
+Donec et nibh pulvinar, commodo magna ac, interdum nibh.
+Sed laoreet ipsum quis quam lacinia, in dictum erat laoreet.
+Etiam rhoncus mauris arcu, ac egestas lacus facilisis nec.
+Pellentesque a orci vitae felis faucibus ultrices sed non massa.
+In posuere vel ex in elementum.
+Aenean ac neque a enim gravida tristique eu ac lacus.
+Ut tincidunt venenatis lectus, et maximus mauris.
+
+Proin ornare nulla id tristique placerat.
+Integer nec elementum tellus.
+Cras at tempus lectus.
+Cras lorem est, ornare id condimentum vitae, scelerisque eu lectus.
+Aenean libero arcu, consequat vel sagittis eget, sagittis ut neque.
+Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
+Aenean nec quam sodales, mollis est sit amet, vulputate turpis.
+Praesent tempus diam eu sem laoreet finibus.
+Nulla finibus, sapien vel lobortis ultricies, mauris ipsum aliquet urna, quis rutrum sapien dui id ipsum.
+Quisque sodales ex sapien, ac egestas lacus suscipit et.
+Suspendisse accumsan rhoncus tristique.
+Pellentesque ut eros lectus.
+Praesent viverra sit amet odio nec vulputate.
+Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
+Nulla molestie metus vitae libero vulputate, at interdum enim tempus.
 \subsection{Simulated data}

### 2017-01-18 ms.tex
Wrote stuff

@@
 In order to perform data analysis on the model parameters, we developed a code that uses Monte Carlo Markov chains in order to produce samples for building the likelihood.
-Before doing so, we need a model measure - a quantitative method for uniquely identifying the interrnal structure of the porous media.
-Say something about pore distribution etc.
+Before doing so, we need a model measure - a quantitative method for uniquely identifying the internal structure of the porous media.
+This quantity should be computed for both the original data set and the one produced by our noise model.
 
-\subsubsection{Distance-to-atom (DTA)}
- The Distance-To-Atom (DTA) measure is a 1D histogram that measures the average distance between any point and the nearest wall in a 3D medium. HAFREAGER SKRIV
+Ideally, the model measure should pick up all geometrical properties that are relevant for the physical study of interest.
+For instance, pore size distribution and porosity are import for studying transport in porous media (REF).
+One such measure is the Distance-To-Atom (DTA).
+
+\subsubsection{Distance-to-atom}
+The DTA measure is the distribution of distances between $N_r$ random points and their nearest wall atom defined by the geometry.
+We get a list of $N_r$ such distances which can be histogrammed.
+By computing this histogram for both the data and the model, we can compare the two as seen in figure \ref{fig_dta:example}.
+
+\subsubsection{Pore size distribution}
+A common method of characterizing the geometry is the pore size distribution as first used in \cite{gelb1999pore}.
+It works by filling the pore space with spheres, starting with the largest possible.
+After each sphere, the pore space is now considered filled and we repeat this process until the next sphere has radius smaller than some $R_\text{min}$.
 
@@
 \includegraphics[width=.45\textwidth]{DTA.png}
-\caption{Distance-to-atom function for two data sets with varying persistence.
+\caption{Distance-to-atom measure for two data sets with varying persistence.
 The green line corresponds to a mock data set with low persistence (dominated by large-scale modes), while the blue line corresponds to high persistence (dominated by small scale modes).

### 2017-01-18 plot_dta.m

@@ -1,0 +1,10 @@
+model = dlmread('/projects/NanoPores/build-NanoPores-Desktop_Qt_5_7_1_clang_64bit-Release/NanoPores/NanoPores.app/Contents/MacOS/model.txt');
+data = dlmread('/projects/NanoPores/build-NanoPores-Desktop_Qt_5_7_1_clang_64bit-Release/NanoPores/NanoPores.app/Contents/MacOS/data.txt');
+plot(model(:,1), model(:,2), 'LineWidth', 2.0)
+hold on
+plot(data(:,1), data(:,2), 'LineWidth', 2.0)
+xlabel('d [\AA]','Interpreter','LaTeX');
+ylabel('P(d)');
+title('Distance-to-atom');
+legend({'Model', 'Data'});
+set(gca,'FontSize', 18)

### 2017-01-24 ms.tex
Added things about porosity
Added surface are
Updated stuff
whooooooo

--- 2017-01-18 ms.tex
@@
 \usepackage{epstopdf}
+\usepackage{color}
+\definecolor{Blue}{rgb}{0.1,0.1,1.0} 
+\definecolor{Magenta}{rgb}{1.0,0.1,0.5} 
+\definecolor{LRed}{rgb}{0.8,0.0,0.0}
+
+ \newcommand{\todo}[1]{ {\color{Magenta} TODO: \color{Blue} \textbf{#1} }}
+
 \begin{document}
@@
 \begin{figure}
-\includegraphics[width=.5\textwidth]{perlinstats.eps}
+\includegraphics[width=.5\textwidth]{noise_comparison.png}
 %\mbox{\epsfig{file=perlinstats.eps, width=\linewidth}}
@@
   Gaussian distribution. Note how both noise functions produce
-  near-Gaussian, but slightly offset distributions.}
+  near-Gaussian, but slightly different distributions.}
 \label{fig:properties}
@@
 
-\ref{burganos1998gas}
+%\ref{burganos1998gas}
 
-The generic noise model presented in \ref{eq:procedural} can easily be extrapolated to 3 dimensions. In Lammps, we
-simulate a uniform box of N $SiO_2$ particles without constraints, and proceed by cutting away particles above a given threshold $T$.
 %\begin{figure}
@@
 
-\section{Model}
-We construct a generic sum over procedural noise octaves with the expression: 
-\begin{equation}
-  M_1(\vec x) = \sum_{\omega=0}^{\Omega} \psi^{(\frac{1}{2})^{\omega +1}}   N\big((\vec x + \sigma \cdot \omega)\nu \cdot 2^\omega \big),
-\label{eq:noisemodel1}
-\end{equation}
-where $\Omega$ is the total amount of octaves, $\psi$ the persistence (scale falloff), $N$ a generic $N$-dimensional noise function, $\sigma$ a predefined random vector that removes anisotropic artifacts and $\nu$ the frequency modifier.
-For each mode, the frequency is thus multiplied, while the overall amplitude is damped by a given persistence.
-In the end, we simulate a large set of bulk particles without any walls, and use the noise function to remove particles that are not in a procedural void given by a threshold $\tau$.
-Three examples of this model are depicted in figure \ref{fig:model_example}, where we have used $\Omega=6$ ocataves and the scale is $\nu=0.01$ and cutoff $\tau=0$.
-The left image depicts low persistence ($\psi = 0$), where only the first octave dominates (corresponding to large scales).
-The middle picture shows the same parameters with a higher persistence $(\psi = 0.5)$, where smaller scales become more dominating.
-In the right picture, the cutoff threshold $\tau = -0.3$.
+\subsection{Model}
+In this paper, we are interested in developing a model using Simplex noise to reproduce nanoporous structures that are as statistically close to actual simulations of porous media.
+For a given data set and a model, we wish to estimate the parameters that best represent the data.
+This is performed through a full likelihood analsys, where we in addition validate the model by comparing the radial pore size distribution $g(r)$, porosity and surface area of the system.
 
@@
 \end{figure*}
+We construct a generic sum over procedural noise octaves with the expression: 
+\begin{equation}
+  M_1(\vec x) = \sum_{\omega=0}^{\Omega} \psi^{\omega -1}   N\big((\vec x + \sigma \cdot \omega)\nu \cdot 2^{\omega-1} \big),
+\label{eq:noisemodel1}
+\end{equation}
+where $\Omega$ is the total amount of octaves, $\psi$ the persistence (scale falloff), $N$ a generic $N$-dimensional noise function, $\sigma$ a predefined random vector that removes anisotropic artifacts and $\nu$ the frequency modifier representing the initial largest scales in the model.
+For each mode, the frequency is thus multiplied, while the overall amplitude is damped by a given persistence.
+In the end, we simulate a large set of bulk particles without any walls, and use the noise function to remove particles that are not in a procedural void given by a threshold $\tau$.
+Three examples of this model are depicted in figure \ref{fig:model_example}, where we have used $\Omega=6$ ocataves and the scale is $\nu=0.01$ and cutoff $\tau=0$.
+The left image depicts low persistence ($\psi = 0.0$), where only the first octave dominates (corresponding to large scales).
+The middle picture shows the same parameters with a higher persistence $(\psi = 0.5)$, where smaller scales become more dominating.
+In the rightmost picture, the effect of the cutoff threshold $\tau = -0.3$ is shown.
 
@@
 In order to perform data analysis on the model parameters, we developed a code that uses Monte Carlo Markov chains in order to produce samples for building the likelihood.
-Before doing so, we need a model measure - a quantitative method for uniquely identifying the interrnal structure of the porous media.
-Say something about pore distribution etc.
+Before doing so, we need a model measure - a quantitative method that characterizes the internal structure of the porous media.
+This quantity should be computed for both the original data set and the one produced by our noise model.
 
-\subsubsection{Distance-to-atom (DTA)}
- The Distance-To-Atom (DTA) measure is a 1D histogram that measures the average distance between any point and the nearest wall in a 3D medium. HAFREAGER SKRIV
+Ideally, the model measure should pick up all geometrical properties that are relevant for the physical study of interest.
+For instance, pore size distribution and porosity are import for studying transport in porous media.
+Another such measure is the Distance-To-Atom (DTA).
+The method used for calculating the porosity of a media is defined as in \cite{gelb1998characterization}, and we obtain the surface area of a porous media by using marching cubes triangulation to sum the surface area of the triangles.
 
-\begin{figure}[htb!]
-\includegraphics[width=.45\textwidth]{DTA.png}
-\caption{Distance-to-atom function for two data sets with varying persistence.
+In this paper, we choose to use the radial pore size distribution $g(r)$ as defined and implemented in LAMMPS \cite{plimpton1995fast}.  \todo{skriv litt mer om hvorfor vi velger g(r)}
+
+%\begin{figure}[htb!]
+%\includegraphics[width=.45\textwidth]{DTA.png}
+%\caption{Distance-to-atom measure for two data sets with varying persistence.
 The green line corresponds to a mock data set with low persistence (dominated by large-scale modes), while the blue line corresponds to high persistence (dominated by small scale modes).
 }
-\label{fig:dta_example}
-\end{figure}
+%\label{fig:dta_example}
+%\end{figure}
 
@@
 In our model, we use the $\chi^2$ test to fit our one-dimensional measure from the data $\bf d$ with our model $\bf M(\theta)$.
-The $\chi^2$ is defined as
+Pearson's $\chi^2$ is defined as
 \begin{equation}
-  \chi^2 = \sum_i \frac{ \Big(d_i - M(\theta)_i \Big)^2}{\sigma_i^2}
+  \chi^2 = \sum_i \frac{ \Big(d_i - M(\theta)_i \Big)^2}{M(\theta)_i^2},
 \end{equation}    
-where $\sigma$ is the standard deviation. Our likelihood then becomes
+and the likelihood is thus
 \begin{equation}
@@
 Then, we create a new set of parameters for the next step $M(\theta)_{n+1}$ by choosing new parameters randomly from a $N$-dimensional sphere around the previous parameters.
+The model also chooses a new random seed.
 We calculate the likelihood at the new step and perform the MCMC test: Accept the new step if $\frac{\mathcal L_n}{\mathcal L_n+1} \ge 1$, or reject if the value is lower than a uniform random value $\in [0,1]$.
 
-\subsection{Model verification}
+\section{Model verification and data analysis}
 \begin{figure}[htb!]
 \includegraphics[width=.45\textwidth]{model_test.png}
-\caption{Simulated mock data. }
+\caption{Simulated mock data used for the initial model testing analysis. Input parameters are $\nu=0.03$,$\tau=-0.02$ and $\psi=0.6$.}
 \label{fig:mockdata}
@@
 In order to verify our likelihood code, we perform a model verification test on a mock data set with known input values.
-In this example, we only use standard simplex noise with two free parameters: the persistence $\psi$ and threshold $\tau$.
-We simulate a data set of $??
-^3 Å$ of $1 000 000$ atoms, and perform the simplex cutoff given input parameters $\Omega=6$, $\Psi = 0.6$, $\tau=-0.05$, $\nu=0.01$.
+In this example, we only use standard simplex noise with three free parameters: the persistence $\psi$, threshold $\tau$ and scale $\nu$.
+We simulate a data set of $159 ^3 Å$ of $1 000 000$ atoms, and perform the cutoff given input parameters $\Omega=3$, $\Psi = 0.6$, $\tau=-0.2$, $\nu=0.03$.
 The resulting 3D structure is depicted in figure \ref{fig:mockdata}.
 
+We continue by performing a full MCMC analysis of the mock data with using a parallelised code running on a 48 core cluster.
+In order to produce 300 000 samples, about 48 hours was needed.
+The results from this analysis can be seen in figure \ref{fig:mockdataresults_pts}, where the upper panel depicts the marginalised 1D posteriors while the lower shows the marginalised 2D posteriors.
+The center and left image shows that the input parameters of $\tau=-0.2$(threshold), $\psi=0.6$(persistence) and $\nu=0.03$ are all reproduced well within $1 \sigma$.
+
 \begin{figure*}
-\includegraphics[width=.99\textwidth]{mock_data_results.png}
+\includegraphics[width=.99\textwidth]{mock_data_results_pts.png}
 \caption{Results from the model validation simulation.
-The left-hand image displays the 2D likelihood function of $\tau$ and $\psi$, with $1$ and $2\sigma$ contours.
-The correlation between these parameters are $r=-0.7$.
-The input parameters of $\tau=-0.05$ and $\psi=0.6$ are reproduced accurately well within $1\sigma$.
-}
-\label{fig:mockdataresults}
+The upper panel displays the marginalised 1D posteriors for each parameter, while the lower panel depicts the marginalised 2D posteriors.
+Note that all input parameters were reproduced well within $1\sigma$ (inner contour line).}
+\label{fig:mockdataresults_pts}
 \end{figure*}
 
-We continue by performing a full MCMC analysis of the mock data with using a parallelised code running on a 48 core cluster using 10000 random vectors for the DTA measure function.
-In order to produce 300 000 samples, about 4 hours was needed.
-The results from the analysis can be seen in figure \ref{fig:mockdataresults}, where the right-hand image depicts the correlation between $\tau$ and $\psi$ at $r=-0.7$.
-The center and left image shows that the input parameters of $\tau=-0.05$(threshold) and $\psi=0.6$(persistence) are reproduced well within $1 \sigma$.
+\section{Results}
+%\begin{figure}
+%\includegraphics[width=.49\textwidth]{sio_porous.png}
+%\caption{Simulated silica created with LAMMPS.
+The box is $159^3 \AA$, with $100 000$ particles, having been created by cooling an expanding system until quenching was achieved.
+Note the similarities between the pores here and in figure \ref{fig:mockdata}.}
+%\label{fig:si2data}
+%\end{figure}
+\begin{figure*}
+\includegraphics[width=.99\textwidth]{results_porous_full.png}
+\caption{Results from the analysis of a simulated $159^3$ \AA $SiO^2$ nanoporous data set.
+The upper panel displays the marginalised 1D posteriors for each parameter, while the lower panel depicts the marginalised 2D posteriors.
+Best fit parameters from the maximum 3D likelihood suggest $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$.}
+\label{fig:porous_results1}
+\end{figure*}
 
-\section{Results}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Duis consequat turpis vel mollis accumsan.
-Etiam eleifend magna vel quam porta, sed porta nibh placerat.
-Sed in faucibus risus.
-Duis suscipit tincidunt leo, nec laoreet leo elementum elementum.
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
-Sed eros risus, convallis sit amet turpis a, dapibus accumsan quam.
-Donec fringilla rhoncus turpis, quis ultrices odio vulputate elementum.
-Cras nec viverra urna.
-Phasellus laoreet tortor sed est sagittis laoreet.
-Nullam a volutpat ex.
+We now wish to perform a parameter fitting analysis on realistic simulated nanoporous media.
+Using LAMMPS \cite{plimpton1995fast}, we simulate a $100^3$ \AA box of 100K $SiO^2$ silica particles at an initial temperature of $4500 K$, using the Vashishta-potential.
+After an initial period, we expand the box to $159^3 \AA$ and start cooling until quenching is achieved at about $300K$.
+The resulting nanoporous media is depicted in figure \ref{fig:sio2data}, with a porosity of $75\%$.
 
-Mauris vestibulum sit amet libero eget facilisis.
-Vivamus a nunc sed felis ultrices convallis luctus et mauris.
-Sed tincidunt mauris nunc, vehicula placerat eros faucibus quis.
-Morbi egestas, nisi quis luctus dignissim, mauris odio pulvinar tellus, non efficitur risus erat vel nunc.
-Cras vel felis magna.
-Mauris rutrum magna a interdum sodales.
-Nam tempor, diam non sodales tristique, nulla ipsum eleifend dui, non aliquam felis tellus id enim.
-Donec eget nulla leo.
+With the model described in \ref{eq:noisemodel1}, we perform a full likelihood analysis of the data set with the three free parameters $\tau$, $\nu$ and $\psi$ using $g(r)$ as the measure.
+The resulting marginalised posteriors are depicted in figure \ref{fig:porous_results1}, with best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$ is extracted from the full 3D likelihood.
+We also display the radial pore size distribution $g(r)$ for the input data set, a best-fit parameter data set and a bad-fit data set in figure \ref{fig:gofr1}, to illustrated the sensitivity of $g(r)$ given the model parameters.
+Note that we are not able to accurately reproduce the structures of the $SiO^2$ $g(r)$ with our current model, but hope that we during future work will be able to produce another model that provides a better fit with this measure.
+
+\begin{figure}
+\includegraphics[width=.55\textwidth]{gofr_plot.png}
+\caption{The radial pore size distribution $g(r)$ for three data sets: the simulated $SiO^2$ input data set, the best-fit parameter data set and a random low-fit parameter data set. }
+\label{fig:gofr1}
+\end{figure}
+Finally, we depict a comparison of the actual simulated input $SiO^2$ data set together with a representation of the best-fit parameters in figure \ref{fig:porous_vs_model}   
+\begin{figure}
+\includegraphics[width=.45\textwidth]{comparison.png}
+\caption{Right side: A realisation of the best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$. Left side: Simulated $SiO^2$ input data set used for parameter estimation. }
+\label{fig:porous_vs_model}
+\end{figure}
+
+The porosity of the input $SiO^2$ data set is $75\%$, while the best-fit data set has porosity $77\% \pm 3 \%$, depending on the random seed.
+In addition, the surface area of the original data set was $\tilde 29 000 \AA^2$, while the best-fit data shows similar values with surface are ranging betweem $\tilde 28 000 \AA^2$ and $\tilde 30 000 \AA^2$.
+
+\subsection{Some notes on the results}
+It is worth to note that since we are chiefly interested in obtaining a set of parameters that reproduces a statistical correct representation of the input data, we are relatively free to choose parameters as desired.
+In the previous analysis, we selected a best fit threshold value of $\tau=-0.2$, but from the width of the likelihood it is clear that nearby parameters would also be suitable.
+However, the surface area is strongly correlated with this value, so even slight deviations (e.g.
+$\delta \tau = 0.05$) yields a relatively large change in surface area.
+
+In addition, we do not claim that this model accurately reproduces the pore structure observed in nanoporous simulations, as is evident from the radial pore size distribution shown in figure \ref{fig:gofr1}.
+
+%persistence : 0.875111019125
+%scale : 0.036638
+%threshold : -0.236784875   78% porosity
+
+% bad fit : scale 0.1, p = 1, threshold = -0.3
+
+\section{Conclusion}
+In this paper, we have proposed a novel method for generating a simple kind of nanoporous media using procedural methods, namely Simplex noise.
+We have developed tools for creating such porous materials, and also estimating model parameters given data.
+We have shown that it is possible to recreate
+
+\subsection{Future work}
+
+\begin{figure*}
+\includegraphics[width=1\textwidth]{future_models.png}
+\caption{Example of future models.
+Left panel: A multifractal model with thin surfaces and large pores.
+Center panel: a multifractal with cave-like structures.
+Right panel: A model with strong asymmetric noise.
+}
+\label{fig:future_models}
+\end{figure*}
 
@@
 
-\section{Conclusion}
-It works! Hopefully.
-\subsection{Future work}
-Stress / shear test both models. Etc.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-01-24 org.tex
whooooooo

\documentclass[aps,pre,twocolumn,letterpaper,floatfix,showpacs]{revtex4}
\usepackage{graphicx} 
\usepackage{amsmath,amssymb,amsfonts} 
\usepackage{mathtools}
\usepackage{afterpage}
\usepackage[hidelinks]{hyperref} 
\usepackage{epstopdf}
\begin{document}
\title{Transient cluster formation in sheared non-Brownian suspensions}
\author{Kjetil Th\o gersen $^1$}
\email{kjetil.thogersen@fys.uio.no}
\author{Marcin Dabrowski $^{1,2}$}
\author{Anders Malthe-S\o renssen $^1$}
\affiliation{$^1$Department of Physics - University of Oslo\\Sem S{\ae}lands vei 24, NO-0316, Oslo, Norway  \\
$^2$Polish Geological Institute - National Research Institute\\ Wroclaw, Poland}
\date{\today}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
We perform numerical simulations of non-Brownian suspensions in the laminar flow regime to study the scaling behavior of particle clusters and collisions under shear. As the particle fraction approaches the maximum packing fraction, large transient clusters appear in the system. We use methods from percolation theory to discuss the cluster size distribution. We also give a scaling relation for the percolation threshold as well as system size effects through time-dependent fluctuations of this threshold and relate them to system size. System size effects are important close to the maximum packing fraction due to the divergence of the cluster length scale. We then investigate the transient nature of the clusters through characterization of particle collisions, and show that collision times exhibit scale invariant properties. We then show that particle collision times can be modeled as first passage processes.
\end{abstract}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Suspensions are ubiquitous in nature. They are important in fields ranging from swimming bacteria \cite{lushi2014fluid} and blood flow \cite{freund2014numerical}, to lava flows \cite{yamato2012numerical,glazner2014magmatic}, debris flows \cite{coussot1996recognition} and subsurface contaminant transport \cite{mccarthy1989subsurface}. Understanding suspensions is also important for several industrial and technological applications, e.g for the gas industry \cite{meakin2013shale,patankar2002power,kuzkin2014computer} and in microfluidics \cite{whitesides2006origins,xuan2010particle}. Since Einstein's work on the effective viscosity of dilute suspensions \cite{einstein1906eineneue,einstein1911berichtigung}, a large number of studies on particle suspensions have been performed.  In 1972, Batchelor and Green extended Einsteins expression of effective viscosity to second order \cite{batchelor1972determination}.  The model of Batchelor and Green has since been thoroughly tested experimentally for low volume fractions ($\phi<0.1$), and the rheology of dilute suspensions is now well understood \cite{ovarlez2006local}. The rheology of dense suspensions is less trivial. Several phenomenological models have been proposed, including the classical Krieger-Dougherty relation where the viscosity diverges at a maximum packing fraction $\phi_m$ \cite{krieger1959mechanism}.

A number of studies have related rheology to the interaction between close particle pairs \cite{batchelor1972determination,frankel1967viscosity,marrucci1985viscosity,iskakova2013shear,sierou2002rheology,brady1997microstructure,mills2009apparent}. Frankel and Acrivos stated that the dominant interaction mode is the compression mode of approaching particles \cite{frankel1967viscosity}, and developed a constitutive relation for effective viscosity for high particle volume fractions.  
Marucci and Denn later pointed out that the tangential mode is dominating, but they also stated that correlated motion of over length scales larger than a single particle, i.e. the formation of particle clusters \cite{marrucci1985viscosity}. The decrease of the size of the lubrication layer between close particle distances with shear rate has also been used to demonstrate shear induced thickening \cite{iskakova2013shear}.
Several studies have been performed that include interparticle forces. Brady et al. have performed several studies of suspension rheology of suspensions using a short range repulsive force accounting for particle roughness (e.g. \cite{brady1997microstructure,sierou2002rheology}). Recent studies have also explained discontinuous shear thickening as the sudden increase of frictional contacts with shear rate \cite{seto2013discontinuous}.  Studies have also been carried out to unify suspension and granular rheology close to the jamming threshold \cite{boyer2011unifying}, as well as to understand the rheology of suspensions of non-spherical particles \cite{mueller2009rheology}

In 1979, de Gennes proposed a series of conjectures about the microstructure of sheared suspensions \cite{de1979conjectures}. Branched and highly transient particle clusters form in sheared suspensions, but there is a well defined time-independent size distribution of clusters. He also stated that an infinite cluster would appear as $\phi$ exceeds 
a critical threshold value $\phi_c$, and that the only motion allowed by such clusters would be uniform translations and rotation, and that the divergence of the cluster correlation length follows standard percolation laws.  de Gennes' conjectures were recently tested by Gallier et al. using numerical simulations based on the fictitious domain method including frictional contacts. They found that most of de Gennes' conjectures hold, but only find weak correlation between transient spanning clusters and measured effective viscosity \cite{gallier2015percolation}.

Suspension microstructure is essential to understand suspension rheology \cite{stickel2005fluid,morris2009review}. However, while dilute suspensions are fairly well understood, the relation between microstructure and rheology of dense suspensions is still lacking, and analytical treatment of microstructure evolution is complicated. The pair distribution function has been used only with limited success to relate rheology and microstructure, even though tensor descriptions of microstructure look more promising \cite{stickel2005fluid}. This calls for a different approach to describe the microstructure of suspensions.

From the transient nature of cluster formation as first proposed by de Gennes \cite{de1979conjectures}, we expect the maximum cluster size to diverge as the maximum packing fraction is approached. This means that system size effects will be increasingly important for high particle volume fractions since percolation cluster can occur well before the maximum packing fraction \cite{gallier2015percolation}. Understanding the transient nature of particle contacts and clustering could therefore provide useful insights to the overall micro-structural fluctuations of particle suspensions.

In this work we focus on the transient nature of cluster formation and breakup in suspensions of rigid particles in shear flow, as well as close particle interactions. The methods we use are motivated by percolation theory as well as theory of first passages. We define particles to be in contact below a length scale $\epsilon$, and the percolation threshold $\epsilon_c$ as the smallest $\epsilon$ that results in a particle cluster that spans the system. We establish a scaling relation between percolation and particle fraction, both for the mean percolation threshold $\langle \epsilon_c \rangle$ and the fluctuations in percolation threshold $\delta \epsilon_c$ which is system size dependent. As $\phi$ approaches the maximum packing fraction $\phi_m$, the percolation threshold decreases towards zero. For a system of finite size, we then expect transient percolating clusters well before the maximum packing fraction if we define particle contact with a small rim separation $r\epsilon$. We then show that particle contact times are invariant under the choice of $\epsilon$.

In this paper we will use a Stokes flow description. Non-Brownian suspensions can often be considered as purely viscous materials \cite{ovarlez2006local}, i.e. the viscosity does not depend on the shear rate. Even though two particles in Stokes flow show reversible dynamics, and their trajectories are symmetric, non-Brownian suspensions show rich dynamics even at zero Reynolds number. Recent studies demonstrate chaotic particle trajectories leading to a threshold of irreversibility related to the growth of the Lyapunov exponent \cite{pine2005chaos}. Chaotic behavior can be demonstrated even for three particles settling under gravity \cite{janosi1997chaotic}, and chaotic rotation has been demonstrated for a single triaxial ellipsoid \cite{yarin1997chaotic}. This failure of time-reversibility shows that chaos is an intrinsic property of slowly sheared suspensions.

The paper is structured as follows: In section \ref{section:model_description} we introduce the finite element model we use to solve for the particle trajectories. In section \ref{section:dynamics} we first introduce a length scale $\epsilon$ that defines particle contact, we then discuss the dynamics of the system and give the scaling relations with $\epsilon$ based on percolation theory, including cluster formation, and particle collisions. We then sum up and conclude in section \ref {section:discussion}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Model description} \label{section:model_description} 
\begin{figure}
\includegraphics[width=.5\textwidth]{Thogersen_fig1.eps}
\caption{Sketch of the system setup. The left and right boundaries are periodic, while the top and bottom have velocities in opposite directions to give a background shear rate $\dot \gamma$. The particle configuration is a snapshot from a simulation of $N=1024$ particles at volume fraction $\phi = 0.4$. $\epsilon = 1$. $\dot \gamma t = 30$. The colored segments are generated from a Delaunay triangulation where only segments shorter than $r(2+\epsilon)$, where $r$ is the particle radius and $\epsilon$ is a dimensionless parameter that defines contact, are kept. Here we can see that $\epsilon \geq \epsilon_c$, where $\epsilon_c$ is the smallest $\epsilon$ that gives a spanning cluster in the vertical direction. Segments that form closed triangles have a different shading to highlight the branched cluster structure.
\label{fig:bc_sketch}}
\end{figure}

While numerous numerical techniques have been used to study particle suspensions (e.g. dissipative particle dynamics \cite{hoogerbrugge1992simulating}, lattice boltzmann methods \cite{ladd2001lattice} and lagrange multiplier ficticious domain method \cite{glowinski1999distributed}), numerical investigation of particles suspensions has until recently been dominated by studies using Stokesian dynamics simulations \cite{brady1988stokesian,brady1985rheology,ekiel2008stokesian}.
In this study we perform direct numerical simulations using finite element method. Recent developments allows us to go to large enough strains and particle numbers to perform statistical analysis. We solve the incompressible Stokes equations. Conservation of mass yields
 \begin{equation}
 \nabla \cdot \vec v  = 0,
 \end{equation}
 where $\vec v$ is the vector field. Conservation of momentum yields
 \begin{equation}
 \nabla \cdot \boldsymbol \sigma = \vec 0,
 \end{equation}
 where $\boldsymbol \sigma$ is the stress tensor
 \begin{equation}
\boldsymbol \sigma = -p  \boldsymbol I +  \boldsymbol T,
\end{equation}
where $p$ is the pressure, $\boldsymbol I$ is the identity matrix, and
 \begin{equation}
\boldsymbol T = \mu \left (   \nabla \vec v + (\nabla \vec v )^T \right )
\end{equation}
is the deviatoric stress tensor. We use a an adaptive two-dimensional mixed finite element discretization with quadratic shape functions for the velocities and linear shape functions for the pressure (the Crouzeix-Raviart seven-node triangular element). We use a modified version of MILAMIN for efficient computation in MATLAB \cite{dabrowski2008milamin}, and Triangle for meshing \cite{shewchuk1996triangle}. For rigid particles we replace the degrees of freedom associated withthe particle boundaries with three degrees of freedom, two translational and one rotational.

In this paper we study the purely hydrodynamic limit, i.e. we use no repulsive forces. In order to achieve this, special care has to be taken to spatial discretization as well as time-integration \cite{ekiel2008stokesian}. We use geometrical mesh refinement to ensure that there are at least two elements across the aperture between all close particle pairs at any time during the simulation. In this way we ensure that the velocity field is accurate between close particles. For time-integration we use the second order Runge-Kutta method. We terminate the simulation if particle overlap is detected. For dense suspensions this sets limits on the maximum strains that we can achieve. For statistical measures, it would be beneficial to have one very long run rather than ensemble averaging over multiple shorter ones, but due to computational challenges, the latter is used throughout the paper. Using no repulsive force is especially problematic for large volume fractions where spanning clusters can occur, and particle distances can get infinitesimal, and hence we can not handle this numerically for arbitrary strains. Still, we present results without any artificial repulsive force up to significant strains for small to moderate volume fractions.

Additional strengths of a finite element discretization is that arbitrary domain and particle geometries are trivial to implement, and the velocity and pressure field is solved for at the same time as the particle velocities, which allows for studies of the coupling between fluid and particle motion as well as access to the stress tensor at any position in space and time. However, for simplicity we limit the study to monodisperse disks.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{System setup and dynamics} \label{section:dynamics}
\figurename~\ref{fig:bc_sketch} shows a sketch of the system setup. The left and right boundaries are periodic, while the top and bottom boundaries have Dirichlet boundary conditions of opposite velocities setting up an average shear rate $\dot \gamma$. The system is integrated forward in time with a second order Runge Kutta method with timestep $dt = \frac{0.02}{\dot \gamma}$. We simulate the fully lubricated limit, i.e. the particles always interact through a thin fluid film.

\subsection{Cluster statistics}
\begin{figure}
\includegraphics[width=.5\textwidth]{Thogersen_fig2a.eps}\\
\includegraphics[width=.5\textwidth]{Thogersen_fig2b.eps}
\caption{Top: Dependency of $\langle \epsilon_c \rangle$ on the system size $\frac{1}{N}$ and the volume fraction $\phi$. 
\emph{Main figure:} $\langle \epsilon_c \rangle$ as a function of the volume fraction (\emph{crosses}: $x$-direction, \emph{dots}: $y$-direction). $\langle\epsilon_c\rangle$ scales with $(1-\phi/\phi_m) \left ( (\phi/\phi_m)^{-1/2} - 1 \right )$ (see text), where $\phi_m = 0.81$, in the range of volume fraction we have tested. \emph{Left inset:} System size dependency of $\langle \epsilon_c\rangle$. We find no significant system size effects. \emph{Right inset:} System size effect of the fluctuations of $\epsilon_c$, here given by the standard deviation. We find that the standard deviation scales with the system size approximately as $std(\epsilon_c) \sim N^{-1/3}$.
Bottom: Percolation threshold $\epsilon_c$ as a function of strain (time) for $N=1024$ particles at volume fraction $\phi = 0.3$.
\label{fig:epsilon_c_scaling}}
\end{figure}

While many models assume a roughness parameter to determine when particles are in contact, we do not have a well defined contact distance in our system since the particles are always separated by a thin fluid film.  We define particle clusters using Delaunay triangulation of the particle positions. We define define active segments of the triangulation as segments that are shorter than
\begin{equation}
|\vec \chi_{ij}| \leq r(2+\epsilon),
\end{equation}
where $|\vec \chi_{ij}|$ is the distance between the particle centers, $r$ is the particle radius and $\epsilon$ is a dimensionless distance. The graph of active segments makes up the cluster structure. This definition means that the somewhat arbitrary dimensionless distance $\epsilon$ defines wether or not a particle pair is connected. However, we will show that several clustering parameters follow classical percolation scaling laws for $\epsilon$, rendering results that are to a large degree independent of or insensitive to the choice of $\epsilon$. \figurename~\ref{fig:bc_sketch} shows a snapshot of a simulation containing 1024 particles at volume fraction $\phi = 0.4$ where we have also highlighted the cluster backbone by triangles where all segments are shorter than $r(2+\epsilon)$ to demonstrate the branched structure and the clusters. In the figure we use $\epsilon = 1$. The clusters in the system are highly transient, they stretch, compress, rotate and lead to transport of particles in the vertical direction. Particles continuously attach and detach to the clusters, so that the size and lifetime of a cluster is not trivial to define. Still, we will show that there is a unique definition of the cluster size distribution if we average in time.

\subsubsection{Percolation threshold}
If the particles were actually in contact, we would have a precise cluster definition. Since this is not the case for slowly sheared incompressible flows (without particle roughness), we need an alternative approach. We define $\epsilon_c$ as the lowest threshold at which a spanning cluster occurs. $\epsilon_c$ is then a continuos value that goes to zero if the system is truly percolating, i.e. particles are in direct contact. The maximum packing fraction is then characterized by
\begin{equation}
\lim_{\phi \rightarrow \phi_m} \epsilon_c(\phi) = 0
\label{eq:jamming_eps}
\end{equation}
Since $\epsilon_c \rightarrow 0$ will not occur in our system (except at maximum packing), minima in $\epsilon_c$ is the best approximation we can obtain. \figurename~\ref{fig:epsilon_c_scaling} shows the percolation threshold as a function of time for $N=1024$ and $\phi = 0.4$. We see that $\epsilon_c$ varies by a factor $\sim 2$, indicating the transient nature of the clusters.

Using scaling arguments we can predict the behavior of $\langle \epsilon_c \rangle$ with $\phi$. At the maximum packing fraction $\phi_m$, the system is spanning and $\epsilon_c$ goes to zero according to equation \ref{eq:jamming_eps}
\footnote{Note that there could be subtle differences between the isotropic maximum packing fraction and jamming when the system is anisotropic, and we do observe minor systematic differences in percolation threshold in the $x$ and $y$-direction.}.
This means that for $\phi$ approaching $\phi_m$, we expect that
\begin{equation}
\epsilon_c \sim (1-\frac{\phi}{\phi_m})^\alpha,
\end{equation}
where $\alpha < 0$ is an exponent that could in principle depend on $\phi$. For small fractions we expect a different scaling behavior. In very dilute systems, we can assume that the configuration does not depend on the particle radius, so that changing $\phi$ is equivalent to reducing $r$ for all particles. Then
\begin{equation}
r(2+\epsilon_c) \approx \text{constant}
\end{equation}
as $r$ is reduced. Using the relation $r \sim \phi^\frac{1}{2}$, we find that 
\begin{equation}
\epsilon_c \sim \phi^{-\frac{1}{2}} - 1
\end{equation}
One scaling relation that gives the expected behavior in both limits is
\begin{equation}
\epsilon_c \sim (1-\frac{\phi}{\phi_m})^\alpha \left ((\frac{\phi}{\phi_m} )^{-\frac{1}{2}}-1 \right )
\end{equation}
\figurename~\ref{fig:epsilon_c_scaling} shows $\langle \epsilon_c \rangle$ as a function of $\phi$. We obtain a good fit using $\alpha = 1$.
\subsubsection{Cluster size distribution}
\begin{figure}
\includegraphics[width=.5\textwidth]{Thogersen_fig3.eps}
\caption{Cluster number density scaled with traditional percolation parameters. N = 1024, $\phi = 0.3$. Scaled with universal percolation exponents $\tau = \frac{187}{91}$, the exponent of the power law distribution at the measured percolation threshold $\epsilon_c = 1.34$, and $\sigma = \frac{36}{91}$, which describes how the cutoff in cluster size diverges as the threshold is increased towards the percolation threshold. The inset shows the unscaled distributions. All data is equilibrated for $\gamma = 20$.
\label{fig:cluster_number_density_epsilon_collapse}}
\end{figure}

We define the cluster size $s$ as the number of particles that form a connected structure. We then measure the cluster number density $n(s,\epsilon)$ for different values of $\epsilon$. The cluster size distribution is highly dependent on the choice of the threshold parameter $\epsilon$.  This is expected if we look at the system as a percolation system where we have a percolation threshold $\epsilon_c(\phi)$ that depends on the volume fraction $\phi$. \figurename~\ref{fig:cluster_number_density_epsilon_collapse} shows the cluster number density for 1024 particles at $\phi = 0.3$ for various $\epsilon$, where the largest clusters are removed because they could in principle be spanning. As $\epsilon$ approaches the percolation threshold $\epsilon_c$, the cutoff in cluster size diverges and we get a power law distribution of cluster sizes, i.e. all clusters of all sizes exist. We obtain a decent collapse of the data using the universal percolation scaling exponents $\sigma = 36/91$, which describes the cutoff in the cluster size distribution as $\epsilon_c$ is approached, and  $\tau = 187/91$, which is the power law exponent of the cluster size distribution at the percolation threshold \cite{stauffer1994introduction}. However, even longer time-series are needed to obtain enough statistics to reach a firm conclusion on this matter.

Using the combination of the universal scaling exponents $\tau$ and $\sigma$ for the cluster size distribution as well as the scaling relation between $\langle \epsilon_c \rangle$ and $\phi$, we can uniquely determine the cluster size distribution given $\phi$ and $\epsilon$. The fluctuations are also well described through $\delta \epsilon_c \sim N^{-\frac{1}{3}}$. With this result established, we now turn to the particle collisions.

\subsection{Collision statistics}
\begin{figure}
\includegraphics[width=.5\textwidth]{Thogersen_fig4.eps}
\caption{Trajectory of two-particle collisions. The first particle position is used as a reference frame. In incompressible flow, particles will not actually touch. The blue circle shows the inaccessible region of radius $2r$, while the outer green circle shows the cutoff $\epsilon$ used to define collisions. Notice the existence of closed trajectories. The figure was generated from simulations of two particles in simple shear with $dt = \frac{0.02}{\dot \gamma}$ with a second order Runge-Kutta integrator, and similar resolution as in the rest of the simulations presented in this paper.
\label{fig:two_particle_interaction}}
\end{figure}
\begin{figure*}[p]
\includegraphics[width=.9\textwidth,trim=1.5cm 1.5cm 1.5cm 1.25cm, clip=true]{Thogersen_fig5.eps}
\caption{Example collisions with $\epsilon = 2$ from a simulation of $N=1024$ particles at volume fraction $\phi = 0.1$ in simple shear demonstrating two different collision regimes. \emph{Top left:} Examples of collisions with collision time $\Delta \gamma \leq 5$. \emph{Top right:} Collisions with collision time $\Delta \gamma \geq 50$. The bottom panels show the distance between the two particles as a function of $\gamma$. The line colors correspond to the line colors in the two top figures.
\label{fig:excursion_examples}}
\end{figure*}
\begin{figure*}
\begin{minipage}[t]{.49\textwidth}
\includegraphics[width=\textwidth]{Thogersen_fig6.eps}
\caption{Average excursion $\langle d_{ij} \rangle$ scaled by the minimum excursion $\langle d_{ij} \rangle_m$. Dashed lines show the standard deviation. A decrease of $\epsilon$ pushed the mean excursion from fairly parabolic towards semi-circular. All data is from simulations of $N = 1024$ particles.
\label{fig:mean_excursion}}
\end{minipage}
\hfill \hfill
\begin{minipage}[t]{.49\textwidth}
\includegraphics[width=\textwidth]{Thogersen_fig7.eps}
\caption{Probability density of collision times $C(\Delta \gamma)$. The curve is invariant under $\epsilon$ and $N$, but has a weak dependence on $\phi$. Long collision times scale as $\Delta \gamma ^{-3}$, but we expect a transition to $\Delta \gamma ^{-3/2}$ for even larger times (see text and \figurename~\ref{fig:return_time_random_walk}. Solid grey lines show results from two particles with $\epsilon = 1$ where the initial relative positions show radial symmetry (dashed) and uniformly distributed in $y$ at $-r(2+\epsilon)$ (solid).
\label{fig:contact_time_distribution}}
\end{minipage}
\end{figure*}
Since we are dealing with an incompressible fluid in the laminar regime and smooth particle surfaces, the particles are not touching, they are only interacting through the fluid phase. This means that we need to define what a collision is based on a length scale that is not related to any physical quantities. However, as we will discuss in this section, many results will not be sensitive to this choice.

\subsubsection{Collision trajectories}
In the previous section we already defined a length scale that we will continue to use. We define two particles to be colliding when they are within the distance $|\vec \chi_{ij}| \leq r(2+\epsilon)$. \figurename~\ref{fig:two_particle_interaction} shows the collision between two particles in simple shear, when they are the only particles present in the system. The black lines shows trajectories of the first particle in the reference frame of the second particle. The blue region is inaccessible since the particles have the same radius $r$. The green region is the region defining the collision between the two particles. Note that there are two different types of trajectories. The open trajectories, and the closed trajectories, first described by Batchelor and Green in 1972 \cite{batchelor1972hydrodynamic}, that will stay in contact indefinitely as long as no additional  particles are present that could disturb the trajectory.

The collisions in suspensions are more complex due to the interaction between more than two particles. Data from a simulation of $N = 1024$ particles at volume fraction $\phi = 0.1$ is shown in \figurename~\ref{fig:excursion_examples}.  From the figure one can clearly distinguish two different regimes, one for the short duration collisions, and one for the long duration collisions. Short duration collisions follow open trajectories similar to those demonstrated in \figurename~\ref{fig:two_particle_interaction} with some deviation due to simultaneous interaction with other particles. The long duration collisions follow trajectories more similar to the closed trajectories in \figurename~\ref{fig:two_particle_interaction}. We interpret the long duration collisions as particles that are pushed into closed orbits by random interaction with multiple particles, and then back to open trajectories before they escape. This is visible in the top right panel of \figurename~\ref{fig:excursion_examples} where one can see that the trajectories complete several loops before they escape.

To quantify the trajectory shape, we introduce the excursion profile which is a common measure in first passage random walks \cite{baldassarri2003average,colaiori2004average}. We reduce the problem to one dimension by only considering the radial distance $d_{ij}$ and velocity $v_{r,{ij}}$ between colliding particles. Collisions that are ongoing at the start or end of the simulations are not counted. A few examples of such excursions are shown in the two bottom panels of \figurename~\ref{fig:excursion_examples}. The short collisions exhibit close to parabolic shapes, while the longer collisions fluctuate in $d_{ij}$. To quantify the shape of the excursions, we scale the time axis of the individual excursions by the excursion time $\Delta \gamma$, and average. \figurename~\ref{fig:mean_excursion} shows the mean excursion shape for various $\phi$ and $\epsilon$ scaled by the maximum vale of the curves. As expected from the symmetry of Stokes equations, the average excursions are symmetric in time. Decreasing $\epsilon$ pushes the shape from parabolic towards semi-circular, which is related to the hydrodynamic forces of very close particles. Increasing $\phi$ has the same effect. However, the effects is not particularly large over the range we have tested ($\epsilon \in [0.1, 2]$ and $\phi \in [0.1, 0.4]$).

\subsubsection{Collision times}
Another quantity of interest is the distribution of collision times, i.e. how long do particle contacts last.  \figurename~\ref{fig:contact_time_distribution} shows the distribution of contact times for various $\epsilon$ and $\phi$. As previously, collisions that are ongoing at the start or end of the simulations are not counted. The distribution of collision times has the surprising feature that it is scale invariant, i.e. independent of $\epsilon$ for a wide range of $\epsilon$ tested. It is also only weakly dependent on the particle volume fraction $\phi$. The short collision times are dominated by open trajectories where the probability density scales as $\Delta \gamma$ with the most likely coillsion time of order $\Delta \gamma \approx 1$. The short collision times are well described by the hydrodynamic interaction between an isolated particle pair. Long collision times, however, exhibit a different scaling; we observe a scaling exponent $\approx \Delta \gamma ^{-3}$.

In an attempt to explain the scaling behavior of \figurename~\ref{fig:contact_time_distribution}, we turn to theory of first passages.  It follows from straightforward scaling arguments that the first passage distribution of a random walk is scale invariant under the condition that the standard deviation of the velocity over the mean value of the velocity is kept constant, i.e. $\delta v / \langle v \rangle = \text{constant}$. This is consistent with the standard deviation and mean velocity measured for collusions in our data. The results are shown in \figurename~\ref{fig:std_mean_coll}, where we see that equal $\phi$ have equal ratios $\delta v_r / \langle v_r \rangle$.

\begin{figure}
\includegraphics[width=.5\textwidth]{Thogersen_fig8.eps}
\caption{The ratio between the standard deviation $\delta v_r$ and mean radial velocity $\langle v_r \rangle$ for colliding particles. Only particles within distance $r(2+\epsilon)$ contribute to the curves. Equal $\phi$ (equal colors) collapse onto the same lines for different $\epsilon$, consistent with scale invariance of the return time of random walks.
\label{fig:std_mean_coll}}
\end{figure}

The first passage time distribution of a random walk is a power law distribution with exponent $\beta = -\frac{3}{2}$ \cite{redner2001guide}, which does not match the long time limit observed in our data. One possible explanation for this dicrepancy is long time correlations of the collision velocities. We would then expect a transition from $\beta = -3$ to $\beta = -\frac{3}{2}$ in the very long time limit. The reason we would not see this is because our time-series are not long enough.  Even though collisions that long are very unlikely, the scaling behavior for long collision times is important to determine wether the mean collision time diverges or not because the first moment of a power law distribution diverges for $\beta \textgreater -2$.

\begin{figure}[h!]
\includegraphics[width=.5\textwidth]{Thogersen_fig9.eps}
\caption{\emph{Main figure:} First passage probability density for the random walk model with data from $N = 1024$, $\phi = 0.1$ and $\epsilon = 2$.  The features from the finite element data are recovered, and a $\beta = -\frac{3}{2}$ scaling is found for large $\Delta \gamma$. The gray line shows the boundaries of \figurename~\ref{fig:contact_time_distribution}. \emph{Top inset:} Mean and standard deviation of $v_r$ from finite element data. The dashed lines show the parameterized curves that are used as inputs for the random walk. \emph{Bottom inset:} Frequency spectrum of $v_r$ from finite element data where we have subtracted $\langle v_r \rangle$. Only data from collisions with $\Delta \gamma > 50$ is used. The dashed line shows the parameterized curve used in the random walk model. Note the transition from pink ($1/f^a$) noise at high frequencies to white noise (constant) at low frequencies.
\label{fig:return_time_random_walk}}
\end{figure}

\subsubsection{First passage random walk model}
We attempt to model the collision time distribution using correlated random walks. The first thing we need to do is to parameterize the parameters that we need to input to the random walk. They are the mean radial velocity during a collision and the standard deviation, as well as the self correlation of each walk. For the self correlation we use the power spectrum because we will use the same power spectrum to generate random walks. We perform a discrete fourier transform of $v_r$ for all collisions that last longer than $\Delta \gamma = 50$. This is because the short collisions are not long enough to get reliable frequency spectra. The insets in \figurename~\ref{fig:return_time_random_walk} shows $\delta v_r$, $\langle v_r \rangle$ and the frequency spectrum $P_{v_r} (1/\gamma)$ for a single run with $N=1024$, $\phi = 0.1$ and $\epsilon = 2$, as well as fits used as inputs for the random walk model. $\delta v_r$ and $\langle v_r \rangle$ show the correlations on the time scale of order $\delta \gamma \sim 20$. The frequency spectrum shows power law noise at high frequencies and a transition to white noise for low frequencies.

We construct our random walk in fourier space that satisfies the measured power spectrum. Each collision velocity $v_r$ follows a power spectrum made from uniform random weights scaled with $P_{v_r} (1/\gamma)$. $v_r(\gamma)$ is then obtained from an inverse discrete fourier transform. We then multiply with the measured standard deviation and add the mean value that we observed. The resulting collision time distribution $C$ is plotted in \figurename~\ref{fig:return_time_random_walk}. The correspondence with our data from finite element simulations is very good, and the simplified model also predicts a crossover from $\beta = -3$ to $\beta = -\frac{3}{2}$ at very large times. This transition from pink noise at high frequencies to white noise at low frequencies is responsible the power law behavior with $\beta = -\frac{3}{2}$ in $C(\Delta \gamma)$ for $\Delta \gamma \gtrsim 10^2$. The good correspondence between the collision time probability density and the return time of the correlated random walk suggests that our interpretation of the collisions as particles that are pushed into what would in the two particle case be closed orbits, and are then subject to random interactions with additional particles in the suspensions is justified. Details of the effect of $P_{v_r}$, $\delta v_r$ and $\langle v_r \rangle$ are discussed in appendix \ref{appendix:random_walk}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion and conclusion} \label{section:discussion}
In this paper we have introduced a two-dimensional finite element model to study particle suspensions in simple shear at zero Reynolds number. We use geometrical mesh refinement in the regions between close particles to assure that lubrication forces are modeled accurately. This allows us to avoid the use of repulsive forces that are commonly implemented to avoid particle overlaps. This does set constraints on the $\gamma$ we can reach for high particle volume fractions, but we have reached $\gamma$ large enough to perform statistical analysis when combined with ensemble averaging.

The particles form transient clusters that that rotate and transport particles in the vertical direction. To study the cluster statistics we introduced a dimensionless contact length $\epsilon$ and investigated how particle collision statistics depend on this length.  The cluster number density follows classical percolation theory, and scales with the universal scaling exponents $\tau$ and $\sigma$. While the percolation threshold fluctuates in time with the fluctuations scaling with system size as $N^{-\frac{1}{3}}$, the time-averaged cluster number density is well defined. Even though de Gennes proposed that spanning clusters are directly related to an increase in effective viscosity we have not found direct correlation between $\epsilon_c(\gamma)$ and $\mu_\text{eff}(\gamma)$. This is supported by a recent study by Gallier et al. \cite{gallier2015percolation}, where they find that although spanning clusters can support more stress, the correlation with fluctuations in effective viscosity and the occurence of transient spanning clusters is weak. We intend to elaborate on the relation between structure and effective viscosity in a separate publication.

Two particles in simple shear exhibit to different collision trajectories; open trajectories and closed trajectories. For particle collisions where many particles are present and can change the trajectories of a particle pair, this distinction is still visible. Short collisions are similar to the open collision trajectories involving only two particles, while we interpret long collisions as particles pushed into what would in the two particle case be closed trajectories. The trajectories are then subject to random interactions with other particles, until they are pushed out of the closed trajectory. This interpretation is supported by the close resemblance between the collision time probability density and the return time of a correlated random walk. Using $\langle v_r \rangle$, $\delta v_r$ as well as the power spectrum of each collision trajectory, we reproduced the collision time probability density. We also predict that the long time behavior scales as a power law with exponent $\beta = -\frac{3}{2}$, but we have not reached strains large enough to confirm this prediction. $\beta = -\frac{3}{2}$ in the long time limit indicates that the mean collision time diverges since the first moment of a power law distribution with exponent $\beta > -2$ diverges. Due to the conservation of the ratio $\delta v_r / \langle v_r \rangle$ with changing $\epsilon$, the distribution of collision time is scale invariant. This is again due to the self affine properties of $v_r$, that are visible in the power spectrum $P_{v_r}$. While lubrication forces are slightly different in three dimensions, we believe that our results are qualitatively valid also in three dimensions.

Our results on the collision time probability density highlight some of the challenges we face with regards to long time series using finite element simulations of particle suspensions. The interpretation of contacts as first passage processes implies that it is quite likely to find very close particle pairs in sheared suspensions. The probability of finding such pairs would scale with the total number of contacts, i.e. $N$ and $\phi$, as well as total strain $\gamma$. Since close particle pairs are difficult to treat numerically due to the large variation in element size and the singularity of the zero spacing between particles, this puts severe constraints on the likelihood of running large systems to high strains. To overcome this difficulty, approaches such as upscaling of the lubrication layer between close particles or short range repulsive forces are necessary.

Several studies have been performed using repulsive forces motivated by friction, electrostatic repulsion or van der Waals forces (see e.g. \cite{brady1997microstructure,sierou2002rheology,morris2009review,seto2013discontinuous}). Even short range repulsive forces will break the symmetry of the two-particle interaction in \figurename~\ref{fig:two_particle_interaction}. We do not expect short range repulsive forces to change the short collision trajectories significantly, but the long collisions will be altered, and if the repulsive force is large enough, closed trajectories will cease to exist.

An interesting question is wether simple random walk models can be used to predict the transient properties of the cluster size distribution. Given the probability density of particle collision times, one could imagine combining it with a collision rate could lead to predictions of the cluster size distribution using a purely statistical approach. However, we leave this open for possible future investigations. Understanding the microstructure of particle suspensions, and in particular, the transient nature of close particle contacts is an important step towards understanding the relation between suspension microstructure and effective properties, including fluctuations. In particular, close to the maximum packing fraction, the fluctuations in the percolation threshold show that transient percolation clusters appear. Gallier et al. found only a weak correlation between spanning clusters and instantaneous effective viscosity \cite{gallier2015percolation}, but it could be that systems even closer with volume fractions very close to jamming are needed to observe significant effect. It could also be the case that there are other properties of the micro-structure that govern fluctuations in the effective viscosity.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Ackowledgements}
K.T. acknowledges support from VISTA, a basic research program funded by Statoil, conducted in close collaboration with The Norwegian Academy of Science and Letters.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix

\section{Details of the random walk model} \label{appendix:random_walk}
\begin{figure}
\includegraphics[width=.5\textwidth]{Thogersen_fig10.eps}
\caption{
Measurements of the return time distribution for random walks following equations \ref{eq:P_vr}, \ref{eq:mean_vr} and \ref{eq:delta_vr}.
\emph{Top left:} Effect of $a$ on the return time distribution $C(\Delta \gamma)$. $\langle v_r \rangle = 0$, $\delta v_r = \text{constant}$.
\emph{Top right:} Effect of $b$ on the return time distribution $C(\Delta \gamma)$. $\langle v_r \rangle = 0$, $\delta v_r = \text{constant}$.
\emph{Bottom left:} Effect of $c$ on the return time distribution $C(\Delta \gamma)$. $a = 1$, $b = 12.5$, $\frac{\delta v_r}{\langle v_r \rangle}= \frac{2}{5}$.
\emph{Bottom right:} Effect of $d$ on the return time distribution $C(\Delta \gamma)$. $a = 1$, $b = 12.5$, $\frac{\delta v_r}{\langle v_r \rangle} = \frac{2}{5}$.
\label{fig:random_walk_param_study}}
\end{figure}
In this section we give the details of the random walk model used in the discussion of the collision times. The parameterized curves we use at inputs are 
\begin{equation}
P_{v_r}(1/ \gamma) \sim \frac{1}{(1+\frac{b}{\gamma})^{a}},
\label{eq:P_vr}
\end{equation}
\begin{equation}
\frac{\langle v_r \rangle}{\langle v_r(0) \rangle} = \frac{5}{3}e^{-\gamma/c} -\frac{2}{3}e^{-\gamma/d},
\label{eq:mean_vr}
\end{equation}
and
\begin{equation}
\frac{\delta v_r}{\langle v_r(0) \rangle} = \frac{2}{5}\gamma^2 e^{-\gamma} + \frac{1}{4}e^{-\gamma/40} + \frac{1}{20},
\label{eq:delta_vr}
\end{equation}
where we have used $a = 1.1$, $b = 12.5$, $c = 1$ and $d = 8$ in the model in the main text. The random walk is constructed using uniform random weights scaled with $P_{v_r}$. In addition, we use a cutoff frequencies $1/\gamma > 5$. We then perform an inverse discrete fourier transform, scale it to get the correct standard deviation$\frac{\delta v_r}{\langle v_r(0) \rangle}$ and add $\frac{\langle v_r \rangle}{\langle v_r(0) \rangle}$ to get $v_r$.

\figurename~\ref{fig:random_walk_param_study} shows how $a$, $b$, $c$ and $d$ control the shape of $C(\Delta \gamma)$, where we have performed a series of random walks and measured $C(\Delta \gamma)$. $a$ and $b$ control the behavior at long times, while $c$ and $d$ control the short time behavior. $a$ controls the fractal dimension of the walk at high frequencies, while $b$ controls the crossover from a self affine walk to white noise. $c$ shifts the maximum point of $C(\Delta \gamma)$, while $d$ controls the shape of the peak. It is the combination of these parameters that make up \figurename~\ref{fig:return_time_random_walk}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{Thogersen_main_text.bbl}

\end{document}

### 2017-01-27 ms.tex
Lots of additions

@@
 \begin{document}
-\title{Generating realistic nanopores using procedural methods}
+\title{Structural replication of nanoporous media using procedural noise}
 %\author{A. Hafreager, N. Groeneboom, A. Malthe-S\oe renssen$^1$}
@@
 \begin{abstract} 
+In this paper, we aim to develop a method for producing statistically correct simulations of nanoporous media without the use of conventional Molecular Dynamics Simulator (MDS).
+In short, the proposed model is a binary function that returns either true if the input position is in a void / pore, or false if within a wall.
+By using procedural noise, we are not restricted to generating media within a limited box, as the noise functions is continuous on all scales.
+It is therefore possible to validated pores at theoretically arbitrary range.
+In addition, this method is extremely fast, requiring very little computational power to calculate whether a particle is within a pore or not.
+After presenting such a model, we choose a method to measure the structural properties of nanoporous media that are created either through regular methods (e.g.
+LAMMPS) or with our method.
+The measure we selected is the radial pore size distribution $g(r)$, which is regarded as the de-facto standard method for capturing the structural properties of porous media.
+Using this measure, we define a full likelihood framework for estimating model parameters given any data set, and continue by validating our own framework by analysing procedural noise data sets with known input parameters.
+Finally, we perform a full likelihood analysis of silica $SiO^2$ data set simulated in LAMMPS, and show that our model correctly reproduces several of the internal properties of the data set, including $g(r)$, porosity and surface area.
+
 \end{abstract}
@@
 \section{Introduction}
+
 Lorem ipsum dolor sit amet, consectetur adipiscing elit.
@@
 \section{Method}
+\todo{her står det nå det samme som i abstract}
+In this paper, we aim to develop a method for producing statistically correct simulations of nanoporous media without the use of conventional Molecular Dynamics Simulator (MDS).
+In short, the proposed model is a binary function that returns either true if the input position is in a void / pore, or false if within a wall.
+By using procedural noise, we are not restricted to generating media within a limited box, as the noise functions is continuous on all scales.
+It is therefore possible to validated pores at theoretically infinite range.
+In addition, this method is extremely fast, requiring very little computational power to calculate whether a particle is within a pore or not.
+
+After presenting such a model, we choose a method to measure the structural properties of nanoporous media that are created either through regular methods (e.g.
+LAMMPS \cite{plimpton1995fast}) or with our method.
+The measure we selected is the radial pore size distribution $g(r)$, which is regarded as the de-facto standard method for capturing the structural properties of porous media.
+
+Using this measure, we define a full likelihood framework for estimating model parameters given any data set, and continue by validating our own framework by analysing procedural noise data sets with known input parameters.
+Finally, we perform a full likelihood analysis of
 
@@
 \subsubsection{Properties of procedural noise}
-Standard Perlin noise is known for having several anisotropic features
-\citep{lagae:2010}. Perlin noise also gives rise to a ``textile''
-pattern in the $x$ and $y$-directions, which is evident from the upper
-left part of figure \ref{fig:noise}, depicting 2D Perlin noise in
-a single octave. The upper-right part shows the same node for Simplex noise,
-with less obvious patterns. However, when combining several octaves,
-as depicted in the two lower parts, these anisotropic patterns tend to cancel
-out. In any case, for observations of weak gravitational lensing, these anisotropies would occur on much smaller 
-scales than the typical resolution. 
-In addition, we have calculated the statistical distributions for Perlin and simplex noise
-with $\sigma = 0.2$, and have plotted the results in figure \ref{fig:properties}.
-Each procedural noise method has distributions that are close to normal, 
-but with some intrinsically shifted sigma and mean.
-
 \begin{figure}
 \includegraphics[width=.5\textwidth]{noise.eps}
-\caption{Perlin noise (left) vs simplex noise (right). Note the
+\caption{Perlin noise (left) versus simplex noise (right). Note the
   straight x-y pattern in the single upper Perlin mode and the less
@@
 \end{figure}
+Standard Perlin noise is known for having several anisotropic features
+\citep{lagae:2010}. Perlin noise also gives rise to a ``textile''
+pattern in the $x$ and $y$-directions, which is evident from the upper
+left part of figure \ref{fig:noise}, depicting 2D Perlin noise in
+a single octave. The upper-right part shows the same node for Simplex noise,
+with less obvious patterns. However, when combining several octaves,
+as depicted in the two lower parts, these anisotropic patterns tend to cancel
+out. In any case, for observations of weak gravitational lensing, these anisotropies would occur on much smaller 
+scales than the typical resolution. 
+In addition, we have calculated the statistical distributions for Perlin and simplex noise
+with $\sigma = 0.3$, and have plotted the results in figure \ref{fig:properties}.
+Each procedural noise method has distributions that are close to normal, 
+but with some asymmetrical properties.
 
@@
 \label{eq:perlinlinear}
- \Phi(\vec x) = \sum_k \frac{1}{k} P( k\vec x), 
+ \Phi(\vec x) = \sum_k \frac{1}{k} P( 2k\vec x), 
 \end{equation}
@@
 part of figure \ref{fig:noise}.
-
-%\ref{burganos1998gas}
-
-%\begin{figure}
-%\includegraphics[width=.45\textwidth]{pores2.png}
-%#\includegraphics[width=.5\textwidth]{pores2.eps}
-%\includepdf{ pores1.pdf }
-%\caption{Model with best-fit parameters}
-%\label{fig:model12}
-%\end{figure}
-
-%\subsection{Simulations}
-%LAMMPS $SiO_2$ simulation heated -> expand box -> cool down
-%\begin{figure}
-%\includegraphics[width=.45\textwidth]{pores1.png}
-%%#\includegraphics[width=.5\textwidth]{pores2.eps}
-%%\includepdf{ pores1.pdf }
-%\caption{LAMMPS $SiO_2$ simulation.}
-%\label{fig:model12}
-%\end{figure}
+Finally, since simplex noise both scales better with higher dimensions and produces less asymmetric effects, we choose to use this noise method over regular Perlin noise in our method.
 
 \subsection{Model}
-In this paper, we are interested in developing a model using Simplex noise to reproduce nanoporous structures that are as statistically close to actual simulations of porous media.
+In this paper, we are interested in developing a model using Simplex/Perlin noise to reproduce nanoporous structures that are as statistically close to actual simulations of porous media.
 For a given data set and a model, we wish to estimate the parameters that best represent the data.
-This is performed through a full likelihood analsys, where we in addition validate the model by comparing the radial pore size distribution $g(r)$, porosity and surface area of the system.
+This is performed through a full likelihood analysis, where we in addition validate the model by comparing the radial pore size distribution $g(r)$, porosity and surface area of the system.
 
@@
 \end{equation}
-where $\Omega$ is the total amount of octaves, $\psi$ the persistence (scale falloff), $N$ a generic $N$-dimensional noise function, $\sigma$ a predefined random vector that removes anisotropic artifacts and $\nu$ the frequency modifier representing the initial largest scales in the model.
+where $\Omega$ is the total amount of octaves, $\psi$ the persistence (scale falloff), $N$ a generic $N$-dimensional noise function, $\sigma$ a predefined random vector that removes anisotropic artefacts and $\nu$ the frequency modifier representing the initial largest scales in the model.
 For each mode, the frequency is thus multiplied, while the overall amplitude is damped by a given persistence.
 In the end, we simulate a large set of bulk particles without any walls, and use the noise function to remove particles that are not in a procedural void given by a threshold $\tau$.
-Three examples of this model are depicted in figure \ref{fig:model_example}, where we have used $\Omega=6$ ocataves and the scale is $\nu=0.01$ and cutoff $\tau=0$.
+Three examples of this model are depicted in figure \ref{fig:model_example}, where we have used $\Omega=6$ octaves and the scale is $\nu=0.01$ and cut-off $\tau=0$.
 The left image depicts low persistence ($\psi = 0.0$), where only the first octave dominates (corresponding to large scales).
 The middle picture shows the same parameters with a higher persistence $(\psi = 0.5)$, where smaller scales become more dominating.
-In the rightmost picture, the effect of the cutoff threshold $\tau = -0.3$ is shown.
+In the rightmost picture, the effect of the cut-off threshold $\tau = -0.3$ is shown.
 
@@
 \subsection{Likelihood}
-The likelihood $\mathcal{L}$ of paremeters $\theta$ of a model $M(\theta)$ given outcome $x$ is defined as $\mathcal{L} (M(\theta) | x) = P( x | M(\theta))$, where P is the probability distribution.
+The likelihood $\mathcal{L}$ of parameters $\theta$ of a model $M(\theta)$ given outcome $x$ is defined as $\mathcal{L} (M(\theta) | x) = P( x | M(\theta))$, where P is the probability distribution.
 In our model, we use the $\chi^2$ test to fit our one-dimensional measure from the data $\bf d$ with our model $\bf M(\theta)$.
@@
 In this example, we only use standard simplex noise with three free parameters: the persistence $\psi$, threshold $\tau$ and scale $\nu$.
-We simulate a data set of $159 ^3 Å$ of $1 000 000$ atoms, and perform the cutoff given input parameters $\Omega=3$, $\Psi = 0.6$, $\tau=-0.2$, $\nu=0.03$.
+We simulate a data set of $159 ^3 Å$ of $1 000 000$ atoms, and perform the cut-off given input parameters $\Omega=3$, $\Psi = 0.6$, $\tau=-0.2$, $\nu=0.03$.
 The resulting 3D structure is depicted in figure \ref{fig:mockdata}.
 
-We continue by performing a full MCMC analysis of the mock data with using a parallelised code running on a 48 core cluster.
+We continue by performing a full MCMC analysis of the mock data with using a parallelized code running on a 48 core cluster.
 In order to produce 300 000 samples, about 48 hours was needed.
 The results from this analysis can be seen in figure \ref{fig:mockdataresults_pts}, where the upper panel depicts the marginalised 1D posteriors while the lower shows the marginalised 2D posteriors.
-The center and left image shows that the input parameters of $\tau=-0.2$(threshold), $\psi=0.6$(persistence) and $\nu=0.03$ are all reproduced well within $1 \sigma$.
+The centre and left image shows that the input parameters of $\tau=-0.2$(threshold), $\psi=0.6$(persistence) and $\nu=0.03$ are all reproduced well within $1 \sigma$.
 
@@
 
-We now wish to perform a parameter fitting analysis on realistic simulated nanoporous media.
+We now perform a parameter fitting analysis on realistic simulated nanoporous media.
 Using LAMMPS \cite{plimpton1995fast}, we simulate a $100^3$ \AA box of 100K $SiO^2$ silica particles at an initial temperature of $4500 K$, using the Vashishta-potential.
@@
 The porosity of the input $SiO^2$ data set is $75\%$, while the best-fit data set has porosity $77\% \pm 3 \%$, depending on the random seed.
-In addition, the surface area of the original data set was $\tilde 29 000 \AA^2$, while the best-fit data shows similar values with surface are ranging betweem $\tilde 28 000 \AA^2$ and $\tilde 30 000 \AA^2$.
+In addition, the surface area of the original data set was $\tilde 29 000 \AA^2$, while the best-fit data shows similar values with surface are ranging between $\tilde 28 000 \AA^2$ and $\tilde 30 000 \AA^2$.
 
@@
 However, the surface area is strongly correlated with this value, so even slight deviations (e.g.
-$\delta \tau = 0.05$) yields a relatively large change in surface area.
+$\Delta \tau = 0.05$) yields a relatively large change in surface area.
 
@@
 We have developed tools for creating such porous materials, and also estimating model parameters given data.
-We have shown that it is possible to recreate
+We have shown that it is possible through a full likelihood analysis to reproduce the input parameters of a simulated model.
+In addition, we performed an analysis on a simulated $159^3 \AA$ data set of $SiO^2$, and explain that while porosity and surface area corresponds well between the model and data, the radial pore size distribution $g(r)$ show that the current model does not exactly reproduce correct data.
+However, we do not claim that the current model should do so yet, since we have chosen the simplest kind of summation with regards to procedural noise modes.
 
 \subsection{Future work}
+In this paper, we have only considered the simplest kind of procedural noise: a simple sum over modes with threshold cut-off $\tau$, scale fall-off persistence $\psi$ and initial scale $\nu$.
+We have also seen that, even through both porosity and surface area accurately reproduced, there are features in the radial pore size distribution $g(r)$ that are not fully captured with the current model.
+This is evident from the images of the simulated structure seen in \ref{fig:porous_vs_model}, where the connecting nearly one-atom thin walls are difficult to reproduce with the current model.
+However, as procedural noise exert properties closely related to that of fractals, it is possible to extend the number of parameters and define new models with very different properties.
+Some examples of these models are depicted in figure \ref{fig:future_models}, where the left- and centre images show a new model with two new parameters (offset and gain), while the rightmost panel shows a simulation where the scale $\nu$ is itself being perturbed by procedural noise, yielding strong circular asymmetric patterns.
+
+Another immediate application would be to estimate model parameters given actual observations of porous media.
+This would enable us to estimate new parameters such as instrumental noise, deviations from symmetry (yielding asymmetric properties) and  pureness of samples.
+In addition, by analysing just a few samples and producing an accurate statistical model, we are able to produce an arbitrary amount of statistically correct samples, free from instrumental noise and artefacts.
+
+A final key analysis would be to compare properties such as percolation through finite-element simulations of gas/liquid flow through our simulated structures, in addition to exerting shear forces, and comparing with actual simulated data from e.g.
+LAMMPS.
 
@@
 \caption{Example of future models.
-Left panel: A multifractal model with thin surfaces and large pores.
-Center panel: a multifractal with cave-like structures.
+Left panel: A multi-fractal model with thin surfaces and large pores.
+Centre panel: a multi fractal with cave-like structures.
 Right panel: A model with strong asymmetric noise.

### 2017-02-04 ms.tex
Hyttetur work paper

@@
 LAMMPS) or with our method.
-The measure we selected is the radial pore size distribution $g(r)$, which is regarded as the de-facto standard method for capturing the structural properties of porous media.
+The measure we selected is the radial distribution function (RDF) $g(r)$, which is regarded as the de-facto standard method for capturing the structural properties of porous media.
 Using this measure, we define a full likelihood framework for estimating model parameters given any data set, and continue by validating our own framework by analysing procedural noise data sets with known input parameters.
-Finally, we perform a full likelihood analysis of silica $SiO^2$ data set simulated in LAMMPS, and show that our model correctly reproduces several of the internal properties of the data set, including $g(r)$, porosity and surface area.
+Finally, we perform a full likelihood analysis of silica SiO2 data set simulated in LAMMPS, and show that our model correctly reproduces several of the internal properties of the data set, including $g(r)$, porosity and surface area.
+\todo{Abstract boer tenkes noe gjennom: faa frem hovedpoenger.}
 
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Duis consequat turpis vel mollis accumsan.
-Etiam eleifend magna vel quam porta, sed porta nibh placerat.
-Sed in faucibus risus.
-Duis suscipit tincidunt leo, nec laoreet leo elementum elementum.
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
-Sed eros risus, convallis sit amet turpis a, dapibus accumsan quam.
-Donec fringilla rhoncus turpis, quis ultrices odio vulputate elementum.
-Cras nec viverra urna.
-Phasellus laoreet tortor sed est sagittis laoreet.
-Nullam a volutpat ex.
+%During recent years, increase in computational power has allowed for a vast improvement  
+\todo{Skriv en fin intro! Ogsaa om nanoporous media}
+Properties of structural patterns that regularly occur in nature are often investigated either through direct observations or through simulations of theoretical models.
+Model simulations typically require a large computational framework to accurately reproduce the observed effects, while experimental data can be both expensive and hard to obtain, in addition to being riddled with artefacts and instrumental errors.
 
-Mauris vestibulum sit amet libero eget facilisis.
-Vivamus a nunc sed felis ultrices convallis luctus et mauris.
-Sed tincidunt mauris nunc, vehicula placerat eros faucibus quis.
-Morbi egestas, nisi quis luctus dignissim, mauris odio pulvinar tellus, non efficitur risus erat vel nunc.
-Cras vel felis magna.
-Mauris rutrum magna a interdum sodales.
-Nam tempor, diam non sodales tristique, nulla ipsum eleifend dui, non aliquam felis tellus id enim.
-Donec eget nulla leo.
+Procedural generation of physical structures have several advantages over conventional simulations.
+For example, when producing simulations of nanoporous silica, we are required to calculate and propagate the physical properties of each particle for every time-step throughout the simulation, using a computational framework such as LAMMPS \cite{plimpton1995fast}.
+This is both time-consuming, and the computational cost can sometimes be be quite large.
+In addition, the simulation is constrained to a box, so particles are not allowed to move outside the defined boundaries.
+On the other hand, using procedural noise to carve out nanoporous media is both extremely fast and computationally cost-efficient, in addition to being continuous on all scales.
+This means that it is possible to simulate
 
+%In addition, by creating a model that yields properties that are statistically consistent with observations, we are ar
+
+In this paper, we aim to investigate whether procedural methods are suitable for producing statistically correct 3D physical structures, focusing on nanoporous media.
+In a sense, we are combining well-known methods from computer games and real-time graphics-oriented software with experimental observations and simulations.
+Procedural methods for automated generation of computer game content have been around since the early 1980's, and in those early days algorithms were less advanced, often yielding plain, simple geometrical structures.
+Nowadays, almost every single computer game produced will contain some level of automated generated content, from textures to level design to landscapes and gameplay.
+
+We do not claim that procedural noise methods completely reproduces all structural properties of nanoporous silica, but show that our approach can be used to efficiently create materials that are structurally similar to simulations.
+Naturally, this all depends on the statistics used for comparing the two models, which in this case is based on the radial distribution function (RDF) $g(r)$.
+
+%\section{Method}
+\section{Procedural content generation}
+By procedural content generation, we generally mean a mathematical algorithm that uses limited input data to create complex, sometimes infinite repeating patterns.
+A firm example would be fractals, such as the Julaia/Mandelbrot [ref mandelbrot] set, where one tests whether the complex quadratic recurrence equation $z_{n+1} = z_n^2 +C$ for $z \in \mathbb C$ converges for each position in a 2D complex plane.
+With this seemingly simple equation, the level of complexity generated on the boundary between convergence and divergence is infinite, with self-similar patterns repeating themselves to arbitrary detail.
+\begin{figure}
+\includegraphics[width=.5\textwidth]{orphancity.png}
+\caption{Example of city building using procedural methods: Landscape, city infrastructure, buildings and all textures are all automatically generated using procedural methods.
+Created by one of the authors.
+\todo{Denne er jo med bare for moro skyld, boer jo fjernes} }
+\label{fig:orphancity}
+\end{figure}
+
+Another example, widely used in contemporary game content generation, is having prefabs (pre-fabricated 3D models such as doors, walls, tiles and roof segments) automatically being stitched together to create buildings.
+When designing a city, instead of creating each building manually, you only need specify the position, area and height of a structure.
+Then, by using an algorithm that combines these prefabs to render the building, it is possible to create cities of arbitrary size with a single command.
+Naturally, it is also possible to create procedural methods that design the layout and placement of these buildings.
+See figure \ref{fig:orphancity} for an example of such a city generator tool, developed by one of the authors of this paper.
+
+A much simpler example would be generating a 1 (or 2)-dimensional landscape by choosing random points on a grid and performing a spline interpolation between them to create the illusion of a smooth landscape.
+In fact, this is the very essence of procedural noise, which we are focusing on in this paper, and the first person to create such a fast algorithm for N-dimensional spline interpolation between pseudo-random points was Ken Perlin.
+
+\subsection{Perlin/Simplex noise}
+\label{sec:perlin}
+
+Procedural algorithmic methods for creating realistic-looking
+fractal-like patterns have been around for more than 30 years. The most famous of these
+algorithms is Perlin noise, first developed by Ken Perlin in 1982 for
+use in the movie ``Tron''.
+
+Perlin noise is a type of pseudo-random gradient noise that mimics a Gaussian field, but at low computational cost.
+An improved version of Perlin noise is called simplex noise, which has several advantages over standard Perlin noise in terms of scalability and speed.
+Perlin noise was developed in response to the lack of detail 1980's contemporary graphics, where polygonal surfaces often were bland and left without any detail.
+In response to this problem, Perlin noise was developed in order to continuously perturb the light bouncing off flat surfaces, effectively yielding bump mapping and producing a much more realistic image.
+
+An overview of the current state of procedural noise functions can be
+found in \cite{lagae:2010}.  Perlin noise \citep{perlin:2002} 
+is a band limited repeatable pseudo-random function
+from $\mathbb R^n \to \mathbb R$ that is computationally fast, and has
+properties which make it suitable for creating patterns that mimic
+nature. Perlin noise is an approximation to Gaussian filtered noise
+implemented as a pseudo-random spline. The algorithm uses a predefined
+grid of gradients pointing in a random direction, where for any point
+within a grid, we interpolate the gradients from the four corners. The
+algorithm is as follows:
+\begin{itemize}
+  \item[1.] For a point, obtain the four closest gradients in the grid.
+   \item [2.] For each of the gradients, calculate the dot product
+     between the gradient and a vector defined by the  the distance
+     between the the point and the current grid corner. 
+   \item [3.] Instead of linear interpolation $F(t) = t$, use a fade function to produce smoother interpolated values.
+Perlin noise typically uses the following fade function: $F(t) = t ^3 (t  (6 - 15) + 10)$.
+    \item [4.] Obtain the final value by linearly interpolating between the x-axis and
+      use this result to linearly interpolate the y-axis. 
+\end{itemize} 
+Simplex noise works in a similar manner, but requires fewer
+computational steps, scales better in higher ($\ge 3$) dimensions and has fewer 
+anisotropic properties.
+
+\subsection{Properties of Perlin/Simplex noise}
+\begin{figure}
+\includegraphics[width=.5\textwidth]{noise.eps}
+\caption{Perlin noise (left) versus simplex noise (right). Note the
+  straight x-y pattern in the single upper Perlin mode and the less
+  anisotropic simplex mode. Lower part: combining octaves tend
+  to remove these patterns.}
+\label{fig:noise}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{noise_comparison.png}
+%\mbox{\epsfig{file=perlinstats.eps, width=\linewidth}}
+\caption{Normalized distributions of Perlin noise and simplex noise compared with a
+  Gaussian distribution. Note how both noise functions produce
+  near-Gaussian, but slightly different distributions.}
+\label{fig:properties}
+\end{figure}
+Standard Perlin noise is known for having several anisotropic features
+\citep{lagae:2010}. Perlin noise also gives rise to a ``textile''
+pattern in the $x$ and $y$-directions, which is evident from the upper
+left part of figure \ref{fig:noise}, depicting 2D Perlin noise in
+a single octave. The upper-right part shows the same node for simplex noise,
+with less obvious patterns. However, when combining several octaves and adding a random shift to the coordinates,
+as depicted in the two lower parts, these anisotropic patterns tend to cancel
+out.  In addition, we have calculated the statistical distributions for Perlin and simplex noise
+with $\sigma = 0.3$, and have plotted the results in figure \ref{fig:properties}.
+Each procedural noise method has distributions that are close to normal, 
+but with some asymmetrical properties.
+
+\subsection{Multiple octaves: creating patterns}
+\label{sec:octaves}
+The different octaves from procedural noise functions can be combined
+to produce a wide range of nature-like patterns. A generic
+N-dimensional procedural function $\Phi(\vec x): \mathbb R^N \to
+\mathbb R$ could be expressed as
+\begin{equation}
+\label{eq:procedural}
+ \Phi(\vec x) = \Theta \Big(\sum_k^\omega \kappa (k) P(  f_s (\vec x,k) )) \Big),
+ \end{equation}
+where $f_s(\vec x,k)$ describes the frequency scale modifier (e.g. $f_s(\vec x,k) =
+k\vec x$), $\omega$ is the amount of octaves (modes), $\kappa(k)$ is the octave amplitude (e.g. $\kappa(k) = 1/k$)
+and $\Theta(x)$ an overall modifier (e.g. $\Theta(x) = x$ or
+$\Theta(x) = \frac{1}{x}$). In a sense, $\kappa(k)$ is similar to Fourier coefficients. 
+In its simplest case, summing directly over the octaves with amplitude $1/k$ results in 
+\begin{equation}
+\label{eq:perlinlinear}
+ \Phi(\vec x) = \sum_k^\omega \frac{1}{k} P( 2k\vec x).
+\end{equation}
+In 1 dimension, equation \ref{eq:perlinlinear} with 1, 3 and 30 octaves is depicted in figure \ref{fig:1dperlin}, while the 2D procedural noise pattern depicted in the lower
+part of figure \ref{fig:noise}.
+Finally, since simplex noise both scales better with higher dimensions and produces less asymmetric effects, we choose to use this noise method over regular Perlin noise in our method.
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{1d_perlin.png}
+\caption{1D simplex noise with 1 (blue), 3 (green) and 30 (red) octaves. }
+\label{fig:1dperlin}
+\end{figure}
+
+\section{Model}
+In this paper, we are interested in developing a model using simplex noise to reproduce nanoporous structures that are as statistically close to actual simulations of porous media.
+For a given data set and a model, we wish to estimate the parameters that best represent the data.
+This is performed through a full likelihood analysis, where we in addition validate the model by comparing the radial distribution function $g(r)$, porosity and surface area of the system.
+The simulated system is confined to a cube of size $159 \AA^3$, but since the procedural noise function is continuous on all scales, the box could be of arbitrary size.
+
+\begin{figure*}
+\includegraphics[width=.95\textwidth]{model_examples.png}
+\caption{Examples of 3D simplex noise structures in nanoporous media.
+Left image: low persistence (large scale dominates), middle image: medium persistence (small scales starts dominating) and right image: high threshold (more atoms are discarded).
+}
+\label{fig:model_example}
+\end{figure*}
+We construct a generic sum over procedural noise octaves with the expression: 
+\begin{equation}
+  M_1(\bar \theta) = M_1(\vec x, \psi, \Omega, \nu) = \sum_{\omega=0}^{\Omega} \psi^{\omega -1}   N\big((\vec x + \sigma \cdot \omega)\nu \cdot 2^{\omega-1} \big),
+\label{eq:noisemodel1}
+\end{equation}
+where $\Omega$ is the total amount of octaves, $\psi$ the persistence (scale falloff), $N$ a generic $N$-dimensional noise function, $\sigma$ a predefined random vector that removes anisotropic artefacts and $\nu$ the frequency modifier representing the initial largest scales in the model.
+For simplicity, $\bar \theta$ here denotes the set of model parameters.
+For each mode, the frequency is thus multiplied, while the overall amplitude is damped by a given persistence.
+In the end, we simulate a large set of bulk particles without any walls, and use the noise function to remove particles that are not in a procedural void given by a threshold $\tau$.
+Three examples of this model are depicted in figure \ref{fig:model_example}, where we have used $\Omega=6$ octaves and the scale is $\nu=0.01$ and cut-off $\tau=0$.
+The left image depicts low persistence ($\psi = 0.0$), where only the first octave dominates (corresponding to large scales).
+The middle picture shows the same parameters with a higher persistence $(\psi = 0.5)$, where smaller scales become more dominating.
+In the rightmost picture, the effect of the cut-off threshold $\tau = -0.3$ is shown.
+
+\subsection{Model Measure}
+In order to perform data analysis on the model parameters, we developed a code that uses Monte Carlo Markov chains in order to produce samples for building the likelihood.
+Before doing so, we need a model measure - a quantitative method that characterizes the internal structure of the porous media.
+This quantity should be computed for both the original data set and the one produced by our noise model.
+
+Ideally, the model measure should pick up all geometrical properties that are relevant for the physical study of interest.
+For instance, pore size distribution and porosity are import for studying transport in porous media.
+Another such measure is the Distance-To-Atom (DTA).
+The method used for calculating porosity of a media is defined as in \cite{gelb1998characterization}, and we obtain the surface area of a porous media by using marching cubes triangulation to sum the surface area of the triangles.
+
+In this paper, we choose to use the radial distribution function $g(r)$ as defined and implemented in LAMMPS \cite{plimpton1995fast}.
+The $g(r)$ (or pair correlation function) describes how density varies as a function of distance from a reference particle.
+Hereafter, $g_d(r)$ denotes the RDF calculated from simulated or experimental data, while $g_{M(\bar \theta)}(r)$ represents the RDF based on the procedural noise model $M(\bar \theta)$, where $\bar \theta$ are the set of model parameters.
+
+\todo{skriv litt mer om hvorfor vi velger g(r)}
+
+%\begin{figure}[htb!]
+%\includegraphics[width=.45\textwidth]{DTA.png}
+%\caption{Distance-to-atom measure for two data sets with varying persistence.
+The green line corresponds to a mock data set with low persistence (dominated by large-scale modes), while the blue line corresponds to high persistence (dominated by small scale modes).
+}
+%\label{fig:dta_example}
+%\end{figure}
+
+%\subsubsection{Fractal Dimension}
+%The fractal dimension is a ratio providing a statistical index of complexity comparing how detail in a pattern changes with the scale at which it is measured. We provide a fracal
+
+%\subsubsection{Octree Counting Measure (OCM)}
+
+%\subsection{Diffusion Measure}
+
+\subsection{Data analysis and the Likelihood}
+
+The posterior probability is defined as the probability of the parameters $\bar \theta$ given the evidence $d$, or $P(\bar \theta | d)$. Bayes' theorem states that 
+\begin{equation}
+P(\bar \theta | d) = \frac{P(d | \bar \theta)P(\bar \theta)}{P(d)} \propto P(d | \bar \theta)P(\bar \theta),
+\end{equation}
+where $P(d | \bar \theta)$ is called the Likelihood $\mathcal L$ (the probability of evidence given parameters) and $P(\bar \theta)$ is the prior (the beliefs about the quantity before some evidence is taken into account).
+$P(d)$ is called the evidence, and is the probability of the observations (here assumed to be 1).
+Since the prior is a constant, we can therefore assume that the posterior distribution is proportional to the likelihood, or $P(d | \bar \theta) \propto \mathcal L$.
+In other words, to calculate the posterior and obtaining the best fit model parameters, we need only to estimate the likelihood function, which is connected to the $\chi^2$ by
+\begin{equation}
+\mathcal L \propto e^{-\chi^2}.
+\end{equation}
+
+In order to test whether two 3D data sets are statistically equivalent up to the model measure $g(r)$, we calculate the $\chi^2$ between the two one-dimensional RDFs. Pearson's $\chi^2$ is defined as 
+\begin{equation}
+  \chi^2 = \sum_r \frac{ \Big(g_d(r) - g_{M(\bar \theta)}(r) \Big)^2}{g_{M(\bar \theta)}(r)^2},
+\end{equation}
+so for low values of $\chi^2$, the better the fit between data and model (up to $g(r)$).
+For a single-parameter model, we could simply just iterate over the lone parameter to obtain the minimum $\chi^2$ (where the best-fit parameter value is), however, with increased number of parameters comes increased computational cost.
+If we divide the parameter space into $N$ bins, the number of $\chi^2$ computations scales as $N^d$, where $d$ is the number of model parameters.
+To be able to build the full $\chi^2$ (and likelihood) parameter landscape, we there need to use Monte Carlo Markov chain methods (MCMC).
+
+We are now interested in maximising the likelihood $\mathcal L$ as opposed to minimising the $\chi^2$, with the goal of mapping out the full $d$-dimensional likelihood space for the model parameters.
+We build the $d$-dimensional likelihood by letting random walkers traverse the parameter space guided by the MCMC test: for each step $M_n(\bar \theta)$ in parameter space, we calculate the likelihood for this parameter configuration.
+Then, we create a new set of parameters for the next step $M_{n+1}(\bar \theta)$ by choosing new parameters randomly from a $d$-dimensional sphere around the previous parameter composition.
+The model also selects a new random seed.
+We calculate the likelihood at the new step and perform the MCMC test: accept the new step if $\frac{\mathcal L_n}{\mathcal L_n+1} \ge 1$, or reject if the value is lower than a uniform random value $\in [0,1]$.
+
+Finally, we obtain the posterior $P(\bar \theta | d) \propto \mathcal L$ by building a $d$-dimensional histogram of the likelihood using the stored positions of the parameters.
+For a sufficiently large amount of steps, the MCMC algorithm ensures that the posterior converges to the likelihood, as long at the surface does not contain unreachable areas such as discontinuous patches or an infinite amount of maxima (as is the case for the function $f(x)=sin(\frac{1}{x}$).
+Luckily, as we shall see, the likelihood surface for our models tend to turn out with a single, well-defined maximum.
+
+\section{Results}
+We now turn our attention to applying the MCMC framework on two kinds of data.
+First, we wish to test and verify the code by performing a full data analysis on mock data with known input parameters.
+This enables us to investigate the posterior distribution, and determine whether any potential problems could arise with the likelihood landscape, such as radom walkers trapped in local minima.
+We then continue with a full analysis of simulated SiO2 data, and see whether the resulting best-fit parameter data set is consistent with the SiO2 data set.
+
+\subsection{Model verification: Parameter estimation of mock data}
+\begin{figure}[htb!]
+\includegraphics[width=.45\textwidth]{model_test.png}
+\caption{Simulated mock data used for the initial model testing analysis. Input parameters are $\nu=0.03$,$\tau=-0.02$ and $\psi=0.6$.}
+\label{fig:mockdata}
+\end{figure}
+In order to verify our likelihood code, we perform a model verification test on a mock data set with known input values.
+In this example, we only use standard simplex noise with three free parameters: the persistence $\psi$, threshold $\tau$ and scale $\nu$.
+We simulate a data set of $159 ^3 Å$ of $\sim 255 000$ atoms, and perform the cut-off given input parameters $\Omega=3$, $\Psi = 0.6$, $\tau=-0.2$, $\nu=0.03$.
+The resulting 3D structure is depicted in figure \ref{fig:mockdata}.
+
+We continue by performing a full MCMC analysis of the mock data with using a parallelized code running on a 48 core cluster.
+In order to produce 300 000 samples, about 48 hours was needed.
+The results from this analysis can be seen in figure \ref{fig:mockdataresults_pts}, where the upper panel depicts the marginalised 1D posteriors while the lower shows the marginalised 2D posteriors.
+The centre and left image shows that the input parameters of $\tau=-0.2$(threshold), $\psi=0.6$(persistence) and $\nu=0.03$ are all reproduced well within $1 \sigma$.
+
+Note that there seems to be no problems with the posterior distribution in terms of local minima or multiple maxima.
+
+\begin{figure*}
+\includegraphics[width=.99\textwidth]{mock_data_results_pts.png}
+\caption{Results from the model validation simulation.
+The upper panel displays the marginalised 1D posteriors for each parameter, while the lower panel depicts the marginalised 2D posteriors.
+Note that all input parameters were reproduced well within $1\sigma$ (inner contour line).}
+\label{fig:mockdataresults_pts}
+\end{figure*}
+
+\subsection{Parameter estimation of simulated SiO2 silica}
+%\begin{figure}
+%\includegraphics[width=.49\textwidth]{sio_porous.png}
+%\caption{Simulated silica created with LAMMPS.
+The box is $159^3 \AA$, with $100 000$ particles, having been created by cooling an expanding system until quenching was achieved.
+Note the similarities between the pores here and in figure \ref{fig:mockdata}.}
+%\label{fig:si2data}
+%\end{figure}
+\begin{figure*}
+\includegraphics[width=.99\textwidth]{results_porous_full.png}
+\caption{Results from the analysis of a simulated $159^3$ \AA SiO2 nanoporous data set.
+The upper panel displays the marginalised 1D posteriors for each parameter, while the lower panel depicts the marginalised 2D posteriors.
+Best fit parameters from the maximum 3D likelihood suggest $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$.}
+\label{fig:porous_results1}
+\end{figure*}
+
+We now perform a parameter fitting analysis on realistic simulated nanoporous media.
+Using LAMMPS \cite{plimpton1995fast}, we simulate a $100^3\AA$  box of 100K SiO2 silica particles at an initial temperature of $4500 K$, using the Vashishta-potential.
+After an initial period, we expand the box to $159^3 \AA$ and start cooling until quenching is achieved at about $300K$.
+The resulting nanoporous media is depicted on the right side in figure \ref{fig:porous_vs_model}, with a porosity of $75\%$.
+\todo{Skrive mer om lagingen?}
+
+With the model described in equation \ref{eq:noisemodel1}, we perform a full likelihood analysis of the data set with the three free parameters $\tau$, $\nu$ and $\psi$ using $g(r)$ as the measure.
+The resulting marginalised posteriors are depicted in figure \ref{fig:porous_results1}, with best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$ is extracted from the full 3D likelihood.
+We also display the RFD $g(r)$ for the input data set, a best-fit parameter data set and a bad-fit data set in figure \ref{fig:gofr1}, to illustrated the sensitivity of $g(r)$ given the model parameters.
+Note that we are not able to accurately reproduce the structures of the SiO2 $g(r)$ with our current model, but hope that we during future work will be able to produce another model that provides a better fit with this measure.
+
+\begin{figure}
+\includegraphics[width=.45\textwidth]{gofr_plot.png}
+\caption{The RDF $g(r)$ for three data sets: the simulated SiO2 input data set (red), the best-fit parameter model (blue) and a random low-fit parameter model (green). }
+\label{fig:gofr1}
+\end{figure}
+Finally, we depict a comparison of the actual simulated input SiO2 data set together with a representation of the best-fit parameters in figure \ref{fig:porous_vs_model}   
+\begin{figure}
+\includegraphics[width=.45\textwidth]{comparison.png}
+\caption{Right side: A realisation of the best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$. Left side: Simulated SiO2 input data set used for parameter estimation. }
+\label{fig:porous_vs_model}
+\end{figure}
+
+The porosity of the input SiO data set is $75\%$, while the best-fit data set has porosity $77\% \pm 3 \%$, depending on the random seed.
+In addition, the surface area of the original data set was $\sim 29 000 \AA^2$, while the best-fit data shows similar values with surface are ranging between $\sim 28 000 \AA^2$ and $\sim 30 000 \AA^2$.
+\todo{legg ved 3D plot av likelihood for aa vise at den er kjekk og fin}
+
+\subsection{Some notes on the results}
+It is worth to note that since we are chiefly interested in obtaining a set of parameters that reproduces a statistical correct representation of the input data, we are relatively free to choose parameters as desired.
+In the previous analysis, we selected a best fit threshold value of $\tau=-0.2$, but from the width of the likelihood it is clear that nearby parameters would also be suitable.
+However, the surface area is strongly correlated with this value, so even slight deviations (e.g.
+$\Delta \tau = 0.05$) yields a relatively large change in surface area.
+
+In addition, we do not claim that this model accurately reproduces the pore structure observed in nanoporous simulations, as is evident from the RDF shown in figure \ref{fig:gofr1}.
+
+\subsection{$g(r)$ variation}
+We also calculated the variation in the RDF $g(r)$ given initial conditions on both our model and simulated data.
+In terms of the data, the variation is attributed as the initial velocities of the particles in the LAMMBS simulation.
+For the model, it corresponds to the random seed.
+We found that varying these initial conditions have negligible effect on the calculation of $g(r)$.
+
+%persistence : 0.875111019125
+%scale : 0.036638
+%threshold : -0.236784875   78% porosity
+
+% bad fit : scale 0.1, p = 1, threshold = -0.3
+
+\section{Conclusion}
+\todo{konklusjon maa renskrives}
+In this paper, we have proposed a novel method for generating a simple kind of nanoporous media using procedural methods, namely simplex noise.
+We have developed tools for creating such porous materials, and also estimating model parameters given data.
+We have shown that it is possible through a full likelihood analysis to reproduce the input parameters of a simulated model.
+In addition, we performed an analysis on a simulated $159^3 \AA$ data set of SiO2, and explain that while porosity and surface area corresponds well between the model and data, the RDF $g(r)$ shows that the current model does not exactly reproduce correct data.
+However, we do not claim that the current model should do so yet, since we have chosen the simplest kind of summation with regards to procedural noise modes.
+
+\subsection{Future work}
+In this paper, we have only considered the simplest kind of procedural noise: a simple sum over modes with threshold cut-off $\tau$, scale fall-off persistence $\psi$ and initial scale $\nu$.
+We have also seen that, even through both porosity and surface area accurately reproduced, there are features in the RDF that are not fully captured with the current model.
+This is evident from the images of the simulated structure seen in figure \ref{fig:porous_vs_model}, where the connecting nearly one-atom thin walls are difficult to reproduce with the current model.
+However, as procedural noise exert properties closely related to that of fractals, it is possible to extend the number of parameters and define new models with very different properties.
+Some examples of these models are depicted in figure \ref{fig:future_models}, where the left- and centre images show a new model with two new parameters (offset and gain), while the rightmost panel shows a simulation where the scale $\nu$ is itself being perturbed by procedural noise, yielding strong circular asymmetric patterns.
+
+Another immediate application would be to estimate model parameters given experimental observations of porous media.
+This would enable us to estimate new parameters such as instrumental noise, deviations from symmetry (yielding asymmetric properties) and pureness of samples.
+In addition, by analysing just a few samples and producing an accurate model, we are able to produce an arbitrary amount of statistically correct samples, free from instrumental noise and artefacts.
+
+A final key analysis would be to compare properties such as permeability through finite-element simulations of gas/liquid flow through our simulated structures, in addition to exerting shear forces, and comparing with actual simulated data from e.g.
+LAMMPS.
+Since the noise function is a continuous boolean function that only checks whether a position is within a wall or void given a threshold, we are not limited to any size of the simulated system, and simulated particles traversing the media has arbitrary range.
+
+%Minkowski functionals and topology, genus etc
+
+%Swiss noise
+
+\begin{figure*}
+\includegraphics[width=1\textwidth]{future_models.png}
+\caption{Example of future models.
+Left panel: A multi-fractal model with thin surfaces and large pores.
+Centre panel: a multi fractal with cave-like structures.
+Right panel: A model with strong asymmetric noise.
+}
+\label{fig:future_models}
+\end{figure*}
+
+\todo{Kanskje fake-latin burde fjernes? Usikker!}
 Ut hendrerit cursus libero, in scelerisque tortor sollicitudin accumsan.
@@
 
-\section{Method}
-\todo{her står det nå det samme som i abstract}
-In this paper, we aim to develop a method for producing statistically correct simulations of nanoporous media without the use of conventional Molecular Dynamics Simulator (MDS).
-In short, the proposed model is a binary function that returns either true if the input position is in a void / pore, or false if within a wall.
-By using procedural noise, we are not restricted to generating media within a limited box, as the noise functions is continuous on all scales.
-It is therefore possible to validated pores at theoretically infinite range.
-In addition, this method is extremely fast, requiring very little computational power to calculate whether a particle is within a pore or not.
-
-After presenting such a model, we choose a method to measure the structural properties of nanoporous media that are created either through regular methods (e.g.
-LAMMPS \cite{plimpton1995fast}) or with our method.
-The measure we selected is the radial pore size distribution $g(r)$, which is regarded as the de-facto standard method for capturing the structural properties of porous media.
-
-Using this measure, we define a full likelihood framework for estimating model parameters given any data set, and continue by validating our own framework by analysing procedural noise data sets with known input parameters.
-Finally, we perform a full likelihood analysis of
-
-\subsection{Procedural noise}
-\label{sec:perlin}
-
-Procedural algorithmic methods for creating realistic-looking
-fractal-like patterns have been around for more than 30 years. The most famous of these
-algorithms is Perlin noise, first developed by Ken Perlin in 1982 for
-use in the movie ``Tron''.  Perlin noise is a type of pseudo-random
-gradient noise that mimics a Gaussian field, but at low computational cost. 
-An improved version of Perlin noise is called Simplex
-noise, which has several advantages over standard Perlin noise in
-terms of scalability and speed.
-
-\subsubsection{Perlin noise - an overview}
-An overview of the current state of procedural noise functions can be
-found in \cite{lagae:2010}.  Perlin noise \citep{perlin:2002} 
-is a band limited repeatable pseudo-random function
-from $\mathbb R^n \to \mathbb R$ that is computationally fast and has
-properties which make it suitable for creating patterns that mimic
-nature. Perlin noise is an approximation to Gaussian filtered noise
-implemented as a pseudo-random spline. The algorithm uses a predefined
-grid of gradients pointing in a random direction, where for any point
-within a grid, we interpolate the gradients from the four corners. The
-algorithm is as follows:
-\begin{itemize}
-  \item[1.] For a point, obtain the four closest gradients in the grid.
-   \item [2.] For each of the gradients, calculate the dot product
-     between the gradient and a vector defined by the  the distance
-     between the the point and the current grid corner. 
-   \item [3.] Instead of linear interpolation $F(t) = t$, use a fade function to produce smoother interpolated values.
-Perlin noise typically uses the following fade function: $F(t) = t ^3 (t  (6 - 15) + 10)$.
-    \item [4.] Obtain the final value by linearly interpolating between the x-axis and
-      use this result to linearly interpolate the y-axis. 
-\end{itemize} 
-Simplex noise works in a similar manner, but requires fewer
-computational steps, scales better in higher ($\ge 3$) dimensions and has fewer 
-anisotropic properties.
-
-\subsubsection{Properties of procedural noise}
-\begin{figure}
-\includegraphics[width=.5\textwidth]{noise.eps}
-\caption{Perlin noise (left) versus simplex noise (right). Note the
-  straight x-y pattern in the single upper Perlin mode and the less
-  anisotropic simplex mode. Lower part: combining octaves tend
-  to remove these patterns.}
-\label{fig:noise}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=.5\textwidth]{noise_comparison.png}
-%\mbox{\epsfig{file=perlinstats.eps, width=\linewidth}}
-\caption{Normalized distributions of Perlin noise and simplex noise compared with a
-  Gaussian distribution. Note how both noise functions produce
-  near-Gaussian, but slightly different distributions.}
-\label{fig:properties}
-\end{figure}
-Standard Perlin noise is known for having several anisotropic features
-\citep{lagae:2010}. Perlin noise also gives rise to a ``textile''
-pattern in the $x$ and $y$-directions, which is evident from the upper
-left part of figure \ref{fig:noise}, depicting 2D Perlin noise in
-a single octave. The upper-right part shows the same node for Simplex noise,
-with less obvious patterns. However, when combining several octaves,
-as depicted in the two lower parts, these anisotropic patterns tend to cancel
-out. In any case, for observations of weak gravitational lensing, these anisotropies would occur on much smaller 
-scales than the typical resolution. 
-In addition, we have calculated the statistical distributions for Perlin and simplex noise
-with $\sigma = 0.3$, and have plotted the results in figure \ref{fig:properties}.
-Each procedural noise method has distributions that are close to normal, 
-but with some asymmetrical properties.
-
-\subsection{Multiple octaves: creating patterns}
-\label{sec:octaves}
-The different octaves from procedural noise functions can be combined
-to produce a wide range of nature-like patterns. A generic
-N-dimensional procedural function $\Phi(\vec x): \mathbb R^N \to
-\mathbb R$ could be expressed as
-\begin{equation}
-\label{eq:procedural}
- \Phi(\vec x) = \Theta \Big(\sum_k \kappa (k) P(  f_s (\vec x,k) )) \Big),
- \end{equation}
-where $f_s(\vec x,k)$ describes the frequency scale modifier (e.g. $f_s(\vec x,k) =
-k\vec x$), $\kappa(k)$ is the octave amplitude (eg $\kappa(k) = 1/k$)
-and $\Theta(x)$ an overall modifier (e.g. $\Theta(x) = x$ or
-$\Theta(x) = \frac{1}{x}$). In a sense, $\kappa(k)$ is similar to Fourier coefficients. 
-In its simplest case, summing directly
-over the octaves with amplitude $1/k$ results in 
-\begin{equation}
-\label{eq:perlinlinear}
- \Phi(\vec x) = \sum_k \frac{1}{k} P( 2k\vec x), 
-\end{equation}
-and produces an 2D procedural noise pattern depicted in the lower
-part of figure \ref{fig:noise}.
-Finally, since simplex noise both scales better with higher dimensions and produces less asymmetric effects, we choose to use this noise method over regular Perlin noise in our method.
-
-\subsection{Model}
-In this paper, we are interested in developing a model using Simplex/Perlin noise to reproduce nanoporous structures that are as statistically close to actual simulations of porous media.
-For a given data set and a model, we wish to estimate the parameters that best represent the data.
-This is performed through a full likelihood analysis, where we in addition validate the model by comparing the radial pore size distribution $g(r)$, porosity and surface area of the system.
-
-\begin{figure*}
-\includegraphics[width=.95\textwidth]{model_examples.png}
-\caption{Examples of 3D simplex noise structures in nanoporous media.
-Left image: low persistence (large scale dominates), middle image: medium persistence (small scales starts dominating) and right image: high threshold (more atoms are discarded).
-}
-\label{fig:model_example}
-\end{figure*}
-We construct a generic sum over procedural noise octaves with the expression: 
-\begin{equation}
-  M_1(\vec x) = \sum_{\omega=0}^{\Omega} \psi^{\omega -1}   N\big((\vec x + \sigma \cdot \omega)\nu \cdot 2^{\omega-1} \big),
-\label{eq:noisemodel1}
-\end{equation}
-where $\Omega$ is the total amount of octaves, $\psi$ the persistence (scale falloff), $N$ a generic $N$-dimensional noise function, $\sigma$ a predefined random vector that removes anisotropic artefacts and $\nu$ the frequency modifier representing the initial largest scales in the model.
-For each mode, the frequency is thus multiplied, while the overall amplitude is damped by a given persistence.
-In the end, we simulate a large set of bulk particles without any walls, and use the noise function to remove particles that are not in a procedural void given by a threshold $\tau$.
-Three examples of this model are depicted in figure \ref{fig:model_example}, where we have used $\Omega=6$ octaves and the scale is $\nu=0.01$ and cut-off $\tau=0$.
-The left image depicts low persistence ($\psi = 0.0$), where only the first octave dominates (corresponding to large scales).
-The middle picture shows the same parameters with a higher persistence $(\psi = 0.5)$, where smaller scales become more dominating.
-In the rightmost picture, the effect of the cut-off threshold $\tau = -0.3$ is shown.
-
-\subsection{Model Measure}
-In order to perform data analysis on the model parameters, we developed a code that uses Monte Carlo Markov chains in order to produce samples for building the likelihood.
-Before doing so, we need a model measure - a quantitative method that characterizes the internal structure of the porous media.
-This quantity should be computed for both the original data set and the one produced by our noise model.
-
-Ideally, the model measure should pick up all geometrical properties that are relevant for the physical study of interest.
-For instance, pore size distribution and porosity are import for studying transport in porous media.
-Another such measure is the Distance-To-Atom (DTA).
-The method used for calculating the porosity of a media is defined as in \cite{gelb1998characterization}, and we obtain the surface area of a porous media by using marching cubes triangulation to sum the surface area of the triangles.
-
-In this paper, we choose to use the radial pore size distribution $g(r)$ as defined and implemented in LAMMPS \cite{plimpton1995fast}.  \todo{skriv litt mer om hvorfor vi velger g(r)}
-
-%\begin{figure}[htb!]
-%\includegraphics[width=.45\textwidth]{DTA.png}
-%\caption{Distance-to-atom measure for two data sets with varying persistence.
-The green line corresponds to a mock data set with low persistence (dominated by large-scale modes), while the blue line corresponds to high persistence (dominated by small scale modes).
-}
-%\label{fig:dta_example}
-%\end{figure}
-
-%\subsubsection{Fractal Dimension}
-%The fractal dimension is a ratio providing a statistical index of complexity comparing how detail in a pattern changes with the scale at which it is measured. We provide a fracal
-
-%\subsubsection{Octree Counting Measure (OCM)}
-
-%\subsection{Diffusion Measure}
-
-\subsection{Likelihood}
-The likelihood $\mathcal{L}$ of parameters $\theta$ of a model $M(\theta)$ given outcome $x$ is defined as $\mathcal{L} (M(\theta) | x) = P( x | M(\theta))$, where P is the probability distribution.
-In our model, we use the $\chi^2$ test to fit our one-dimensional measure from the data $\bf d$ with our model $\bf M(\theta)$.
-Pearson's $\chi^2$ is defined as
-\begin{equation}
-  \chi^2 = \sum_i \frac{ \Big(d_i - M(\theta)_i \Big)^2}{M(\theta)_i^2},
-\end{equation}    
-and the likelihood is thus
-\begin{equation}
-\mathcal L \propto e^{-\chi^2}.
-\end{equation}
-For a model with $N$ free parameters, we build the full $N$-dimensional likelihood by letting random walkers map out the parameter space through the Monte Carlo Markov chain method (MCMC).
-For each step $M(\theta)_n$ in parameter space, we calculate the likelihood for this parameter configuration.
-Then, we create a new set of parameters for the next step $M(\theta)_{n+1}$ by choosing new parameters randomly from a $N$-dimensional sphere around the previous parameters.
-The model also chooses a new random seed.
-We calculate the likelihood at the new step and perform the MCMC test: Accept the new step if $\frac{\mathcal L_n}{\mathcal L_n+1} \ge 1$, or reject if the value is lower than a uniform random value $\in [0,1]$.
-
-\section{Model verification and data analysis}
-\begin{figure}[htb!]
-\includegraphics[width=.45\textwidth]{model_test.png}
-\caption{Simulated mock data used for the initial model testing analysis. Input parameters are $\nu=0.03$,$\tau=-0.02$ and $\psi=0.6$.}
-\label{fig:mockdata}
-\end{figure}
-In order to verify our likelihood code, we perform a model verification test on a mock data set with known input values.
-In this example, we only use standard simplex noise with three free parameters: the persistence $\psi$, threshold $\tau$ and scale $\nu$.
-We simulate a data set of $159 ^3 Å$ of $1 000 000$ atoms, and perform the cut-off given input parameters $\Omega=3$, $\Psi = 0.6$, $\tau=-0.2$, $\nu=0.03$.
-The resulting 3D structure is depicted in figure \ref{fig:mockdata}.
-
-We continue by performing a full MCMC analysis of the mock data with using a parallelized code running on a 48 core cluster.
-In order to produce 300 000 samples, about 48 hours was needed.
-The results from this analysis can be seen in figure \ref{fig:mockdataresults_pts}, where the upper panel depicts the marginalised 1D posteriors while the lower shows the marginalised 2D posteriors.
-The centre and left image shows that the input parameters of $\tau=-0.2$(threshold), $\psi=0.6$(persistence) and $\nu=0.03$ are all reproduced well within $1 \sigma$.
-
-\begin{figure*}
-\includegraphics[width=.99\textwidth]{mock_data_results_pts.png}
-\caption{Results from the model validation simulation.
-The upper panel displays the marginalised 1D posteriors for each parameter, while the lower panel depicts the marginalised 2D posteriors.
-Note that all input parameters were reproduced well within $1\sigma$ (inner contour line).}
-\label{fig:mockdataresults_pts}
-\end{figure*}
-
-\section{Results}
-%\begin{figure}
-%\includegraphics[width=.49\textwidth]{sio_porous.png}
-%\caption{Simulated silica created with LAMMPS.
-The box is $159^3 \AA$, with $100 000$ particles, having been created by cooling an expanding system until quenching was achieved.
-Note the similarities between the pores here and in figure \ref{fig:mockdata}.}
-%\label{fig:si2data}
-%\end{figure}
-\begin{figure*}
-\includegraphics[width=.99\textwidth]{results_porous_full.png}
-\caption{Results from the analysis of a simulated $159^3$ \AA $SiO^2$ nanoporous data set.
-The upper panel displays the marginalised 1D posteriors for each parameter, while the lower panel depicts the marginalised 2D posteriors.
-Best fit parameters from the maximum 3D likelihood suggest $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$.}
-\label{fig:porous_results1}
-\end{figure*}
-
-We now perform a parameter fitting analysis on realistic simulated nanoporous media.
-Using LAMMPS \cite{plimpton1995fast}, we simulate a $100^3$ \AA box of 100K $SiO^2$ silica particles at an initial temperature of $4500 K$, using the Vashishta-potential.
-After an initial period, we expand the box to $159^3 \AA$ and start cooling until quenching is achieved at about $300K$.
-The resulting nanoporous media is depicted in figure \ref{fig:sio2data}, with a porosity of $75\%$.
-
-With the model described in \ref{eq:noisemodel1}, we perform a full likelihood analysis of the data set with the three free parameters $\tau$, $\nu$ and $\psi$ using $g(r)$ as the measure.
-The resulting marginalised posteriors are depicted in figure \ref{fig:porous_results1}, with best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$ is extracted from the full 3D likelihood.
-We also display the radial pore size distribution $g(r)$ for the input data set, a best-fit parameter data set and a bad-fit data set in figure \ref{fig:gofr1}, to illustrated the sensitivity of $g(r)$ given the model parameters.
-Note that we are not able to accurately reproduce the structures of the $SiO^2$ $g(r)$ with our current model, but hope that we during future work will be able to produce another model that provides a better fit with this measure.
-
-\begin{figure}
-\includegraphics[width=.55\textwidth]{gofr_plot.png}
-\caption{The radial pore size distribution $g(r)$ for three data sets: the simulated $SiO^2$ input data set, the best-fit parameter data set and a random low-fit parameter data set. }
-\label{fig:gofr1}
-\end{figure}
-Finally, we depict a comparison of the actual simulated input $SiO^2$ data set together with a representation of the best-fit parameters in figure \ref{fig:porous_vs_model}   
-\begin{figure}
-\includegraphics[width=.45\textwidth]{comparison.png}
-\caption{Right side: A realisation of the best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$. Left side: Simulated $SiO^2$ input data set used for parameter estimation. }
-\label{fig:porous_vs_model}
-\end{figure}
-
-The porosity of the input $SiO^2$ data set is $75\%$, while the best-fit data set has porosity $77\% \pm 3 \%$, depending on the random seed.
-In addition, the surface area of the original data set was $\tilde 29 000 \AA^2$, while the best-fit data shows similar values with surface are ranging between $\tilde 28 000 \AA^2$ and $\tilde 30 000 \AA^2$.
-
-\subsection{Some notes on the results}
-It is worth to note that since we are chiefly interested in obtaining a set of parameters that reproduces a statistical correct representation of the input data, we are relatively free to choose parameters as desired.
-In the previous analysis, we selected a best fit threshold value of $\tau=-0.2$, but from the width of the likelihood it is clear that nearby parameters would also be suitable.
-However, the surface area is strongly correlated with this value, so even slight deviations (e.g.
-$\Delta \tau = 0.05$) yields a relatively large change in surface area.
-
-In addition, we do not claim that this model accurately reproduces the pore structure observed in nanoporous simulations, as is evident from the radial pore size distribution shown in figure \ref{fig:gofr1}.
-
-%persistence : 0.875111019125
-%scale : 0.036638
-%threshold : -0.236784875   78% porosity
-
-% bad fit : scale 0.1, p = 1, threshold = -0.3
-
-\section{Conclusion}
-In this paper, we have proposed a novel method for generating a simple kind of nanoporous media using procedural methods, namely Simplex noise.
-We have developed tools for creating such porous materials, and also estimating model parameters given data.
-We have shown that it is possible through a full likelihood analysis to reproduce the input parameters of a simulated model.
-In addition, we performed an analysis on a simulated $159^3 \AA$ data set of $SiO^2$, and explain that while porosity and surface area corresponds well between the model and data, the radial pore size distribution $g(r)$ show that the current model does not exactly reproduce correct data.
-However, we do not claim that the current model should do so yet, since we have chosen the simplest kind of summation with regards to procedural noise modes.
-
-\subsection{Future work}
-In this paper, we have only considered the simplest kind of procedural noise: a simple sum over modes with threshold cut-off $\tau$, scale fall-off persistence $\psi$ and initial scale $\nu$.
-We have also seen that, even through both porosity and surface area accurately reproduced, there are features in the radial pore size distribution $g(r)$ that are not fully captured with the current model.
-This is evident from the images of the simulated structure seen in \ref{fig:porous_vs_model}, where the connecting nearly one-atom thin walls are difficult to reproduce with the current model.
-However, as procedural noise exert properties closely related to that of fractals, it is possible to extend the number of parameters and define new models with very different properties.
-Some examples of these models are depicted in figure \ref{fig:future_models}, where the left- and centre images show a new model with two new parameters (offset and gain), while the rightmost panel shows a simulation where the scale $\nu$ is itself being perturbed by procedural noise, yielding strong circular asymmetric patterns.
-
-Another immediate application would be to estimate model parameters given actual observations of porous media.
-This would enable us to estimate new parameters such as instrumental noise, deviations from symmetry (yielding asymmetric properties) and  pureness of samples.
-In addition, by analysing just a few samples and producing an accurate statistical model, we are able to produce an arbitrary amount of statistically correct samples, free from instrumental noise and artefacts.
-
-A final key analysis would be to compare properties such as percolation through finite-element simulations of gas/liquid flow through our simulated structures, in addition to exerting shear forces, and comparing with actual simulated data from e.g.
-LAMMPS.
-
-\begin{figure*}
-\includegraphics[width=1\textwidth]{future_models.png}
-\caption{Example of future models.
-Left panel: A multi-fractal model with thin surfaces and large pores.
-Centre panel: a multi fractal with cave-like structures.
-Right panel: A model with strong asymmetric noise.
-}
-\label{fig:future_models}
-\end{figure*}
-
-Ut hendrerit cursus libero, in scelerisque tortor sollicitudin accumsan.
-Nam eleifend velit metus, quis volutpat justo faucibus ut.
-Cras ut sem in nunc fringilla egestas vitae sit amet justo.
-Ut vel condimentum tortor.
-Morbi in massa in ipsum ultrices tincidunt.
-Duis feugiat dignissim nisi ac ultricies.
-Quisque eget risus fermentum, condimentum massa ullamcorper, posuere neque.
-Nam congue consequat mi, vel commodo mi tempus eget.
-
-Nullam ultrices, velit sed venenatis ultrices, risus urna ultricies magna, at pretium lorem nibh et nunc.
-Suspendisse sed mattis leo.
-Nulla facilisi.
-Praesent finibus lobortis leo, ut porttitor massa tincidunt ut.
-Vivamus quis orci eu sem sodales ornare.
-Cras sed faucibus velit.
-Phasellus scelerisque mauris quis lorem dignissim vehicula.
-Sed tempor ut urna sed bibendum.
-Donec et nibh pulvinar, commodo magna ac, interdum nibh.
-Sed laoreet ipsum quis quam lacinia, in dictum erat laoreet.
-Etiam rhoncus mauris arcu, ac egestas lacus facilisis nec.
-Pellentesque a orci vitae felis faucibus ultrices sed non massa.
-In posuere vel ex in elementum.
-Aenean ac neque a enim gravida tristique eu ac lacus.
-Ut tincidunt venenatis lectus, et maximus mauris.
-
-Proin ornare nulla id tristique placerat.
-Integer nec elementum tellus.
-Cras at tempus lectus.
-Cras lorem est, ornare id condimentum vitae, scelerisque eu lectus.
-Aenean libero arcu, consequat vel sagittis eget, sagittis ut neque.
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
-Aenean nec quam sodales, mollis est sit amet, vulputate turpis.
-Praesent tempus diam eu sem laoreet finibus.
-Nulla finibus, sapien vel lobortis ultricies, mauris ipsum aliquet urna, quis rutrum sapien dui id ipsum.
-Quisque sodales ex sapien, ac egestas lacus suscipit et.
-Suspendisse accumsan rhoncus tristique.
-Pellentesque ut eros lectus.
-Praesent viverra sit amet odio nec vulputate.
-Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
-Nulla molestie metus vitae libero vulputate, at interdum enim tempus.
-\subsection{Simulated data}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-02-04 ms.tex
Small fixes at Hyttetur

@@
 For instance, pore size distribution and porosity are import for studying transport in porous media.
-Another such measure is the Distance-To-Atom (DTA).
-The method used for calculating porosity of a media is defined as in \cite{gelb1998characterization}, and we obtain the surface area of a porous media by using marching cubes triangulation to sum the surface area of the triangles.
-
 In this paper, we choose to use the radial distribution function $g(r)$ as defined and implemented in LAMMPS \cite{plimpton1995fast}.
@@
 Hereafter, $g_d(r)$ denotes the RDF calculated from simulated or experimental data, while $g_{M(\bar \theta)}(r)$ represents the RDF based on the procedural noise model $M(\bar \theta)$, where $\bar \theta$ are the set of model parameters.
+
+The method used for calculating porosity of a media is defined as in \cite{gelb1998characterization}, and we obtain the surface area of a porous media by using marching cubes triangulation to sum the surface area of the triangles.
 
@@
 In this example, we only use standard simplex noise with three free parameters: the persistence $\psi$, threshold $\tau$ and scale $\nu$.
-We simulate a data set of $159 ^3 Å$ of $\sim 255 000$ atoms, and perform the cut-off given input parameters $\Omega=3$, $\Psi = 0.6$, $\tau=-0.2$, $\nu=0.03$.
+We simulate a data set of $159 ^3 \AA$ of $\sim 255 000$ atoms, and perform the cut-off given input parameters $\Omega=3$, $\Psi = 0.6$, $\tau=-0.2$, $\nu=0.03$.
 The resulting 3D structure is depicted in figure \ref{fig:mockdata}.

### 2017-02-05 ms.tex
Rewrote some of abstract, added references and updated some noise discussion

@@
 \begin{abstract} 
-In this paper, we aim to develop a method for producing statistically correct simulations of nanoporous media without the use of conventional Molecular Dynamics Simulator (MDS).
-In short, the proposed model is a binary function that returns either true if the input position is in a void / pore, or false if within a wall.
+In this paper, we aim to develop a method for producing statistically correct porous media using noise functions.
+In short, the proposed model is a binary function that returns either true if the input point is inside pore space, or false if within a wall.
 By using procedural noise, we are not restricted to generating media within a limited box, as the noise functions is continuous on all scales.
-It is therefore possible to validated pores at theoretically arbitrary range.
-In addition, this method is extremely fast, requiring very little computational power to calculate whether a particle is within a pore or not.
+It is therefore possible to generate geometries at arbitrary scales.
+In addition, this method is extremely fast and low memory footprints, requiring very little computational power.
 After presenting such a model, we choose a method to measure the structural properties of nanoporous media that are created either through regular methods (e.g.
-LAMMPS) or with our method.
-The measure we selected is the radial distribution function (RDF) $g(r)$, which is regarded as the de-facto standard method for capturing the structural properties of porous media.
+molecular dynamics) or with our method.
+The pair distribution function $g(r)$ is chosen as a measure of the structural properties to compare the simulated geometry to our model.
 Using this measure, we define a full likelihood framework for estimating model parameters given any data set, and continue by validating our own framework by analysing procedural noise data sets with known input parameters.
-Finally, we perform a full likelihood analysis of silica SiO2 data set simulated in LAMMPS, and show that our model correctly reproduces several of the internal properties of the data set, including $g(r)$, porosity and surface area.
+Finally, we generate a nanoporous media with an expand-and-quench technique using molecular dynamics and show that our model correctly reproduces several of the internal properties of the data set, including $g(r)$, porosity and surface area.
 \todo{Abstract boer tenkes noe gjennom: faa frem hovedpoenger.}
@@
 This means that it is possible to simulate
-
+\todo{Nevn finite size effects}
 %In addition, by creating a model that yields properties that are statistically consistent with observations, we are ar
@@
 We do not claim that procedural noise methods completely reproduces all structural properties of nanoporous silica, but show that our approach can be used to efficiently create materials that are structurally similar to simulations.
-Naturally, this all depends on the statistics used for comparing the two models, which in this case is based on the radial distribution function (RDF) $g(r)$.
-
+Naturally, this all depends on the statistics used for comparing the two models, which in this case is based on the pair correlation function $g(r)$.
+\todo{Nevne at kjemi naturligvis ikke vil være riktig siden antall naboer per atom ikke er riktig. }
 %\section{Method}
@@
 fractal-like patterns have been around for more than 30 years. The most famous of these
-algorithms is Perlin noise, first developed by Ken Perlin in 1982 for
-use in the movie ``Tron''.
+algorithms is Perlin noise, first developed by Ken Perlin in 1983 for
+use in the movie ``Tron''\cite{perlin1985image}.
 
 Perlin noise is a type of pseudo-random gradient noise that mimics a Gaussian field, but at low computational cost.
-An improved version of Perlin noise is called simplex noise, which has several advantages over standard Perlin noise in terms of scalability and speed.
+An improved version of Perlin noise is called simplex noise, which has several advantages over standard Perlin noise in terms of scalability and speed \cite{perlin:2002}.
 Perlin noise was developed in response to the lack of detail 1980's contemporary graphics, where polygonal surfaces often were bland and left without any detail.
@@
 An overview of the current state of procedural noise functions can be
-found in \cite{lagae:2010}.  Perlin noise \citep{perlin:2002} 
-is a band limited repeatable pseudo-random function
+found in \cite{lagae:2010}. Perlin noise is a band limited repeatable pseudo-random function
 from $\mathbb R^n \to \mathbb R$ that is computationally fast, and has
@@
 \end{itemize} 
-Simplex noise works in a similar manner, but requires fewer
-computational steps, scales better in higher ($\ge 3$) dimensions and has fewer 
-anisotropic properties.
+Simplex noise works in a similar manner, but works on on simplices instead of hypergrids, and so requires fewer computational steps, scales better in higher ($\ge 3$) dimensions and has fewer anisotropic properties.
 
-\subsection{Properties of Perlin/Simplex noise}
+\subsection{Properties of Perlin/simplex noise}
 \begin{figure}

### 2017-02-05 ms.tex
Wrote more about g(r), lammps simulation

@@
 \usepackage{epstopdf}
+\usepackage{siunitx}
 \usepackage{color}
@@
 We do not claim that procedural noise methods completely reproduces all structural properties of nanoporous silica, but show that our approach can be used to efficiently create materials that are structurally similar to simulations.
-Naturally, this all depends on the statistics used for comparing the two models, which in this case is based on the pair correlation function $g(r)$.
+Naturally, this all depends on the statistics used for comparing the two models, which in this case is based on the radial distribution function $g(r)$ (RDF).
 \todo{Nevne at kjemi naturligvis ikke vil være riktig siden antall naboer per atom ikke er riktig. }
@@
 This is performed through a full likelihood analysis, where we in addition validate the model by comparing the radial distribution function $g(r)$, porosity and surface area of the system.
-The simulated system is confined to a cube of size $159 \AA^3$, but since the procedural noise function is continuous on all scales, the box could be of arbitrary size.
+The simulated system is confined to a cube of size $\SI{159}{\angstrom^{3}}$, but since the procedural noise function is continuous on all scales, the box could be of arbitrary size.
 
@@
 Ideally, the model measure should pick up all geometrical properties that are relevant for the physical study of interest.
-For instance, pore size distribution and porosity are import for studying transport in porous media.
-In this paper, we choose to use the radial distribution function $g(r)$ as defined and implemented in LAMMPS \cite{plimpton1995fast}.
-The $g(r)$ (or pair correlation function) describes how density varies as a function of distance from a reference particle.
+For instance, transport in porous media heavily depends on the pore size distribution and porosity, whereas chemical reactons are sensitive to surface area.
+\todo{referanser, malthe?} In this paper, we choose to focus on the radial distribution function $g(r)$ as defined and implemented in LAMMPS \cite{plimpton1995fast}.
+$g(r)$ describes how density varies as a function of distance from a reference particle, picking up both the intramolecular and intermolecular structure.
+We have chosen to ignore the short range contributions of $g(r)$ since the pore geometry does not depend on the intramolecular structure of SiO$_2$, but only the density variations at larger distances.
 Hereafter, $g_d(r)$ denotes the RDF calculated from simulated or experimental data, while $g_{M(\bar \theta)}(r)$ represents the RDF based on the procedural noise model $M(\bar \theta)$, where $\bar \theta$ are the set of model parameters.
 
-The method used for calculating porosity of a media is defined as in \cite{gelb1998characterization}, and we obtain the surface area of a porous media by using marching cubes triangulation to sum the surface area of the triangles.
+The method used for calculating porosity of a media is defined as in \cite{gelb1998characterization}, and we obtain the surface area of a porous media by using marching cubes triangulation to sum the surface area of the generated triangles.
+As discussed in \cite{gelb1998characterization}, the definition of surface area is ill-defined in atomic simulations, because a definition of pore space is needed.
+In the marching cubes triangulation, we generate a voxelized grid, the scalar field, with values being the distance from the voxel center to the nearest atom.
+The isovalue threshold defining the surface is chosen to be \SI{2}{\angstrom}, large nough to have a dense pore matrix, but small enough to have most of the available pore space defined as such.
+The important note here is that the same definition of surface is being used in all compared systems.
 
@@
 This enables us to investigate the posterior distribution, and determine whether any potential problems could arise with the likelihood landscape, such as radom walkers trapped in local minima.
-We then continue with a full analysis of simulated SiO2 data, and see whether the resulting best-fit parameter data set is consistent with the SiO2 data set.
+We then continue with a full analysis of simulated SiO$_2$ data, and see whether the resulting best-fit parameter data set is consistent with the SiO$_2$ data set.
 
@@
 
-\subsection{Parameter estimation of simulated SiO2 silica}
+\subsection{Parameter estimation of simulated nanoporous silica}
 %\begin{figure}
@@
 \includegraphics[width=.99\textwidth]{results_porous_full.png}
-\caption{Results from the analysis of a simulated $159^3$ \AA SiO2 nanoporous data set.
+\caption{Results from the analysis of a simulated $159^3$ \AA SiO$_2$ nanoporous data set.
 The upper panel displays the marginalised 1D posteriors for each parameter, while the lower panel depicts the marginalised 2D posteriors.
@@
 We now perform a parameter fitting analysis on realistic simulated nanoporous media.
-Using LAMMPS \cite{plimpton1995fast}, we simulate a $100^3\AA$  box of 100K SiO2 silica particles at an initial temperature of $4500 K$, using the Vashishta-potential.
-After an initial period, we expand the box to $159^3 \AA$ and start cooling until quenching is achieved at about $300K$.
-The resulting nanoporous media is depicted on the right side in figure \ref{fig:porous_vs_model}, with a porosity of $75\%$.
-\todo{Skrive mer om lagingen?}
+Using LAMMPS \cite{plimpton1995fast} and the Vashishta-potential \cite{vashishta1990interaction}, we simulate system consisting of $14^3$ betacristobalite unit cells of SiO$_2$, corresponding to system size of approximately \SI{100}{\angstrom}.
+The temperature is then increased to \SI{4500}{\kelvin} using NVT integration.
+After an initial period, we expand the box to \SI{159}{\angstrom} to achieve a porosity $\phi=0.75$.
+Then, the system is quenched by setting the temperature to \SI{300}{\kelvin}, resulting in the nanoporous media on the right side in figure \ref{fig:porous_vs_model}.
+By choosing a different seed for the initial velocities, it is easy to produce other samples.
 
@@
 The resulting marginalised posteriors are depicted in figure \ref{fig:porous_results1}, with best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$ is extracted from the full 3D likelihood.
-We also display the RFD $g(r)$ for the input data set, a best-fit parameter data set and a bad-fit data set in figure \ref{fig:gofr1}, to illustrated the sensitivity of $g(r)$ given the model parameters.
-Note that we are not able to accurately reproduce the structures of the SiO2 $g(r)$ with our current model, but hope that we during future work will be able to produce another model that provides a better fit with this measure.
+We also display $g(r)$ for the input data set, a best-fit parameter data set and a bad-fit data set in figure \ref{fig:gofr1}, to illustrate the sensitivity of $g(r)$ given the model parameters.
+Note that we are not able to accurately reproduce the structures of the SiO$_2$ $g(r)$ with our current model, but hope that we during future work will be able to produce another model that provides a better fit with this measure.
+
+Note that the surface atoms are not in chemical equilibrium since some neighbors of a surface atom might have been removed.
+In order to use the generated geometry in molecular dynamics, one would need either noble gas atoms which may be stable, or apply passivation step.
 
@@
 \includegraphics[width=.45\textwidth]{gofr_plot.png}
-\caption{The RDF $g(r)$ for three data sets: the simulated SiO2 input data set (red), the best-fit parameter model (blue) and a random low-fit parameter model (green). }
+\caption{$g(r)$ for three data sets: the simulated SiO$_2$ input data set (red), the best-fit parameter model (blue) and a random low-fit parameter model (green).
+\todo{Bytt denne figuren med VMD sin beregning.
+Normaliseringen vår er gal.}}
 \label{fig:gofr1}
 \end{figure}
-Finally, we depict a comparison of the actual simulated input SiO2 data set together with a representation of the best-fit parameters in figure \ref{fig:porous_vs_model}   
+Finally, we depict a comparison of the actual simulated input SiO$_2$ data set together with a representation of the best-fit parameters in figure \ref{fig:porous_vs_model}   
 \begin{figure}
 \includegraphics[width=.45\textwidth]{comparison.png}
-\caption{Right side: A realisation of the best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$. Left side: Simulated SiO2 input data set used for parameter estimation. }
+\caption{Right side: A realisation of the best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$. Left side: Simulated SiO$_2$ input data set used for parameter estimation. }
 \label{fig:porous_vs_model}
@@
 \subsection{$g(r)$ variation}
-We also calculated the variation in the RDF $g(r)$ given initial conditions on both our model and simulated data.
-In terms of the data, the variation is attributed as the initial velocities of the particles in the LAMMBS simulation.
+We also calculated the variation in the $g(r)$ given initial conditions on both our model and simulated data.
+In terms of the data, the variation is attributed as the initial velocities of the particles in the LAMMPS simulation.
 For the model, it corresponds to the random seed.
@@
 We have shown that it is possible through a full likelihood analysis to reproduce the input parameters of a simulated model.
-In addition, we performed an analysis on a simulated $159^3 \AA$ data set of SiO2, and explain that while porosity and surface area corresponds well between the model and data, the RDF $g(r)$ shows that the current model does not exactly reproduce correct data.
+In addition, we performed an analysis on a simulated $159^3 \AA$ data set of SiO$_2$, and explain that while porosity and surface area corresponds well between the model and data, the $g(r)$ shows that the current model does not exactly reproduce correct data.
 However, we do not claim that the current model should do so yet, since we have chosen the simplest kind of summation with regards to procedural noise modes.

### 2017-02-05 ms.tex
Smallies

@@
 It is therefore possible to generate geometries at arbitrary scales.
-In addition, this method is extremely fast and low memory footprints, requiring very little computational power.
+In addition, this method is extremely fast and has low memory footprints, requiring very little computational power.
 After presenting such a model, we choose a method to measure the structural properties of nanoporous media that are created either through regular methods (e.g.
@@
 \includegraphics[width=.45\textwidth]{comparison.png}
-\caption{Right side: A realisation of the best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$. Left side: Simulated SiO$_2$ input data set used for parameter estimation. }
+\caption{Right side: A realisation of the best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$.
+Left side: Simulated SiO$_2$ input data set used for parameter estimation.
+Although the pore geometry seems to be similar, we see some artifacts, such as single loose atoms in the generated geometry.}
 \label{fig:porous_vs_model}
@@
 The porosity of the input SiO data set is $75\%$, while the best-fit data set has porosity $77\% \pm 3 \%$, depending on the random seed.
-In addition, the surface area of the original data set was $\sim 29 000 \AA^2$, while the best-fit data shows similar values with surface are ranging between $\sim 28 000 \AA^2$ and $\sim 30 000 \AA^2$.
+In addition, the surface area of the original data set was $\sim \SI{29000}{\angstrom^2}$, while the best-fit data shows similar values with surface are ranging between $\sim \SI{28000}{\angstrom^2}$ and $\sim \SI{30000}{\angstrom^2}$.
 \todo{legg ved 3D plot av likelihood for aa vise at den er kjekk og fin}

### 2017-02-05 ms.tex
thing
smaating

--- 2017-02-05 ms.tex
@@
 \begin{abstract} 
-In this paper, we aim to develop a method for producing statistically correct porous media using noise functions.
-In short, the proposed model is a binary function that returns either true if the input point is inside pore space, or false if within a wall.
+In this paper, we aim to develop a method for producing statistically correct simulations of nanoporous media without the use of conventional Molecular Dynamics Simulator (MDS).
+In short, the proposed model is a binary function that returns either true if the input position is in a void (pore), or false if within a wall.
 By using procedural noise, we are not restricted to generating media within a limited box, as the noise functions is continuous on all scales.
-It is therefore possible to generate geometries at arbitrary scales.
-In addition, this method is extremely fast and low memory footprints, requiring very little computational power.
-After presenting such a model, we choose a method to measure the structural properties of nanoporous media that are created either through regular methods (e.g.
-molecular dynamics) or with our method.
-The pair distribution function $g(r)$ is chosen as a measure of the structural properties to compare the simulated geometry to our model.
+It is therefore possible to simulate nanoporous media of arbitrary size.
+In addition, this method is extremely fast, requiring very little computational power to calculate whether a particle is within a pore or not.
+After presenting such a model, we select a method to measure the structural properties of nanoporous media that are created either through regular methods (e.g.
+LAMMPS) or with our method.
+The measure we selected is the radial distribution function (RDF) $g(r)$, which is regarded as the de-facto standard method for capturing the structural properties of porous media.
 Using this measure, we define a full likelihood framework for estimating model parameters given any data set, and continue by validating our own framework by analysing procedural noise data sets with known input parameters.
-Finally, we generate a nanoporous media with an expand-and-quench technique using molecular dynamics and show that our model correctly reproduces several of the internal properties of the data set, including $g(r)$, porosity and surface area.
+Finally, we perform a full likelihood analysis of silica SiO2 data set simulated in LAMMPS, and show that our model correctly reproduces several of the internal properties of the data set, including $g(r)$, porosity and surface area.
 \todo{Abstract boer tenkes noe gjennom: faa frem hovedpoenger.}
@@
 This means that it is possible to simulate
-\todo{Nevn finite size effects}
+
 %In addition, by creating a model that yields properties that are statistically consistent with observations, we are ar
@@
 We do not claim that procedural noise methods completely reproduces all structural properties of nanoporous silica, but show that our approach can be used to efficiently create materials that are structurally similar to simulations.
-Naturally, this all depends on the statistics used for comparing the two models, which in this case is based on the pair correlation function $g(r)$.
-\todo{Nevne at kjemi naturligvis ikke vil være riktig siden antall naboer per atom ikke er riktig. }
+Naturally, this all depends on the statistics used for comparing the two models, which in this case is based on the radial distribution function (RDF) $g(r)$.
+
 %\section{Method}
@@
 fractal-like patterns have been around for more than 30 years. The most famous of these
-algorithms is Perlin noise, first developed by Ken Perlin in 1983 for
-use in the movie ``Tron''\cite{perlin1985image}.
+algorithms is Perlin noise, first developed by Ken Perlin in 1982 for
+use in the movie ``Tron''.
 
 Perlin noise is a type of pseudo-random gradient noise that mimics a Gaussian field, but at low computational cost.
-An improved version of Perlin noise is called simplex noise, which has several advantages over standard Perlin noise in terms of scalability and speed \cite{perlin:2002}.
+An improved version of Perlin noise is called simplex noise, which has several advantages over standard Perlin noise in terms of scalability and speed.
 Perlin noise was developed in response to the lack of detail 1980's contemporary graphics, where polygonal surfaces often were bland and left without any detail.
@@
 An overview of the current state of procedural noise functions can be
-found in \cite{lagae:2010}. Perlin noise is a band limited repeatable pseudo-random function
+found in \cite{lagae:2010}.  Perlin noise \citep{perlin:2002} 
+is a band limited repeatable pseudo-random function
 from $\mathbb R^n \to \mathbb R$ that is computationally fast, and has
@@
 \end{itemize} 
-Simplex noise works in a similar manner, but works on on simplices instead of hypergrids, and so requires fewer computational steps, scales better in higher ($\ge 3$) dimensions and has fewer anisotropic properties.
+Simplex noise works in a similar manner, but requires fewer
+computational steps, scales better in higher ($\ge 3$) dimensions and has fewer 
+anisotropic properties.
 
-\subsection{Properties of Perlin/simplex noise}
+\subsection{Properties of Perlin/Simplex noise}
 \begin{figure}
@@
 We now perform a parameter fitting analysis on realistic simulated nanoporous media.
-Using LAMMPS \cite{plimpton1995fast}, we simulate a $100^3\AA$  box of 100K SiO2 silica particles at an initial temperature of $4500 K$, using the Vashishta-potential.
+Using LAMMPS \cite{plimpton1995fast}, we simulate a $100^3\AA$  box of 65K SiO2 silica particles at an initial temperature of $4500 K$, using the Vashishta-potential.
 After an initial period, we expand the box to $159^3 \AA$ and start cooling until quenching is achieved at about $300K$.

### 2017-02-06 ms.tex
Fixes

--- 2017-02-05 ms.tex
@@
 molecular dynamics) or with our method.
-The pair distribution function $g(r)$ is chosen as a measure of the structural properties to compare the simulated geometry to our model.
+The radial distribution function $g(r)$ is chosen as a measure of the structural properties to compare the simulated geometry to our model.
 Using this measure, we define a full likelihood framework for estimating model parameters given any data set, and continue by validating our own framework by analysing procedural noise data sets with known input parameters.
@@
 This is performed through a full likelihood analysis, where we in addition validate the model by comparing the radial distribution function $g(r)$, porosity and surface area of the system.
-The simulated system is confined to a cube of size $\SI{159}{\angstrom^{3}}$, but since the procedural noise function is continuous on all scales, the box could be of arbitrary size.
+The simulated system is confined to a cube of size $\SI{159} {\angstrom}^3$, but since the procedural noise function is continuous on all scales, the box could be of arbitrary size.
 
@@
 In this example, we only use standard simplex noise with three free parameters: the persistence $\psi$, threshold $\tau$ and scale $\nu$.
-We simulate a data set of $159 ^3 \AA$ of $\sim 255 000$ atoms, and perform the cut-off given input parameters $\Omega=3$, $\Psi = 0.6$, $\tau=-0.2$, $\nu=0.03$.
+We simulate a data set of $\SI{159} {\angstrom}^3$ of $\sim 255 000$ atoms, and perform the cut-off given input parameters $\Omega=3$, $\Psi = 0.6$, $\tau=-0.2$, $\nu=0.03$.
 The resulting 3D structure is depicted in figure \ref{fig:mockdata}.
@@
 \includegraphics[width=.99\textwidth]{results_porous_full.png}
-\caption{Results from the analysis of a simulated $159^3$ \AA SiO$_2$ nanoporous data set.
+\caption{
+Results from the analysis of a simulated $\SI{159} {\angstrom}^3$ SiO$_2$ nanoporous data set.
 The upper panel displays the marginalised 1D posteriors for each parameter, while the lower panel depicts the marginalised 2D posteriors.
-Best fit parameters from the maximum 3D likelihood suggest $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$.}
+Best fit parameters from the maximum 3D likelihood suggest $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$.
+}
+
 \label{fig:porous_results1}
@@
 
-The porosity of the input SiO data set is $75\%$, while the best-fit data set has porosity $77\% \pm 3 \%$, depending on the random seed.
+The porosity of the input SiO$_2$ data set is $75\%$, while the best-fit data set has porosity $77\% \pm 3 \%$, depending on the random seed.
 In addition, the surface area of the original data set was $\sim \SI{29000}{\angstrom^2}$, while the best-fit data shows similar values with surface are ranging between $\sim \SI{28000}{\angstrom^2}$ and $\sim \SI{30000}{\angstrom^2}$.
@@
 We have shown that it is possible through a full likelihood analysis to reproduce the input parameters of a simulated model.
-In addition, we performed an analysis on a simulated $159^3 \AA$ data set of SiO$_2$, and explain that while porosity and surface area corresponds well between the model and data, the $g(r)$ shows that the current model does not exactly reproduce correct data.
+In addition, we performed an analysis on a simulated $\SI{159}{\angstrom^{3}}$ data set of SiO$_2$, and explain that while porosity and surface area corresponds well between the model and data, the $g(r)$ shows that the current model does not exactly reproduce correct data.
 However, we do not claim that the current model should do so yet, since we have chosen the simplest kind of summation with regards to procedural noise modes.

### 2017-02-06 ms.tex
Renskrevet mengder, endel nytt

@@
 \begin{abstract} 
-In this paper, we aim to develop a method for producing statistically correct porous media using noise functions.
+In this paper, we aim to develop a method for producing statistically correct porous media using procedural noise functions.
 In short, the proposed model is a binary function that returns either true if the input point is inside pore space, or false if within a wall.
 By using procedural noise, we are not restricted to generating media within a limited box, as the noise functions is continuous on all scales.
-It is therefore possible to generate geometries at arbitrary scales.
+It is therefore possible to generate fractal-like geometries at arbitrary scales, circumventing the need for any physics simulations.
 In addition, this method is extremely fast and has low memory footprints, requiring very little computational power.
@@
 Model simulations typically require a large computational framework to accurately reproduce the observed effects, while experimental data can be both expensive and hard to obtain, in addition to being riddled with artefacts and instrumental errors.
+While physical simulations yields statistically correct results, it is often hard to re-create features that are noticeable in experimental data, and therefore fit model parameters with observations.
+In this paper, we propose a new method for producing realistic-looking nanoporous media using only random noise fields, without the need for physics simulations.
+We do not claim that our method produces results that are more physically accurate than full model simulations, quite the contrary, we aim to show that it is possible to create fast, near-accurate and arbitrary large data sets of nanoporous media that are statistically similar to physics-based simulations with only the use of random noise fields.
 
@@
 On the other hand, using procedural noise to carve out nanoporous media is both extremely fast and computationally cost-efficient, in addition to being continuous on all scales.
-This means that it is possible to simulate
+
 \todo{Nevn finite size effects}
@@
 
-In this paper, we aim to investigate whether procedural methods are suitable for producing statistically correct 3D physical structures, focusing on nanoporous media.
+Since we aim to investigate whether procedural methods are suitable for producing statistically correct 3D physical structures, we choose to focus our attention on silica-based nanoporous media.
 In a sense, we are combining well-known methods from computer games and real-time graphics-oriented software with experimental observations and simulations.
@@
 By procedural content generation, we generally mean a mathematical algorithm that uses limited input data to create complex, sometimes infinite repeating patterns.
-A firm example would be fractals, such as the Julaia/Mandelbrot [ref mandelbrot] set, where one tests whether the complex quadratic recurrence equation $z_{n+1} = z_n^2 +C$ for $z \in \mathbb C$ converges for each position in a 2D complex plane.
+A firm example would be fractals, such as the Julia/Mandelbrot [ref Mandelbrot] set, where one tests whether the complex quadratic recurrence equation $z_{n+1} = z_n^2 +C$ for $z \in \mathbb C$ converges for each position in a 2D complex plane.
 With this seemingly simple equation, the level of complexity generated on the boundary between convergence and divergence is infinite, with self-similar patterns repeating themselves to arbitrary detail.
@@
 
-Another example, widely used in contemporary game content generation, is having prefabs (pre-fabricated 3D models such as doors, walls, tiles and roof segments) automatically being stitched together to create buildings.
+Another example, widely used in contemporary game content generation, is having prefabs, or pre-fabricated 3D models such as doors, walls, tiles and roof segments automatically being stitched together to create buildings.
 When designing a city, instead of creating each building manually, you only need specify the position, area and height of a structure.
@@
 
-A much simpler example would be generating a 1 (or 2)-dimensional landscape by choosing random points on a grid and performing a spline interpolation between them to create the illusion of a smooth landscape.
-In fact, this is the very essence of procedural noise, which we are focusing on in this paper, and the first person to create such a fast algorithm for N-dimensional spline interpolation between pseudo-random points was Ken Perlin.
+A much simpler example would be generating a 1 (or 2)-dimensional terrain by choosing random points on a grid and performing a spline interpolation between them to create the illusion of a smooth landscape.
+In fact, this is the very essence of procedural noise, which we are focusing on in our model, and the first person to create such a fast algorithm for N-dimensional spline interpolation between pseudo-random points was Ken Perlin.
 
@@
 a single octave. The upper-right part shows the same node for simplex noise,
-with less obvious patterns. However, when combining several octaves and adding a random shift to the coordinates,
-as depicted in the two lower parts, these anisotropic patterns tend to cancel
+with less obvious ``textile'' features.
+However, when combining several octaves and adding a random shift to the coordinates, as depicted in the two lower parts, these anisotropic patterns tend to cancel
 out.  In addition, we have calculated the statistical distributions for Perlin and simplex noise
-with $\sigma = 0.3$, and have plotted the results in figure \ref{fig:properties}.
+with $\sigma = 0.35$, and have plotted the results in figure \ref{fig:properties}.
 Each procedural noise method has distributions that are close to normal, 
@@
 \label{eq:procedural}
- \Phi(\vec x) = \Theta \Big(\sum_k^\omega \kappa (k) P(  f_s (\vec x,k) )) \Big),
+ \Phi(\vec x) = \Theta \Big(\sum_k^\Omega \kappa (k) P(  f_s (\vec x,k) )) \Big),
  \end{equation}
@@
 \label{eq:perlinlinear}
- \Phi(\vec x) = \sum_k^\omega \frac{1}{k} P( 2k\vec x).
+ \Phi(\vec x) = \sum_k^\Omega \frac{1}{k} P( 2k\vec x).
 \end{equation}
-In 1 dimension, equation \ref{eq:perlinlinear} with 1, 3 and 30 octaves is depicted in figure \ref{fig:1dperlin}, while the 2D procedural noise pattern depicted in the lower
+In one dimension, we show equation \ref{eq:perlinlinear} with 1, 3 and 30 octaves in figure \ref{fig:1dperlin}, while the 2D procedural noise pattern depicted in the lower
 part of figure \ref{fig:noise}.
-Finally, since simplex noise both scales better with higher dimensions and produces less asymmetric effects, we choose to use this noise method over regular Perlin noise in our method.
+Finally, since simplex noise both scales better with higher dimensions and produces less asymmetric effects, we choose to use simplex over regular Perlin noise as the main noise function throughout this paper.
 
@@
 \section{Model}
-In this paper, we are interested in developing a model using simplex noise to reproduce nanoporous structures that are as statistically close to actual simulations of porous media.
+In this paper, we are interested in developing a model using simplex noise to reproduce nanoporous structures that are as statistically close to physical simulations of porous media.
 For a given data set and a model, we wish to estimate the parameters that best represent the data.
-This is performed through a full likelihood analysis, where we in addition validate the model by comparing the radial distribution function $g(r)$, porosity and surface area of the system.
+This is performed through a full likelihood analysis in order to obtain the posterior probability distribution of the parameters.
+The model measure of the analysis is the radial distribution function $g(r)$, for which the parameters will be maximised to fit.
+In addition, we also investigate the porosity and surface area of the system.
 The simulated system is confined to a cube of size $\SI{159} {\angstrom}^3$, but since the procedural noise function is continuous on all scales, the box could be of arbitrary size.
@@
 For each mode, the frequency is thus multiplied, while the overall amplitude is damped by a given persistence.
-In the end, we simulate a large set of bulk particles without any walls, and use the noise function to remove particles that are not in a procedural void given by a threshold $\tau$.
+In the end, we simulate a large set of bulk particles without any walls, and use the noise function to remove particles that are not within a procedural void given by a threshold $\tau$.
 Three examples of this model are depicted in figure \ref{fig:model_example}, where we have used $\Omega=6$ octaves and the scale is $\nu=0.01$ and cut-off $\tau=0$.
@@
 The middle picture shows the same parameters with a higher persistence $(\psi = 0.5)$, where smaller scales become more dominating.
-In the rightmost picture, the effect of the cut-off threshold $\tau = -0.3$ is shown.
+In the rightmost picture, the effect of the cut-off threshold $\tau = -0.3$ is shown to emphasize threshold cut-off.
 
 \subsection{Model Measure}
-In order to perform data analysis on the model parameters, we developed a code that uses Monte Carlo Markov chains in order to produce samples for building the likelihood.
+In order to perform data analysis on the model parameters, we developed a code that uses Monte Carlo Markov chains to produce samples for building the posterior probability distribution for the parameters.
 Before doing so, we need a model measure - a quantitative method that characterizes the internal structure of the porous media.
@@
 $g(r)$ describes how density varies as a function of distance from a reference particle, picking up both the intramolecular and intermolecular structure.
-We have chosen to ignore the short range contributions of $g(r)$ since the pore geometry does not depend on the intramolecular structure of SiO$_2$, but only the density variations at larger distances.
-Hereafter, $g_d(r)$ denotes the RDF calculated from simulated or experimental data, while $g_{M(\bar \theta)}(r)$ represents the RDF based on the procedural noise model $M(\bar \theta)$, where $\bar \theta$ are the set of model parameters.
+We have chosen to ignore the short range contributions of $g(r)$ since the pore geometry does not depend on the intra-molecular structure of SiO$_2$, but only the density variations at larger distances.
+Hereafter, $g_d(r)$ denotes the RDF calculated from simulated or experimental data, while $g_{M(\bar \theta)}(r)$ represents the RDF based on the procedural noise model $M(\bar \theta)$, where $\bar \theta$ is the set of model parameters.
 
 The method used for calculating porosity of a media is defined as in \cite{gelb1998characterization}, and we obtain the surface area of a porous media by using marching cubes triangulation to sum the surface area of the generated triangles.
-As discussed in \cite{gelb1998characterization}, the definition of surface area is ill-defined in atomic simulations, because a definition of pore space is needed.
-In the marching cubes triangulation, we generate a voxelized grid, the scalar field, with values being the distance from the voxel center to the nearest atom.
-The isovalue threshold defining the surface is chosen to be \SI{2}{\angstrom}, large nough to have a dense pore matrix, but small enough to have most of the available pore space defined as such.
-The important note here is that the same definition of surface is being used in all compared systems.
+As discussed in \cite{gelb1998characterization}, the definition of surface area is ill-defined in atomic simulations, because a clear definition of pore space is needed.
+In the marching cubes triangulation, we create a voxelized grid, a scalar field with values being the distance from the voxel centre to the nearest atom.
+The isovalue threshold defining the surface is chosen to be \SI{2}{\angstrom}, large enough to contribute to a dense pore matrix, but small enough to have most of the available pore space defined as such.
+The important note here is that the same definition of both surface and porosity is being used in all systems that are compared, physical simulations, procedural methods or data.
 
@@
 
-The posterior probability is defined as the probability of the parameters $\bar \theta$ given the evidence $d$, or $P(\bar \theta | d)$. Bayes' theorem states that 
+The posterior probability distribution is defined as the probability of the parameters $\bar \theta$ given the evidence $d$, or $P(\bar \theta | d)$. Bayes' theorem states that 
 \begin{equation}
@@
 $P(d)$ is called the evidence, and is the probability of the observations (here assumed to be 1).
-Since the prior is a constant, we can therefore assume that the posterior distribution is proportional to the likelihood, or $P(d | \bar \theta) \propto \mathcal L$.
-In other words, to calculate the posterior and obtaining the best fit model parameters, we need only to estimate the likelihood function, which is connected to the $\chi^2$ by
+Prior parameters are restricted to freely roam within a minimum and maximum value, so we can therefore assume that the posterior distribution is proportional to the likelihood, or $P(d | \bar \theta) \propto \mathcal L$.
+In other words, to calculate the posterior probability distribution and obtaining the best fit model parameters, we need only to estimate the likelihood function, which is connected to the $\chi^2$ by
 \begin{equation}
@@
 
-In order to test whether two 3D data sets are statistically equivalent up to the model measure $g(r)$, we calculate the $\chi^2$ between the two one-dimensional RDFs. Pearson's $\chi^2$ is defined as 
+In order to estimate whether two 3D data sets are statistically equivalent up to the model measure $g(r)$, we calculate the $\chi^2$ between the two one-dimensional RDFs.
+Pearson's $\chi^2$ is defined as
 \begin{equation}
@@
 \end{equation}
-so for low values of $\chi^2$, the better the fit between data and model (up to $g(r)$).
-For a single-parameter model, we could simply just iterate over the lone parameter to obtain the minimum $\chi^2$ (where the best-fit parameter value is), however, with increased number of parameters comes increased computational cost.
+so for low values of $\chi^2$, the better the fit between data and model, up to $g(r)$.
+For a single-parameter model, we could simply just iterate over the lone parameter to obtain the minimum $\chi^2$ (where the best-fit parameter value resides), however, with increased number of parameters comes increased computational cost.
 If we divide the parameter space into $N$ bins, the number of $\chi^2$ computations scales as $N^d$, where $d$ is the number of model parameters.
-To be able to build the full $\chi^2$ (and likelihood) parameter landscape, we there need to use Monte Carlo Markov chain methods (MCMC).
+To be able to build the full landscape of the posterior probability distribution, we need to use the Monte Carlo Markov chain (MCMC) method .
 
 We are now interested in maximising the likelihood $\mathcal L$ as opposed to minimising the $\chi^2$, with the goal of mapping out the full $d$-dimensional likelihood space for the model parameters.
-We build the $d$-dimensional likelihood by letting random walkers traverse the parameter space guided by the MCMC test: for each step $M_n(\bar \theta)$ in parameter space, we calculate the likelihood for this parameter configuration.
-Then, we create a new set of parameters for the next step $M_{n+1}(\bar \theta)$ by choosing new parameters randomly from a $d$-dimensional sphere around the previous parameter composition.
+We build the $d$-dimensional likelihood by letting random walkers traverse the parameter space guided by the MCMC test: for each step $M(\bar \theta_{n})$ in parameter space, we calculate the likelihood $\mathcal L_n$ for this particular parameter configuration.
+Then, for the next step, we create a new set of parameters $M(\bar \theta_{n+1})$ by choosing new parameters randomly from a $d$-dimensional sphere around the previous parameter composition.
 The model also selects a new random seed.
-We calculate the likelihood at the new step and perform the MCMC test: accept the new step if $\frac{\mathcal L_n}{\mathcal L_n+1} \ge 1$, or reject if the value is lower than a uniform random value $\in [0,1]$.
+We calculate the proposed likelihood $\mathcal L_{n+1}$ at this new step and perform the MCMC test: accept the new step if $\frac{\mathcal L_n}{\mathcal L_n+1} \ge 1$, or reject if the value is lower than a uniform random value $\in [0,1]$.
 
@@
 First, we wish to test and verify the code by performing a full data analysis on mock data with known input parameters.
-This enables us to investigate the posterior distribution, and determine whether any potential problems could arise with the likelihood landscape, such as radom walkers trapped in local minima.
+This enables us to investigate the posterior distribution, and determine whether any potential problems could arise with the likelihood landscape, such as random walkers trapped in local minima.
+In addition, we will obtain an intuition on the width of the estimated parameter distributions.
+But most importantly, we will have verified that our likelihood code is working correctly if we successfully reproduce the original input parameters.
 We then continue with a full analysis of simulated SiO$_2$ data, and see whether the resulting best-fit parameter data set is consistent with the SiO$_2$ data set.
@@
 \end{figure}
-In order to verify our likelihood code, we perform a model verification test on a mock data set with known input values.
+In order to verify our likelihood code, we perform a model verification test on a mock data set with known input values based on equation \ref{eq:noisemodel1}.
 In this example, we only use standard simplex noise with three free parameters: the persistence $\psi$, threshold $\tau$ and scale $\nu$.
-We simulate a data set of $\SI{159} {\angstrom}^3$ of $\sim 255 000$ atoms, and perform the cut-off given input parameters $\Omega=3$, $\Psi = 0.6$, $\tau=-0.2$, $\nu=0.03$.
+We simulate a data set of size $\SI{159} {\angstrom}^3$ with $\sim 255 000$ atoms, and perform the cut-off given by input parameters $\Omega=3$, $\Psi = 0.6$, $\tau=-0.2$ and $\nu=0.03$.
 The resulting 3D structure is depicted in figure \ref{fig:mockdata}.
 
-We continue by performing a full MCMC analysis of the mock data with using a parallelized code running on a 48 core cluster.
-In order to produce 300 000 samples, about 48 hours was needed.
+We continue by executing a full MCMC analysis on the mock data by using a parallelized code written in C++ running on a 48 core cluster.
+About 48 hours was needed to produce about 300 000 samples.
 The results from this analysis can be seen in figure \ref{fig:mockdataresults_pts}, where the upper panel depicts the marginalised 1D posteriors while the lower shows the marginalised 2D posteriors.
 The centre and left image shows that the input parameters of $\tau=-0.2$(threshold), $\psi=0.6$(persistence) and $\nu=0.03$ are all reproduced well within $1 \sigma$.
-
 Note that there seems to be no problems with the posterior distribution in terms of local minima or multiple maxima.
+In addition, the results show that all the parameters are correlated with $r \sim 0.85$, while the width of the marginalised distributions vary by $\sigma_\tau = 0.02 \sim 10\%$, $\sigma_\Psi = 0.062 \sim 20\%$ and $\sigma_\nu = 0.01 \sim 33\%$.
 
@@
 The upper panel displays the marginalised 1D posteriors for each parameter, while the lower panel depicts the marginalised 2D posteriors.
-Note that all input parameters were reproduced well within $1\sigma$ (inner contour line).}
+Note that all input parameters were reproduced well within $1 \sigma$ (inner contour line).}
 \label{fig:mockdataresults_pts}
@@
 \end{figure*}
-
-We now perform a parameter fitting analysis on realistic simulated nanoporous media.
+Having verified that the likelihood code yields correct results when analysing mock data, we turn to 
+performing a parameter fitting analysis on realistic simulated nanoporous media.
 Using LAMMPS \cite{plimpton1995fast} and the Vashishta-potential \cite{vashishta1990interaction}, we simulate system consisting of $14^3$ betacristobalite unit cells of SiO$_2$, corresponding to system size of approximately \SI{100}{\angstrom}.
@@
 Then, the system is quenched by setting the temperature to \SI{300}{\kelvin}, resulting in the nanoporous media on the right side in figure \ref{fig:porous_vs_model}.
-By choosing a different seed for the initial velocities, it is easy to produce other samples.
+By choosing a different seed for the initial velocities, it is possible to produce statistically equal samples but with random realizations.
 
 With the model described in equation \ref{eq:noisemodel1}, we perform a full likelihood analysis of the data set with the three free parameters $\tau$, $\nu$ and $\psi$ using $g(r)$ as the measure.
-The resulting marginalised posteriors are depicted in figure \ref{fig:porous_results1}, with best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$ is extracted from the full 3D likelihood.
+The resulting marginalised posteriors are depicted in figure \ref{fig:porous_results1}, where the best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$ is extracted from the full 3D likelihood.
 We also display $g(r)$ for the input data set, a best-fit parameter data set and a bad-fit data set in figure \ref{fig:gofr1}, to illustrate the sensitivity of $g(r)$ given the model parameters.
 Note that we are not able to accurately reproduce the structures of the SiO$_2$ $g(r)$ with our current model, but hope that we during future work will be able to produce another model that provides a better fit with this measure.
+\todo{skriv om hvorfor vi ikke plukker opp alle features?}
 
-Note that the surface atoms are not in chemical equilibrium since some neighbors of a surface atom might have been removed.
+Note that the surface atoms are not in chemical equilibrium since some neighbours of a surface atom might have been removed.
 In order to use the generated geometry in molecular dynamics, one would need either noble gas atoms which may be stable, or apply passivation step.
@@
 \subsection{Some notes on the results}
-It is worth to note that since we are chiefly interested in obtaining a set of parameters that reproduces a statistical correct representation of the input data, we are relatively free to choose parameters as desired.
+It is worth to note that since we are chiefly interested in obtaining a set of parameters that reproduces a statistical correct representation of the input data, we are relatively free to choose parameters as desired within the posterior distribution width of the parameter.
 In the previous analysis, we selected a best fit threshold value of $\tau=-0.2$, but from the width of the likelihood it is clear that nearby parameters would also be suitable.
@@
 $\Delta \tau = 0.05$) yields a relatively large change in surface area.
-
-In addition, we do not claim that this model accurately reproduces the pore structure observed in nanoporous simulations, as is evident from the RDF shown in figure \ref{fig:gofr1}.
+In addition, we do not claim that this model completely reproduces the pore structure observed in nanoporous simulations, as is evident from the RDF shown in figure \ref{fig:gofr1}, but hope to discover a better model that will produce a $g(r)$ that is more in accordance with the data.
 
@@
 
-%persistence : 0.875111019125
-%scale : 0.036638
-%threshold : -0.236784875   78% porosity
-
-% bad fit : scale 0.1, p = 1, threshold = -0.3
-
 \section{Conclusion}
@@
 In this paper, we have proposed a novel method for generating a simple kind of nanoporous media using procedural methods, namely simplex noise.
-We have developed tools for creating such porous materials, and also estimating model parameters given data.
-We have shown that it is possible through a full likelihood analysis to reproduce the input parameters of a simulated model.
-In addition, we performed an analysis on a simulated $\SI{159}{\angstrom^{3}}$ data set of SiO$_2$, and explain that while porosity and surface area corresponds well between the model and data, the $g(r)$ shows that the current model does not exactly reproduce correct data.
+Procedural methods, albeit having no root in physics, have some advatages over prevalent methods: it computationally efficient, has low memory footprints and is only limited in terms of detail by the model in question.
+With detail, we mean that is possible to create simulations of arbitrary scales and resolution.
+We have also developed a framework for both producing such materials and also estimating model parameters given data.
+We have shown that it is possible through a full likelihood analysis to successfully reproduce the input parameters of procedurally simulated data.
+In addition, we performed an analysis on a simulated $\SI{159}{\angstrom^{3}}$ data set of SiO$_2$, and explain that while porosity and surface area corresponds well between the model and data, the $g(r)$ shows that the current model does not completely reproduce correct data.
 However, we do not claim that the current model should do so yet, since we have chosen the simplest kind of summation with regards to procedural noise modes.
@@
 \subsection{Future work}
-In this paper, we have only considered the simplest kind of procedural noise: a simple sum over modes with threshold cut-off $\tau$, scale fall-off persistence $\psi$ and initial scale $\nu$.
+We have only considered the simplest kind of procedural noise: a simple sum over modes with threshold cut-off $\tau$, scale fall-off persistence $\psi$ and initial scale $\nu$.
 We have also seen that, even through both porosity and surface area accurately reproduced, there are features in the RDF that are not fully captured with the current model.
@@
 Another immediate application would be to estimate model parameters given experimental observations of porous media.
-This would enable us to estimate new parameters such as instrumental noise, deviations from symmetry (yielding asymmetric properties) and pureness of samples.
-In addition, by analysing just a few samples and producing an accurate model, we are able to produce an arbitrary amount of statistically correct samples, free from instrumental noise and artefacts.
+This would enable us to estimate new parameters such as instrumental noise, deviations from symmetry (asymmetric properties), and pureness of samples.
+In addition, by analysing just a few samples and developing a model that produces similar structures as in the specimens, we are able to produce an arbitrary amount of statistically correct samples, free from instrumental noise and artefacts.
 

### 2017-03-24 ms.tex
Structure of intro, still needs to be written
Rewrote introduction. Still missing references. Some changes to conclusions - revision of conclusion needed.

@@
 
-%During recent years, increase in computational power has allowed for a vast improvement  
-\todo{Skriv en fin intro! Ogsaa om nanoporous media}
-Properties of structural patterns that regularly occur in nature are often investigated either through direct observations or through simulations of theoretical models.
-Model simulations typically require a large computational framework to accurately reproduce the observed effects, while experimental data can be both expensive and hard to obtain, in addition to being riddled with artefacts and instrumental errors.
-While physical simulations yields statistically correct results, it is often hard to re-create features that are noticeable in experimental data, and therefore fit model parameters with observations.
+The rapid development of new experimental and observational methods to determine the three-dimensional structure of many types of porous media calls for new statistical tools and methods to characterize complex porous media and to generate synthetic media based on the observed structures.
+New experimental methods allow for studies of three-dimensional structures from nano-meter to meter scales.
+Destructive techniques such as focused ion beam scanning electron microscopy (FIB-SEM) are routinely used to characterize geologically relevant structures in shales and carbonate rocks with resolutions down to tens of nanometers \cite{XX}.
+High-energy X-ray tomography methods using high-energy beams at international research facilites or in-house scanners open for three-dimensional mapping of mineral structures from sub-micrometer to millimeter scales \cite{XX}.
+Thus new experimental techniques provide three-dimensional voxelized images that can form the basis of analysis or modeling.
+However, in many cases we can only scan a few examples experimentally, but need to perform simulations over many statistically similar realizations to find effective properties or to populate upscaling models.
+Thus the need for methods to robustly characterize three-dimensional porous structures and to generate synthetic structures with similar statistical properties.
+
+The field of synthetic rocks is a large field both scientifically and technologically \cite{}.
+Methods to generate synthentic rock samples from experimental characterizations may be based on purely statistical or geometrical models, models of physical processes that generate the samples, or combinations of both.
+For small-scale porous media, physical models are often used to generate initial grain packings \cite{}, then further mechano-chemical and fluid-mediated processes such as precipitation-dissoluton processes and pressure solution \cite{}, or fracture and plastic deformation processes are added to include effects of long-term compaction \cite{}.
+On large scales, reservoir models are populated by statistical models for lithological and geomorphological processes \cite{}, such as river fan formation to generate particular grain size distributions \cite{}.
+More statistical approaches include variations of percolation models \cite{sahimi} or network models \cite{} to characterize and generate uncorrelated or correlated porous media.
+In addition, the high resolution of observational methods allow three-dimensional scans to form the basis for atomic-scale modeling using molecular modeling methods such as molecular dynamics or Monte Carlo methods to study the dynamics of nanoporous media.
+But then procedural or physical models are needed to fill in the atomic scale details, to build many statistical realizations of nanoporous media with atomic-scale details, and to extrapolate to larger scale models.
+
+These physically based models may not always include all relevant effects, can be quite computationally expensive when used to generate many samples, and cannot easily be extended beyond the initial size of the simulated systems.
+For example, to prepare a simulation of nanoporous silica, we need to carry out a sequence of large scale simulations in molecular dynamics, which is time consuming, computationally expensive and the results canno be exteded beyond the intial size of the simulated systems without performing another simulation \cite{Adarsh}.
+On the other hand, the statistical models often do not include all the correlations induced by physical processes or have limited number of tunable parameters.
+In particular, statistical models have challenges in reproducing atomic-scale structures of nanoporous media.
+
+Meanwhile, there has been a recent growth in advanced random noise models that open for new ways to generate structured noise with statistical similarities to real samples, such as the use of Perlin noise.
+Such noise models have been used to model systems with complex statistical correlations such as in (examples from Nicloaas galaxy/astrophysics applications), and show great promise in being adaptable to many different physical systems.
+Also, such noise models are continuous and extendable, so that arbitrary large systems or systems with arbitrary high precision can be modeled.
+While these models show great promise in many fields \cite{}, they have not yet been systematically applied to characterize or generate nanoporous media with statistical properties corresponding to that of real systems.
+
 In this paper, we propose a new method for producing realistic-looking nanoporous media using only random noise fields, without the need for physics simulations.
-We do not claim that our method produces results that are more physically accurate than full model simulations, quite the contrary, we aim to show that it is possible to create fast, near-accurate and arbitrary large data sets of nanoporous media that are statistically similar to physics-based simulations with only the use of random noise fields.
+We provide a systematic study of the use of procedural noise-based models to characterize and generate three-dimensional nanoporous media.
+Our studies will be based on silica-based nanoporous media, for which we have good atomic scale models and physics-based methods to generate porous media as a sequence of simulated processes, and we therefore can generate large-scale samples where we have full knowledge of all aspects of the material.
+These simulated systems are then characterized using the radial distribution function, $g(r)$, statistical methods are generated with similar properties, and compared with the simulated systems.
+Using these approaches we find that procedural noise methods can be tuned to efficiently create materials that are structurally similar to simulationed materials.
 
-Procedural generation of physical structures have several advantages over conventional simulations.
-For example, when producing simulations of nanoporous silica, we are required to calculate and propagate the physical properties of each particle for every time-step throughout the simulation, using a computational framework such as LAMMPS \cite{plimpton1995fast}.
-This is both time-consuming, and the computational cost can sometimes be be quite large.
-In addition, the simulation is constrained to a box, so particles are not allowed to move outside the defined boundaries.
-On the other hand, using procedural noise to carve out nanoporous media is both extremely fast and computationally cost-efficient, in addition to being continuous on all scales.
-
-\todo{Nevn finite size effects}
-%In addition, by creating a model that yields properties that are statistically consistent with observations, we are ar
-
-Since we aim to investigate whether procedural methods are suitable for producing statistically correct 3D physical structures, we choose to focus our attention on silica-based nanoporous media.
-In a sense, we are combining well-known methods from computer games and real-time graphics-oriented software with experimental observations and simulations.
-Procedural methods for automated generation of computer game content have been around since the early 1980's, and in those early days algorithms were less advanced, often yielding plain, simple geometrical structures.
-Nowadays, almost every single computer game produced will contain some level of automated generated content, from textures to level design to landscapes and gameplay.
-
-We do not claim that procedural noise methods completely reproduces all structural properties of nanoporous silica, but show that our approach can be used to efficiently create materials that are structurally similar to simulations.
-Naturally, this all depends on the statistics used for comparing the two models, which in this case is based on the radial distribution function $g(r)$ (RDF).
-\todo{Nevne at kjemi naturligvis ikke vil være riktig siden antall naboer per atom ikke er riktig. }
-%\section{Method}
 \section{Procedural content generation}
@@
 
-\subsection{Some notes on the results}
+\subsection{Discussion}
 It is worth to note that since we are chiefly interested in obtaining a set of parameters that reproduces a statistical correct representation of the input data, we are relatively free to choose parameters as desired within the posterior distribution width of the parameter.
@@
 \todo{konklusjon maa renskrives}
-In this paper, we have proposed a novel method for generating a simple kind of nanoporous media using procedural methods, namely simplex noise.
-Procedural methods, albeit having no root in physics, have some advatages over prevalent methods: it computationally efficient, has low memory footprints and is only limited in terms of detail by the model in question.
-With detail, we mean that is possible to create simulations of arbitrary scales and resolution.
-We have also developed a framework for both producing such materials and also estimating model parameters given data.
+
+In this paper, we have proposed a novel method for generating a nanoporous media using procedural methods in the form of simplex noise.
+Procedural methods, albeit having no root in physics, have some advatages over prevalent methods: they are computationally efficient, have low memory footprints and are only limited in terms of detail by the model in question.
+In particular, it is possible to create simulations of arbitrary scales and resolution, indeed with any level of detail.
+
+\todo{AMS: This is a bit weak currently. State first what you have succeeded in doing. Then state the possible weaknesses of the model.}
+We have developed a framework to generate nanoporous media using molecular dynamics simulations and to estimate model parameters given such data.
 We have shown that it is possible through a full likelihood analysis to successfully reproduce the input parameters of procedurally simulated data.
-In addition, we performed an analysis on a simulated $\SI{159}{\angstrom^{3}}$ data set of SiO$_2$, and explain that while porosity and surface area corresponds well between the model and data, the $g(r)$ shows that the current model does not completely reproduce correct data.
-However, we do not claim that the current model should do so yet, since we have chosen the simplest kind of summation with regards to procedural noise modes.
+In addition, we performed an analysis on a simulated $\SI{159}{\angstrom^{3}}$ data set of SiO$_2$, and explain that while porosity and surface area correspond well between the model and data, analysis of $g(r)$ shows that the current model reproduces the main features of the material, but that there are some deviations.
+We find the correspondence strong, which demonstrates the proof-of-principle of the use of procedural noise methods, but we realize that even better correspondence may be achieved by including other summations in the procedural noise modes.
 

### 2017-05-16 ms.tex
Fixed references, added newlines to all sentences for simpler git history

@@
 Meanwhile, there has been a recent growth in advanced random noise models that open for new ways to generate structured noise with statistical similarities to real samples, such as the use of Perlin noise.
-Such noise models have been used to model systems with complex statistical correlations such as in (examples from Nicloaas galaxy/astrophysics applications), and show great promise in being adaptable to many different physical systems.
+Such noise models have been used to model systems with complex statistical correlations such as in GAMER\cite{groeneboom:2014}, and show great promise in being adaptable to many different physical systems.
 Also, such noise models are continuous and extendable, so that arbitrary large systems or systems with arbitrary high precision can be modeled. 
@@
 With this seemingly simple equation, the level of complexity generated on the boundary between convergence and divergence is infinite, with self-similar patterns repeating themselves to arbitrary detail.
-\begin{figure}
-\includegraphics[width=.5\textwidth]{orphancity.png}
-\caption{Example of city building using procedural methods: Landscape, city infrastructure, buildings and all textures are all automatically generated using procedural methods.
-Created by one of the authors.
-\todo{Denne er jo med bare for moro skyld, boer jo fjernes} }
-\label{fig:orphancity}
-\end{figure}
 
@@
 Ideally, the model measure should pick up all geometrical properties that are relevant for the physical study of interest. 
-For instance, transport in porous media heavily depends on the pore size distribution and porosity, whereas chemical reactons are sensitive to surface area.
-\todo{referanser, malthe?} In this paper, we choose to focus on the radial distribution function $g(r)$ as defined and implemented in LAMMPS \cite{plimpton1995fast}.
+For instance, transport in porous media heavily depends on the pore size distribution and porosity, whereas chemical reactons are sensitive to surface area. \todo{referanser, malthe?} 
+In this paper, we choose to focus on the radial distribution function $g(r)$ as defined and implemented in LAMMPS \cite{plimpton1995fast}.
 $g(r)$ describes how density varies as a function of distance from a reference particle, picking up both the intramolecular and intermolecular structure.
@@
 In the previous analysis, we selected a best fit threshold value of $\tau=-0.2$, but from the width of the likelihood it is clear that nearby parameters would also be suitable. 
-However, the surface area is strongly correlated with this value, so even slight deviations (e.g.
-$\Delta \tau = 0.05$) yields a relatively large change in surface area.
+However, the surface area is strongly correlated with this value, so even slight deviations (e.g. $\Delta \tau = 0.05$) yields a relatively large change in surface area. 
 In addition, we do not claim that this model completely reproduces the pore structure observed in nanoporous simulations, as is evident from the RDF shown in figure \ref{fig:gofr1}, but hope to discover a better model that will produce a $g(r)$ that is more in accordance with the data.
@@
 In terms of the data, the variation is attributed as the initial velocities of the particles in the LAMMPS simulation. 
-For the model, it corresponds to the random seed.
-We found that varying these initial conditions have negligible effect on the calculation of $g(r)$.
+For the model, it corresponds to the random seed. We found that varying these initial conditions have negligible effect on the calculation of $g(r)$.
 
@@
 
-\todo{Kanskje fake-latin burde fjernes? Usikker!}
-Ut hendrerit cursus libero, in scelerisque tortor sollicitudin accumsan.
-Nam eleifend velit metus, quis volutpat justo faucibus ut.
-Cras ut sem in nunc fringilla egestas vitae sit amet justo.
-Ut vel condimentum tortor.
-Morbi in massa in ipsum ultrices tincidunt.
-Duis feugiat dignissim nisi ac ultricies.
-Quisque eget risus fermentum, condimentum massa ullamcorper, posuere neque.
-Nam congue consequat mi, vel commodo mi tempus eget.
-
-Nullam ultrices, velit sed venenatis ultrices, risus urna ultricies magna, at pretium lorem nibh et nunc.
-Suspendisse sed mattis leo.
-Nulla facilisi.
-Praesent finibus lobortis leo, ut porttitor massa tincidunt ut.
-Vivamus quis orci eu sem sodales ornare.
-Cras sed faucibus velit.
-Phasellus scelerisque mauris quis lorem dignissim vehicula.
-Sed tempor ut urna sed bibendum.
-Donec et nibh pulvinar, commodo magna ac, interdum nibh.
-Sed laoreet ipsum quis quam lacinia, in dictum erat laoreet.
-Etiam rhoncus mauris arcu, ac egestas lacus facilisis nec.
-Pellentesque a orci vitae felis faucibus ultrices sed non massa.
-In posuere vel ex in elementum.
-Aenean ac neque a enim gravida tristique eu ac lacus.
-Ut tincidunt venenatis lectus, et maximus mauris.
-
-Proin ornare nulla id tristique placerat.
-Integer nec elementum tellus.
-Cras at tempus lectus.
-Cras lorem est, ornare id condimentum vitae, scelerisque eu lectus.
-Aenean libero arcu, consequat vel sagittis eget, sagittis ut neque.
-Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac turpis egestas.
-Aenean nec quam sodales, mollis est sit amet, vulputate turpis.
-Praesent tempus diam eu sem laoreet finibus.
-Nulla finibus, sapien vel lobortis ultricies, mauris ipsum aliquet urna, quis rutrum sapien dui id ipsum.
-Quisque sodales ex sapien, ac egestas lacus suscipit et.
-Suspendisse accumsan rhoncus tristique.
-Pellentesque ut eros lectus.
-Praesent viverra sit amet odio nec vulputate.
-Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos.
-Nulla molestie metus vitae libero vulputate, at interdum enim tempus.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2017-06-07 ms.tex
Har lagt til referanser og gjort noen smaa endringer i introduction. Det som naa mangler er en litt mer utfyllende diskusjon.

@@
 \begin{abstract} 
-In this paper, we aim to develop a method for producing statistically correct porous media using procedural noise functions.
-In short, the proposed model is a binary function that returns either true if the input point is inside pore space, or false if within a wall.
-By using procedural noise, we are not restricted to generating media within a limited box, as the noise functions is continuous on all scales.
-It is therefore possible to generate fractal-like geometries at arbitrary scales, circumventing the need for any physics simulations.
-In addition, this method is extremely fast and has low memory footprints, requiring very little computational power.
-After presenting such a model, we choose a method to measure the structural properties of nanoporous media that are created either through regular methods (e.g.
+New experimental and observational methods of the structure of three-dimensional porous media calls for improved approaches for statistical characterization and the generation of random samples with given statistical properties.
+We aim to develop a method for producing statistically consistent realization of porous media using procedural noise functions.
+By using procedural noise, we are not restricted to generating media with a predetermined resolution, since the noise functions is continuous on all scales.
+It is therefore possible to generate fractal-like geometries at arbitrary scales.
+This method is extremely fast and has low memory footprints, requiring very little computational power.
+We demonstrate the use of procedural noise functions by choosing a method to measure the structural properties of nanoporous media that are created either through regular methods (e.g.
 molecular dynamics) or with our method.
@@
 Finally, we generate a nanoporous media with an expand-and-quench technique using molecular dynamics and show that our model correctly reproduces several of the internal properties of the data set, including $g(r)$, porosity and surface area.
-  \todo{Abstract boer tenkes noe gjennom: faa frem hovedpoenger.}
-
 \end{abstract}
@@
 New experimental methods allow for studies of three-dimensional structures from nano-meter to meter scales.
-Destructive techniques such as focused ion beam scanning electron microscopy (FIB-SEM) are routinely used to characterize geologically relevant structures in shales and carbonate rocks with resolutions down to tens of nanometers \cite{XX}.
-High-energy X-ray tomography methods using high-energy beams at international research facilites or in-house scanners open for three-dimensional mapping of mineral structures from sub-micrometer to millimeter scales \cite{XX}.
+Destructive techniques such as focused ion beam scanning electron microscopy (FIB-SEM) are routinely used to characterize geologically relevant structures in shales and carbonate rocks with resolutions down to tens of nanometers \cite{cnudde2013high,curtis2010structural}.
+High-energy X-ray tomography methods using high-energy beams at international research facilites or in-house scanners open for three-dimensional mapping of mineral structures from sub-micrometer to millimeter scales \cite{kobchenko20114d,jamtveit2014pore,zhu2011microtomography,fusseis2009creep}.
 Thus new experimental techniques provide three-dimensional voxelized images that can form the basis of analysis or modeling.
@@
 
-The field of synthetic rocks is a large field both scientifically and technologically \cite{}. 
+The field of synthetic rocks is a large field both scientifically and technologically \cite{biswal2007stochastic,hamzehpour2006development,Reconstructing random media}. 
 Methods to generate synthentic rock samples from experimental characterizations may be based on purely statistical or geometrical models, models of physical processes that generate the samples, or combinations of both.
-For small-scale porous media, physical models are often used to generate initial grain packings \cite{}, then further mechano-chemical and fluid-mediated processes such as precipitation-dissoluton processes and pressure solution \cite{}, or fracture and plastic deformation processes are added to include effects of long-term compaction \cite{}.
-On large scales, reservoir models are populated by statistical models for lithological and geomorphological processes \cite{}, such as river fan formation to generate particular grain size distributions \cite{}.
-More statistical approaches include variations of percolation models \cite{sahimi} or network models \cite{} to characterize and generate uncorrelated or correlated porous media. 
+For small-scale porous media, physical models are often used to generate initial grain packings \cite{pilotti1998generation,scott1969density,herrmann2013physics}, then further mechano-chemical and fluid-mediated processes such as precipitation-dissoluton processes and pressure solution \cite{utter1983pressure}, or fracture and plastic deformation processes are added to include effects of long-term compaction \cite{rutter1983pressure,renard1997pressure}.
+On large scales, reservoir models are populated by statistical models for lithological and geomorphological processes \cite{pyrcz2014geostatistical}, such as river fan formation to generate particular grain size distributions \cite{oltermann1996heterogeneity,lair1994alluvial}.
+More statistical approaches include variations of percolation models \cite{sahimi} or network models \cite{fatt1956network,sahimi1993flow} to characterize and generate uncorrelated or correlated porous media.
 In addition, the high resolution of observational methods allow three-dimensional scans to form the basis for atomic-scale modeling using molecular modeling methods such as molecular dynamics or Monte Carlo methods to study the dynamics of nanoporous media.
@@
 In this paper, we have proposed a novel method for generating a nanoporous media using procedural methods in the form of simplex noise. 
-Procedural methods, albeit having no root in physics, have some advatages over prevalent methods: they are computationally efficient, have low memory footprints and are only limited in terms of detail by the model in question.
+Procedural methods are not physical, but rather statistical models, and this gives them some advatages over prevalent methods: they are computationally efficient, have low memory footprints and are only limited in terms of detail by the model in question.
 In particular, it is possible to create simulations of arbitrary scales and resolution, indeed with any level of detail.

### 2017-06-07 ms.tex
Ref til Porous media lagt til

@@
 New experimental and observational methods of the structure of three-dimensional porous media calls for improved approaches for statistical characterization and the generation of random samples with given statistical properties.
-We aim to develop a method for producing statistically consistent realization of porous media using procedural noise functions.
+We have developed a method for producing statistically consistent realization of porous media using procedural noise functions.
 By using procedural noise, we are not restricted to generating media with a predetermined resolution, since the noise functions is continuous on all scales.
@@
 Ideally, the model measure should pick up all geometrical properties that are relevant for the physical study of interest. 
-For instance, transport in porous media heavily depends on the pore size distribution and porosity, whereas chemical reactons are sensitive to surface area. \todo{referanser, malthe?} 
+For instance, transport in porous media heavily depends on the pore size distribution and porosity, whereas chemical reactons are sensitive to surface area \cite{coussy2011mechanics}.
 In this paper, we choose to focus on the radial distribution function $g(r)$ as defined and implemented in LAMMPS \cite{plimpton1995fast}.

### 2017-08-17 ms.tex
Fixed spelling of references

@@
 Methods to generate synthentic rock samples from experimental characterizations may be based on purely statistical or geometrical models, models of physical processes that generate the samples, or combinations of both.
-For small-scale porous media, physical models are often used to generate initial grain packings \cite{pilotti1998generation,scott1969density,herrmann2013physics}, then further mechano-chemical and fluid-mediated processes such as precipitation-dissoluton processes and pressure solution \cite{utter1983pressure}, or fracture and plastic deformation processes are added to include effects of long-term compaction \cite{rutter1983pressure,renard1997pressure}.
-On large scales, reservoir models are populated by statistical models for lithological and geomorphological processes \cite{pyrcz2014geostatistical}, such as river fan formation to generate particular grain size distributions \cite{oltermann1996heterogeneity,lair1994alluvial}.
+For small-scale porous media, physical models are often used to generate initial grain packings \cite{pilotti1998generation,scott1969density,herrmann2013physics}, then further mechano-chemical and fluid-mediated processes such as precipitation-dissoluton processes and pressure solution \cite{rutter1983pressure}, or fracture and plastic deformation processes are added to include effects of long-term compaction \cite{rutter1983pressure,renard1997pressure}.
+On large scales, reservoir models are populated by statistical models for lithological and geomorphological processes \cite{pyrcz2014geostatistical}, such as river fan formation to generate particular grain size distributions \cite{koltermann1996heterogeneity,blair1994alluvial}.
 More statistical approaches include variations of percolation models \cite{sahimi} or network models \cite{fatt1956network,sahimi1993flow} to characterize and generate uncorrelated or correlated porous media.
@@
 A much simpler example would be generating a 1 (or 2)-dimensional terrain by choosing random points on a grid and performing a spline interpolation between them to create the illusion of a smooth landscape.
-In fact, this is the very essence of procedural noise, which we are focusing on in our model, and the first person to create such a fast algorithm for N-dimensional spline interpolation between pseudo-random points was Ken Perlin.
+In fact, this is the very essence of procedural noise, which we are focusing on in our model, and the first person to create such a fast algorithm for N-dimensional spline interpolation between pseudo-random points was Ken Perlin\cite{perlin1985image}.
 
@@
 \end{equation}
-In one dimension, we show equation \ref{eq:perlinlinear} with 1, 3 and 30 octaves in figure \ref{fig:1dperlin}, while the 2D procedural noise pattern depicted in the lower
+In one dimension, we show equation \ref{eq:perlinlinear} with 1, 3 and 30 octaves in figure \ref{fig:1dperlin}, while the 2D procedural noise pattern is depicted in the lower
 part of figure \ref{fig:noise}.
@@
 \section{Model}
-In this paper, we are interested in developing a model using simplex noise to reproduce nanoporous structures that are as statistically close to physical simulations of porous media.
+In this paper, we are interested in developing a model using simplex noise to reproduce nanoporous structures that are statistically similar to physical simulations of porous media.
 For a given data set and a model, we wish to estimate the parameters that best represent the data. 
@@
 For instance, transport in porous media heavily depends on the pore size distribution and porosity, whereas chemical reactons are sensitive to surface area \cite{coussy2011mechanics}.
-In this paper, we choose to focus on the radial distribution function $g(r)$ as defined and implemented in LAMMPS \cite{plimpton1995fast}.
+In this paper, we choose to focus on the radial distribution function (RDF) $g(r)$ as defined and implemented in LAMMPS \cite{plimpton1995fast}.
 $g(r)$ describes how density varies as a function of distance from a reference particle, picking up both the intramolecular and intermolecular structure.
@@
 
-The method used for calculating porosity of a media is defined as in \cite{gelb1998characterization}, and we obtain the surface area of a porous media by using marching cubes triangulation to sum the surface area of the generated triangles.
+The method used for calculating porosity of a media is defined as in \cite{gelb1998characterization}, and we obtain the surface area of a porous media by using m arching cubes triangulation to sum the surface area of the generated triangles.
 As discussed in \cite{gelb1998characterization}, the definition of surface area is ill-defined in atomic simulations, because a clear definition of pore space is needed. 
@@
 Note that the surface atoms are not in chemical equilibrium since some neighbours of a surface atom might have been removed. 
-In order to use the generated geometry in molecular dynamics, one would need either noble gas atoms which may be stable, or apply passivation step.
+In order to use the generated geometry in molecular dynamics, one would need either noble gas atoms which may be stable, or to apply passivation step.
 

### 2018-05-29 ms.tex
Updated gofr and refs

@@
 
-The field of synthetic rocks is a large field both scientifically and technologically \cite{biswal2007stochastic,hamzehpour2006development,Reconstructing random media}. 
+The field of synthetic rocks is a large field both scientifically and technologically \citep{biswal2007stochastic,hamzehpour2006development, yeong1998reconstructing}. 
 Methods to generate synthentic rock samples from experimental characterizations may be based on purely statistical or geometrical models, models of physical processes that generate the samples, or combinations of both.
@@
 \begin{figure}
-\includegraphics[width=.45\textwidth]{gofr_plot.png}
-\caption{$g(r)$ for three data sets: the simulated SiO$_2$ input data set (red), the best-fit parameter model (blue) and a random low-fit parameter model (green).
-\todo{Bytt denne figuren med VMD sin beregning.
-Normaliseringen vår er gal.}}
+\includegraphics[width=0.5\textwidth]{gofr_figure.pdf}
+\caption{$g(r)$ for three data sets: the simulated SiO$_2$ input data set (red), the best-fit parameter model (blue) and a random low-fit parameter model (green).}
 \label{fig:gofr1}
@@
 
-%\begin{thebibliography}{9}
-
-%\bibitem[Groeneboom2014]{groeneboom2014} Groeneboom, N.~E., \& Dahle, H.\ 2014, \apj, 783, 138
-
-%\end{thebibliography}
-
 \end{document}

### 2018-05-29 ms.tex
Fixed errors and tex structure

@@
 
-The rapid development of new experimental and observational methods to determine the three-dimensional structure of many types of porous media calls for new statistical tools and methods to characterize complex porous media and to generate synthetic media based on the observed structures.
+The rapid development of new experimental and observational methods to determine the three-dimensional
+structure of many types of porous media calls for new statistical tools and methods to characterize
+complex porous media and to generate synthetic media based on the observed structures.
 New experimental methods allow for studies of three-dimensional structures from nano-meter to meter scales.
-Destructive techniques such as focused ion beam scanning electron microscopy (FIB-SEM) are routinely used to characterize geologically relevant structures in shales and carbonate rocks with resolutions down to tens of nanometers \cite{cnudde2013high,curtis2010structural}.
-High-energy X-ray tomography methods using high-energy beams at international research facilites or in-house scanners open for three-dimensional mapping of mineral structures from sub-micrometer to millimeter scales \cite{kobchenko20114d,jamtveit2014pore,zhu2011microtomography,fusseis2009creep}.
-Thus new experimental techniques provide three-dimensional voxelized images that can form the basis of analysis or modeling.
-However, in many cases we can only scan a few examples experimentally, but need to perform simulations over many statistically similar realizations to find effective properties or to populate upscaling models.
-Thus the need for methods to robustly characterize three-dimensional porous structures and to generate synthetic structures with similar statistical properties.
+Destructive techniques such as focused ion beam scanning electron microscopy (FIB-SEM) are routinely
+used to characterize geologically relevant structures in shales and carbonate rocks with resolutions
+down to tens of nanometers \cite{cnudde2013high,curtis2010structural}.
+High-energy X-ray tomography methods using high-energy beams at international research facilites or
+in-house scanners open for three-dimensional mapping of mineral structures from sub-micrometer to
+millimeter scales \cite{kobchenko20114d,jamtveit2014pore,zhu2011microtomography,fusseis2009creep}.
+Thus new experimental techniques provide three-dimensional voxelized images that can form the basis
+of analysis or modeling.
+However, in many cases we can only scan a few examples experimentally, but need to perform simulations
+over many statistically similar realizations to find effective properties or to populate upscaling models.
+Thus the need for methods to robustly characterize three-dimensional porous structures and to generate
+synthetic structures with similar statistical properties.
 
 The field of synthetic rocks is a large field both scientifically and technologically \citep{biswal2007stochastic,hamzehpour2006development, yeong1998reconstructing}. 
-Methods to generate synthentic rock samples from experimental characterizations may be based on purely statistical or geometrical models, models of physical processes that generate the samples, or combinations of both.
-For small-scale porous media, physical models are often used to generate initial grain packings \cite{pilotti1998generation,scott1969density,herrmann2013physics}, then further mechano-chemical and fluid-mediated processes such as precipitation-dissoluton processes and pressure solution \cite{rutter1983pressure}, or fracture and plastic deformation processes are added to include effects of long-term compaction \cite{rutter1983pressure,renard1997pressure}.
-On large scales, reservoir models are populated by statistical models for lithological and geomorphological processes \cite{pyrcz2014geostatistical}, such as river fan formation to generate particular grain size distributions \cite{koltermann1996heterogeneity,blair1994alluvial}.
-More statistical approaches include variations of percolation models \cite{sahimi} or network models \cite{fatt1956network,sahimi1993flow} to characterize and generate uncorrelated or correlated porous media.
-In addition, the high resolution of observational methods allow three-dimensional scans to form the basis for atomic-scale modeling using molecular modeling methods such as molecular dynamics or Monte Carlo methods to study the dynamics of nanoporous media.
-But then procedural or physical models are needed to fill in the atomic scale details, to build many statistical realizations of nanoporous media with atomic-scale details, and to extrapolate to larger scale models.
+Methods to generate synthentic rock samples from experimental characterizations may be based on 
+purely statistical or geometrical models, models of physical processes that generate the samples, or combinations of both. 
+For small-scale porous media, physical models are often used to generate initial grain
+packings \cite{pilotti1998generation,scott1969density,herrmann2013physics}, then further mechano-chemical
+and fluid-mediated processes such as precipitation-dissoluton processes and pressure
+solution \cite{rutter1983pressure}, or fracture and plastic deformation processes are added
+to include effects of long-term compaction \cite{rutter1983pressure,renard1997pressure}. 
+On large scales, reservoir models are populated by statistical models for lithological and
+geomorphological processes \cite{pyrcz2014geostatistical}, such as river fan formation to
+generate particular grain size distributions \cite{koltermann1996heterogeneity,blair1994alluvial}. 
+More statistical approaches include variations of percolation models \cite{sahimi} or
+network models \cite{fatt1956network,sahimi1993flow} to characterize and generate uncorrelated or correlated porous media. 
+In addition, the high resolution of observational methods allow three-dimensional scans to
+form the basis for atomic-scale modeling using molecular modeling methods such as
+molecular dynamics or Monte Carlo methods to study the dynamics of nanoporous media. 
+But then procedural or physical models are needed to fill in the atomic scale details, to
+build many statistical realizations of nanoporous media with atomic-scale details, and to extrapolate to larger scale models.
 
-These physically based models may not always include all relevant effects, can be quite computationally expensive when used to generate many samples, and cannot easily be extended beyond the initial size of the simulated systems.
-For example, to prepare a simulation of nanoporous silica, we need to carry out a sequence of large scale simulations in molecular dynamics, which is time consuming, computationally expensive and the results canno be exteded beyond the intial size of the simulated systems without performing another simulation \cite{Adarsh}.
-On the other hand, the statistical models often do not include all the correlations induced by physical processes or have limited number of tunable parameters. 
+These physically based models may not always include all relevant effects, can be quite
+computationally expensive when used to generate many samples, and cannot easily be extended
+beyond the initial size of the simulated systems. 
+For example, to prepare a simulation of nanoporous silica, we need to carry out a sequence
+of large scale simulations in molecular dynamics, which is time consuming, computationally
+expensive and the results cannot be exteded beyond the intial size of the simulated systems
+without performing another simulation \cite{Adarsh}. 
+On the other hand, the statistical models often do not include all the correlations induced
+by physical processes or have limited number of tunable parameters. 
 In particular, statistical models have challenges in reproducing atomic-scale structures of nanoporous media.
 
-Meanwhile, there has been a recent growth in advanced random noise models that open for new ways to generate structured noise with statistical similarities to real samples, such as the use of Perlin noise.
-Such noise models have been used to model systems with complex statistical correlations such as in GAMER\cite{groeneboom:2014}, and show great promise in being adaptable to many different physical systems.
-Also, such noise models are continuous and extendable, so that arbitrary large systems or systems with arbitrary high precision can be modeled. 
-While these models show great promise in many fields \cite{}, they have not yet been systematically applied to characterize or generate nanoporous media with statistical properties corresponding to that of real systems.
+Meanwhile, there has been a recent growth in advanced random noise models that open for new
+ways to generate structured noise with statistical similarities to real samples, such as the use of Perlin noise. 
+Such noise models have been used to model systems with complex statistical correlations such
+as in GAMER\cite{groeneboom:2014}, and show great promise in being adaptable to many different physical systems. 
+Also, such noise models are continuous and extendable, so that arbitrary large systems or
+systems with arbitrary high precision can be modeled. 
+While these models show great promise in several fields \cite{ebert2003texturing, cook2005wavelet, bridson2007curl, goldberg2008anisotropic},
+they have not yet been systematically applied to characterize or generate nanoporous media with statistical properties corresponding to that of real systems.
 
-In this paper, we propose a new method for producing realistic-looking nanoporous media using only random noise fields, without the need for physics simulations. 
-We provide a systematic study of the use of procedural noise-based models to characterize and generate three-dimensional nanoporous media. 
-Our studies will be based on silica-based nanoporous media, for which we have good atomic scale models and physics-based methods to generate porous media as a sequence of simulated processes, and we therefore can generate large-scale samples where we have full knowledge of all aspects of the material.
-These simulated systems are then characterized using the radial distribution function, $g(r)$, statistical methods are generated with similar properties, and compared with the simulated systems. 
-Using these approaches we find that procedural noise methods can be tuned to efficiently create materials that are structurally similar to simulationed materials.
+In this paper, we propose a new method for producing realistic-looking nanoporous media
+using only random noise fields, without the need for physics simulations. 
+We provide a systematic study of the use of procedural noise-based models to characterize
+and generate three-dimensional nanoporous media. 
+Our studies will be based on silica-based nanoporous media, for which we have good atomic
+scale models and physics-based methods to generate porous media as a sequence of simulated
+processes, and we therefore can generate large-scale samples where we have full knowledge of all aspects of the material. 
+These simulated systems are then characterized using the radial distribution function, $g(r)$,
+statistical methods are generated with similar properties, and compared with the simulated systems. 
+Using these approaches we find that procedural noise methods can be tuned to efficiently create
+materials that are structurally similar to simulationed materials.
 
 \section{Procedural content generation}
-By procedural content generation, we generally mean a mathematical algorithm that uses limited input data to create complex, sometimes infinite repeating patterns.
-A firm example would be fractals, such as the Julia/Mandelbrot [ref Mandelbrot] set, where one tests whether the complex quadratic recurrence equation $z_{n+1} = z_n^2 +C$ for $z \in \mathbb C$ converges for each position in a 2D complex plane.
-With this seemingly simple equation, the level of complexity generated on the boundary between convergence and divergence is infinite, with self-similar patterns repeating themselves to arbitrary detail.
+By procedural content generation, we generally mean a mathematical algorithm that uses limited
+input data to create complex, sometimes infinite repeating patterns.
+A firm example would be fractals, such as the Julia/Mandelbrot [ref Mandelbrot] set, where one
+tests whether the complex quadratic recurrence equation $z_{n+1} = z_n^2 +C$ for $z \in \mathbb C$
+converges for each position in a 2D complex plane.
+With this seemingly simple equation, the level of complexity generated on the boundary between
+convergence and divergence is infinite, with self-similar patterns repeating themselves to arbitrary detail.
 
-Another example, widely used in contemporary game content generation, is having prefabs, or pre-fabricated 3D models such as doors, walls, tiles and roof segments automatically being stitched together to create buildings.
-When designing a city, instead of creating each building manually, you only need specify the position, area and height of a structure.
-Then, by using an algorithm that combines these prefabs to render the building, it is possible to create cities of arbitrary size with a single command.
+Another example, widely used in contemporary game content generation, is having prefabs, or
+pre-fabricated 3D models such as doors, walls, tiles and roof segments automatically being
+stitched together to create buildings.
+When designing a city, instead of creating each building manually, you only need specify the
+position, area and height of a structure.
+Then, by using an algorithm that combines these prefabs to render the building, it is possible
+to create cities of arbitrary size with a single command.
 Naturally, it is also possible to create procedural methods that design the layout and placement of these buildings.
-See figure \ref{fig:orphancity} for an example of such a city generator tool, developed by one of the authors of this paper.
 
-A much simpler example would be generating a 1 (or 2)-dimensional terrain by choosing random points on a grid and performing a spline interpolation between them to create the illusion of a smooth landscape.
-In fact, this is the very essence of procedural noise, which we are focusing on in our model, and the first person to create such a fast algorithm for N-dimensional spline interpolation between pseudo-random points was Ken Perlin\cite{perlin1985image}.
+A much simpler example would be generating a 1 (or 2)-dimensional terrain by choosing random
+points on a grid and performing a spline interpolation between them to create the illusion of a smooth landscape.
+In fact, this is the very essence of procedural noise, which we are focusing on in our model,
+and the first person to create such a fast algorithm for N-dimensional spline interpolation
+between pseudo-random points was Ken Perlin\cite{perlin1985image}.
 
@@
 Perlin noise is a type of pseudo-random gradient noise that mimics a Gaussian field, but at low computational cost.
-An improved version of Perlin noise is called simplex noise, which has several advantages over standard Perlin noise in terms of scalability and speed \cite{perlin:2002}.
-Perlin noise was developed in response to the lack of detail 1980's contemporary graphics, where polygonal surfaces often were bland and left without any detail.
-In response to this problem, Perlin noise was developed in order to continuously perturb the light bouncing off flat surfaces, effectively yielding bump mapping and producing a much more realistic image.
+An improved version of Perlin noise is called simplex noise, which has several advantages
+over standard Perlin noise in terms of scalability and speed \cite{perlin:2002}.
+Perlin noise was developed in response to the lack of detail 1980's contemporary graphics,
+where polygonal surfaces often were bland and left without any detail.
+In response to this problem, Perlin noise was developed in order to continuously perturb
+the light bouncing off flat surfaces, effectively yielding bump mapping and producing a much more realistic image.
 
 An overview of the current state of procedural noise functions can be
-found in \cite{lagae:2010}. Perlin noise is a band limited repeatable pseudo-random function
+found in \cite{lagae2010state}. Perlin noise is a band limited repeatable pseudo-random function
 from $\mathbb R^n \to \mathbb R$ that is computationally fast, and has
@@
      between the the point and the current grid corner. 
-   \item [3.] Instead of linear interpolation $F(t) = t$, use a fade function to produce smoother interpolated values.
-Perlin noise typically uses the following fade function: $F(t) = t ^3 (t  (6 - 15) + 10)$.
+   \item [3.] Instead of linear interpolation $F(t) = t$, use a fade function to produce
+   smoother interpolated values. Perlin noise typically uses the following fade function: $F(t) = t ^3 (t  (6 - 15) + 10)$.
     \item [4.] Obtain the final value by linearly interpolating between the x-axis and
@@
 \end{itemize} 
-Simplex noise works in a similar manner, but works on on simplices instead of hypergrids, and so requires fewer computational steps, scales better in higher ($\ge 3$) dimensions and has fewer anisotropic properties.
+Simplex noise works in a similar manner, but works on on simplices instead of hypergrids,
+and so requires fewer computational steps, scales better in higher ($\ge 3$) dimensions
+and has fewer anisotropic properties.
 
@@
 a single octave. The upper-right part shows the same node for simplex noise,
-with less obvious ``textile'' features.
-However, when combining several octaves and adding a random shift to the coordinates, as depicted in the two lower parts, these anisotropic patterns tend to cancel
+with less obvious ``textile'' features. However, when combining several octaves
+and adding a random shift to the coordinates, as depicted in the two lower parts,
+these anisotropic patterns tend to cancel
 out.  In addition, we have calculated the statistical distributions for Perlin and simplex noise
@@
 to produce a wide range of nature-like patterns. A generic
-N-dimensional procedural function $\Phi(\vec x): \mathbb R^N \to
+$N$-dimensional procedural function $\Phi(\vec x): \mathbb R^N \to
 \mathbb R$ could be expressed as
@@
 where $f_s(\vec x,k)$ describes the frequency scale modifier (e.g. $f_s(\vec x,k) =
-k\vec x$), $\omega$ is the amount of octaves (modes), $\kappa(k)$ is the octave amplitude (e.g. $\kappa(k) = 1/k$)
+k\vec x$), $\Omega$ is the amount of octaves (modes), $\kappa(k)$ is the octave amplitude (e.g. $\kappa(k) = 1/k$)
 and $\Theta(x)$ an overall modifier (e.g. $\Theta(x) = x$ or
@@
 \end{equation}
-In one dimension, we show equation \ref{eq:perlinlinear} with 1, 3 and 30 octaves in figure \ref{fig:1dperlin}, while the 2D procedural noise pattern is depicted in the lower
-part of figure \ref{fig:noise}.
-Finally, since simplex noise both scales better with higher dimensions and produces less asymmetric effects, we choose to use simplex over regular Perlin noise as the main noise function throughout this paper.
+In one dimension, we show equation \ref{eq:perlinlinear} with 1, 3 and 30 octaves
+in figure \ref{fig:1dperlin}, while the 2D procedural noise pattern is depicted in the lower
+part of figure \ref{fig:noise}. Finally, since simplex noise both scales better
+with higher dimensions and produces less asymmetric effects, we choose to use
+simplex noise over regular Perlin noise as the main noise function throughout this paper.
 
@@
 \section{Model}
-In this paper, we are interested in developing a model using simplex noise to reproduce nanoporous structures that are statistically similar to physical simulations of porous media.
+In this paper, we are interested in developing a model using simplex noise to reproduce
+nanoporous structures that are statistically similar to physical simulations of porous media.
 For a given data set and a model, we wish to estimate the parameters that best represent the data. 
-This is performed through a full likelihood analysis in order to obtain the posterior probability distribution of the parameters. 
-The model measure of the analysis is the radial distribution function $g(r)$, for which the parameters will be maximised to fit. 
+This is performed through a full likelihood analysis in order to obtain the posterior
+probability distribution of the parameters. 
+The model measure of the analysis is the radial distribution function $g(r)$, for
+which the parameters will be adjusted to maximise.
 In addition, we also investigate the porosity and surface area of the system. 
-The simulated system is confined to a cube of size $\SI{159} {\angstrom}^3$, but since the procedural noise function is continuous on all scales, the box could be of arbitrary size.
+The simulated system is confined to a cube of size $\SI{159} {\angstrom}^3$, but since
+the procedural noise function is continuous on all scales, the box could be of arbitrary size.
 
@@
 \includegraphics[width=.95\textwidth]{model_examples.png}
-\caption{Examples of 3D simplex noise structures in nanoporous media.
-Left image: low persistence (large scale dominates), middle image: medium persistence (small scales starts dominating) and right image: high threshold (more atoms are discarded).
-}
+\caption{Examples of 3D simplex noise structures in nanoporous media. Left image:
+low persistence (large scale dominates), middle image: medium persistence (small
+scales starts dominating) and right image: high threshold (more atoms are discarded). }
 \label{fig:model_example}
@@
 \begin{equation}
-  M_1(\bar \theta) = M_1(\vec x, \psi, \Omega, \nu) = \sum_{\omega=0}^{\Omega} \psi^{\omega -1}   N\big((\vec x + \sigma \cdot \omega)\nu \cdot 2^{\omega-1} \big),
+  M_1(\bar \theta) = M_1(\vec x, \psi, \Omega, \nu) = \sum_{\omega=0}^{\Omega} \psi^{\omega -1}   N\big((\vec x + \vec\sigma \cdot \omega)\nu \cdot 2^{\omega-1} \big),
 \label{eq:noisemodel1}
 \end{equation}
-where $\Omega$ is the total amount of octaves, $\psi$ the persistence (scale falloff), $N$ a generic $N$-dimensional noise function, $\sigma$ a predefined random vector that removes anisotropic artefacts and $\nu$ the frequency modifier representing the initial largest scales in the model.
-For simplicity, $\bar \theta$ here denotes the set of model parameters.
-For each mode, the frequency is thus multiplied, while the overall amplitude is damped by a given persistence.
-In the end, we simulate a large set of bulk particles without any walls, and use the noise function to remove particles that are not within a procedural void given by a threshold $\tau$.
-Three examples of this model are depicted in figure \ref{fig:model_example}, where we have used $\Omega=6$ octaves and the scale is $\nu=0.01$ and cut-off $\tau=0$.
-The left image depicts low persistence ($\psi = 0.0$), where only the first octave dominates (corresponding to large scales).
-The middle picture shows the same parameters with a higher persistence $(\psi = 0.5)$, where smaller scales become more dominating.
+where $\Omega$ is the number of octaves, $\psi$ the persistence (scale falloff),
+$N$ a generic $N$-dimensional noise function, $\vec \sigma$ a predefined random
+vector that removes anisotropic artefacts and $\nu$ the frequency modifier representing
+the initial largest scales in the model. For simplicity, $\bar \theta$ here denotes
+the set of model parameters. For each mode, the frequency is thus multiplied, while
+the overall amplitude is damped by a given persistence. In the end, we simulate a
+large set of bulk particles without any walls, and use the noise function to remove
+particles that are not within a procedural void given by a threshold $\tau$. Three
+examples of this model are depicted in figure \ref{fig:model_example}, where we have
+used $\Omega=6$ octaves and the scale is $\nu=0.01$ and cut-off $\tau=0$. The left
+image depicts low persistence ($\psi = 0.0$), where only the first octave dominates
+(corresponding to large scales). The middle picture shows the same parameters with
+a higher persistence $(\psi = 0.5)$, where smaller scales become more dominating.
 In the rightmost picture, the effect of the cut-off threshold $\tau = -0.3$ is shown to emphasize threshold cut-off.
@@
 \subsection{Model Measure}
-In order to perform data analysis on the model parameters, we developed a code that uses Monte Carlo Markov chains to produce samples for building the posterior probability distribution for the parameters.
-Before doing so, we need a model measure - a quantitative method that characterizes the internal structure of the porous media.
-This quantity should be computed for both the original data set and the one produced by our noise model.
+In order to perform data analysis on the model parameters, we developed a code that
+uses Monte Carlo Markov chains to produce samples for building the posterior probability distribution for the parameters. 
+Before doing so, we need a model measure - a quantitative method that characterizes
+the internal structure of the porous media. This quantity should be computed for both
+the original data set and the one produced by our noise model.
 
-Ideally, the model measure should pick up all geometrical properties that are relevant for the physical study of interest. 
-For instance, transport in porous media heavily depends on the pore size distribution and porosity, whereas chemical reactons are sensitive to surface area \cite{coussy2011mechanics}.
-In this paper, we choose to focus on the radial distribution function (RDF) $g(r)$ as defined and implemented in LAMMPS \cite{plimpton1995fast}.
-$g(r)$ describes how density varies as a function of distance from a reference particle, picking up both the intramolecular and intermolecular structure.
-We have chosen to ignore the short range contributions of $g(r)$ since the pore geometry does not depend on the intra-molecular structure of SiO$_2$, but only the density variations at larger distances.
-Hereafter, $g_d(r)$ denotes the RDF calculated from simulated or experimental data, while $g_{M(\bar \theta)}(r)$ represents the RDF based on the procedural noise model $M(\bar \theta)$, where $\bar \theta$ is the set of model parameters.
+Ideally, the model measure should pick up all geometrical properties that are
+relevant for the physical study of interest. 
+For instance, transport in porous media heavily depends on the pore size
+distribution and porosity, whereas chemical reactons are sensitive to surface area \cite{coussy2011mechanics}.
+In this paper, we choose to focus on the radial distribution function (RDF)
+$g(r)$ as defined and implemented in LAMMPS \cite{plimpton1995fast}.
+$g(r)$ describes how density varies as a function of distance from a 
+reference particle, picking up both the intramolecular and intermolecular structure. 
+We have chosen to ignore the short range contributions of $g(r)$ since
+the pore geometry does not depend on the intra-molecular structure of
+SiO$_2$, but only the density variations at larger distances. 
+Hereafter, $g_d(r)$ denotes the RDF calculated from simulated or experimental data,
+while $g_{M(\bar \theta)}(r)$ represents the RDF based on the procedural noise
+model $M(\bar \theta)$, where $\bar \theta$ is the set of model parameters.
 
-The method used for calculating porosity of a media is defined as in \cite{gelb1998characterization}, and we obtain the surface area of a porous media by using m arching cubes triangulation to sum the surface area of the generated triangles.
-As discussed in \cite{gelb1998characterization}, the definition of surface area is ill-defined in atomic simulations, because a clear definition of pore space is needed. 
-In the marching cubes triangulation, we create a voxelized grid, a scalar field with values being the distance from the voxel centre to the nearest atom. 
-The isovalue threshold defining the surface is chosen to be \SI{2}{\angstrom}, large enough to contribute to a dense pore matrix, but small enough to have most of the available pore space defined as such.
-The important note here is that the same definition of both surface and porosity is being used in all systems that are compared, physical simulations, procedural methods or data.
+The method used for calculating porosity of a media is defined as in \cite{gelb1998characterization},
+and we obtain the surface area of a porous media by using m arching cubes triangulation to sum the
+surface area of the generated triangles.
+As discussed in \cite{gelb1998characterization}, the definition of surface area is ill-defined in
+atomic simulations, because a clear definition of pore space is needed. 
+In the marching cubes triangulation, we create a voxelized grid, a scalar field with values being
+the distance from the voxel centre to the nearest atom. 
+The isovalue threshold defining the surface is chosen to be \SI{2}{\angstrom}, large enough to
+contribute to a dense pore matrix, but small enough to have most of the available pore space defined as such. 
+The important note here is that the same definition of both surface and porosity is being used
+in all systems that are compared, physical simulations, procedural methods or data.
 
@@
 
-The posterior probability distribution is defined as the probability of the parameters $\bar \theta$ given the evidence $d$, or $P(\bar \theta | d)$. Bayes' theorem states that 
+The posterior probability distribution is defined as the probability of the parameters
+$\bar \theta$ given the evidence $d$, or $P(\bar \theta | d)$. Bayes' theorem states that 
 \begin{equation}
@@
 \end{equation}
-where $P(d | \bar \theta)$ is called the Likelihood $\mathcal L$ (the probability of evidence given parameters) and $P(\bar \theta)$ is the prior (the beliefs about the quantity before some evidence is taken into account).
-$P(d)$ is called the evidence, and is the probability of the observations (here assumed to be 1).
-Prior parameters are restricted to freely roam within a minimum and maximum value, so we can therefore assume that the posterior distribution is proportional to the likelihood, or $P(d | \bar \theta) \propto \mathcal L$.
-In other words, to calculate the posterior probability distribution and obtaining the best fit model parameters, we need only to estimate the likelihood function, which is connected to the $\chi^2$ by
+where $P(d | \bar \theta)$ is called the Likelihood $\mathcal L$ (the probability of
+evidence given parameters) and $P(\bar \theta)$ is the prior (the beliefs about the
+quantity before some evidence is taken into account). $P(d)$ is called the evidence,
+and is the probability of the observations (here assumed to be 1). Prior parameters
+are restricted to freely roam within a minimum and maximum value, so we can therefore
+assume that the posterior distribution is proportional to the likelihood, or
+$P(d | \bar \theta) \propto \mathcal L$. In other words, to calculate the posterior
+probability distribution and obtaining the best fit model parameters, we need only
+to estimate the likelihood function, which is connected to the $\chi^2$ by 
 \begin{equation}
@@
 
-In order to estimate whether two 3D data sets are statistically equivalent up to the model measure $g(r)$, we calculate the $\chi^2$ between the two one-dimensional RDFs.
-Pearson's $\chi^2$ is defined as
+In order to estimate whether two 3D data sets are statistically equivalent up to
+the model measure $g(r)$, we calculate the $\chi^2$ between the two one-dimensional
+RDFs. Pearson's $\chi^2$ is defined as 
 \begin{equation}
@@
 \end{equation}
-so for low values of $\chi^2$, the better the fit between data and model, up to $g(r)$.
-For a single-parameter model, we could simply just iterate over the lone parameter to obtain the minimum $\chi^2$ (where the best-fit parameter value resides), however, with increased number of parameters comes increased computational cost.
-If we divide the parameter space into $N$ bins, the number of $\chi^2$ computations scales as $N^d$, where $d$ is the number of model parameters.
-To be able to build the full landscape of the posterior probability distribution, we need to use the Monte Carlo Markov chain (MCMC) method .
+so for low values of $\chi^2$, the better the fit between data and model, up to
+$g(r)$. For a single-parameter model, we could simply just iterate over the lone
+parameter to obtain the minimum $\chi^2$ (where the best-fit parameter value resides),
+however, with increased number of parameters comes increased computational cost.
+If we divide the parameter space into $N$ bins, the number of $\chi^2$ computations
+scales as $N^d$, where $d$ is the number of model parameters. To be able to build
+the full landscape of the posterior probability distribution, we need to use the Monte Carlo Markov chain (MCMC) method .
 
-We are now interested in maximising the likelihood $\mathcal L$ as opposed to minimising the $\chi^2$, with the goal of mapping out the full $d$-dimensional likelihood space for the model parameters.
-We build the $d$-dimensional likelihood by letting random walkers traverse the parameter space guided by the MCMC test: for each step $M(\bar \theta_{n})$ in parameter space, we calculate the likelihood $\mathcal L_n$ for this particular parameter configuration.
-Then, for the next step, we create a new set of parameters $M(\bar \theta_{n+1})$ by choosing new parameters randomly from a $d$-dimensional sphere around the previous parameter composition.
-The model also selects a new random seed.
-We calculate the proposed likelihood $\mathcal L_{n+1}$ at this new step and perform the MCMC test: accept the new step if $\frac{\mathcal L_n}{\mathcal L_n+1} \ge 1$, or reject if the value is lower than a uniform random value $\in [0,1]$.
+We are now interested in maximising the likelihood $\mathcal L$ as opposed to
+minimising the $\chi^2$, with the goal of mapping out the full $d$-dimensional
+likelihood space for the model parameters. We build the $d$-dimensional likelihood
+by letting random walkers traverse the parameter space guided by the MCMC test:
+for each step $M(\bar \theta_{n})$ in parameter space, we calculate the likelihood
+$\mathcal L_n$ for this particular parameter configuration. Then, for the next
+step, we create a new set of parameters $M(\bar \theta_{n+1})$ by choosing new
+parameters randomly from a $d$-dimensional sphere around the previous parameter
+composition. The model also selects a new random seed. We calculate the proposed
+likelihood $\mathcal L_{n+1}$ at this new step and perform the MCMC test: accept
+the new step if $\frac{\mathcal L_n}{\mathcal L_n+1} \ge 1$, or reject if the value
+is lower than a uniform random value $\in [0,1]$.
 
-Finally, we obtain the posterior $P(\bar \theta | d) \propto \mathcal L$ by building a $d$-dimensional histogram of the likelihood using the stored positions of the parameters.
-For a sufficiently large amount of steps, the MCMC algorithm ensures that the posterior converges to the likelihood, as long at the surface does not contain unreachable areas such as discontinuous patches or an infinite amount of maxima (as is the case for the function $f(x)=sin(\frac{1}{x}$).
-Luckily, as we shall see, the likelihood surface for our models tend to turn out with a single, well-defined maximum.
+Finally, we obtain the posterior $P(\bar \theta | d) \propto \mathcal L$ by building
+a $d$-dimensional histogram of the likelihood using the stored positions of the parameters.
+For a sufficiently large amount of steps, the MCMC algorithm ensures that the posterior
+converges to the likelihood, as long at the surface does not contain unreachable areas
+such as discontinuous patches or an infinite amount of maxima (as is the case for the
+function $f(x)=sin(\frac{1}{x}$). Luckily, as we shall see, the likelihood surface for
+our models tend to turn out with a single, well-defined maximum.
 
@@
 We now turn our attention to applying the MCMC framework on two kinds of data.
-First, we wish to test and verify the code by performing a full data analysis on mock data with known input parameters.
-This enables us to investigate the posterior distribution, and determine whether any potential problems could arise with the likelihood landscape, such as random walkers trapped in local minima.
-In addition, we will obtain an intuition on the width of the estimated parameter distributions.
-But most importantly, we will have verified that our likelihood code is working correctly if we successfully reproduce the original input parameters.
-We then continue with a full analysis of simulated SiO$_2$ data, and see whether the resulting best-fit parameter data set is consistent with the SiO$_2$ data set.
+First, we wish to test and verify the code by performing a full data analysis on
+mock data with known input parameters. This enables us to investigate the posterior
+distribution, and determine whether any potential problems could arise with the
+likelihood landscape, such as random walkers trapped in local minima. In addition,
+we will obtain an intuition on the width of the estimated parameter distributions.
+But most importantly, we will have verified that our likelihood code is working correctly
+if we successfully reproduce the original input parameters. We then continue with a
+full analysis of simulated SiO$_2$ data, and see whether the resulting best-fit
+parameter data set is consistent with the SiO$_2$ data set.
 
@@
 \end{figure}
-In order to verify our likelihood code, we perform a model verification test on a mock data set with known input values based on equation \ref{eq:noisemodel1}.
-In this example, we only use standard simplex noise with three free parameters: the persistence $\psi$, threshold $\tau$ and scale $\nu$.
-We simulate a data set of size $\SI{159} {\angstrom}^3$ with $\sim 255 000$ atoms, and perform the cut-off given by input parameters $\Omega=3$, $\Psi = 0.6$, $\tau=-0.2$ and $\nu=0.03$.
+In order to verify our likelihood code, we perform a model verification test
+on a mock data set with known input values based on equation \ref{eq:noisemodel1}.
+In this example, we only use standard simplex noise with three free parameters:
+the persistence $\psi$, threshold $\tau$ and scale $\nu$. We simulate a data set
+of size $\SI{159} {\angstrom}^3$ with $\sim 255 000$ atoms, and perform the cut-off
+given by input parameters $\Omega=3$, $\Psi = 0.6$, $\tau=-0.2$ and $\nu=0.03$.
 The resulting 3D structure is depicted in figure \ref{fig:mockdata}.
 
-We continue by executing a full MCMC analysis on the mock data by using a parallelized code written in C++ running on a 48 core cluster.
-About 48 hours was needed to produce about 300 000 samples.
-The results from this analysis can be seen in figure \ref{fig:mockdataresults_pts}, where the upper panel depicts the marginalised 1D posteriors while the lower shows the marginalised 2D posteriors.
-The centre and left image shows that the input parameters of $\tau=-0.2$(threshold), $\psi=0.6$(persistence) and $\nu=0.03$ are all reproduced well within $1 \sigma$.
-Note that there seems to be no problems with the posterior distribution in terms of local minima or multiple maxima.
-In addition, the results show that all the parameters are correlated with $r \sim 0.85$, while the width of the marginalised distributions vary by $\sigma_\tau = 0.02 \sim 10\%$, $\sigma_\Psi = 0.062 \sim 20\%$ and $\sigma_\nu = 0.01 \sim 33\%$.
+We continue by executing a full MCMC analysis on the mock data by using a
+parallelized code written in C++ running on a 48 core cluster. About 48 hours
+was needed to produce about 300 000 samples. The results from this analysis
+can be seen in figure \ref{fig:mockdataresults_pts}, where the upper panel
+depicts the marginalised 1D posteriors while the lower shows the marginalised
+2D posteriors. The centre and left image shows that the input parameters of
+$\tau=-0.2$(threshold), $\psi=0.6$(persistence) and $\nu=0.03$ are all reproduced
+well within $1 \sigma$. Note that there seems to be no problems with the posterior
+distribution in terms of local minima or multiple maxima. In addition, the results
+show that all the parameters are correlated with $r \sim 0.85$, while the width of
+the marginalised distributions vary by $\sigma_\tau = 0.02 \sim 10\%$, $\sigma_\Psi = 0.062 \sim 20\%$ and $\sigma_\nu = 0.01 \sim 33\%$.
 
@@
 \includegraphics[width=.99\textwidth]{mock_data_results_pts.png}
-\caption{Results from the model validation simulation.
-The upper panel displays the marginalised 1D posteriors for each parameter, while the lower panel depicts the marginalised 2D posteriors.
+\caption{Results from the model validation simulation.The upper panel displays the marginalised 1D posteriors for each parameter, while the lower panel depicts the marginalised 2D posteriors.
 Note that all input parameters were reproduced well within $1 \sigma$ (inner contour line).}
@@
 performing a parameter fitting analysis on realistic simulated nanoporous media.
-Using LAMMPS \cite{plimpton1995fast} and the Vashishta-potential \cite{vashishta1990interaction}, we simulate system consisting of $14^3$ betacristobalite unit cells of SiO$_2$, corresponding to system size of approximately \SI{100}{\angstrom}.
+Using LAMMPS \cite{plimpton1995fast} and the Vashishta-potential \cite{vashishta1990interaction},
+we simulate system consisting of $14^3$ betacristobalite unit cells of SiO$_2$, corresponding
+to system size of approximately \SI{100}{\angstrom}. 
 The temperature is then increased to \SI{4500}{\kelvin} using NVT integration. 
 After an initial period, we expand the box to \SI{159}{\angstrom} to achieve a porosity $\phi=0.75$. 
-Then, the system is quenched by setting the temperature to \SI{300}{\kelvin}, resulting in the nanoporous media on the right side in figure \ref{fig:porous_vs_model}. 
-By choosing a different seed for the initial velocities, it is possible to produce statistically equal samples but with random realizations.
+Then, the system is quenched by setting the temperature to \SI{300}{\kelvin},
+resulting in the nanoporous media on the right side in figure \ref{fig:porous_vs_model}. 
+By choosing a different seed for the initial velocities, it is possible to produce
+statistically equal samples but with random realizations.
 
-With the model described in equation \ref{eq:noisemodel1}, we perform a full likelihood analysis of the data set with the three free parameters $\tau$, $\nu$ and $\psi$ using $g(r)$ as the measure. 
-The resulting marginalised posteriors are depicted in figure \ref{fig:porous_results1}, where the best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$ is extracted from the full 3D likelihood. 
-We also display $g(r)$ for the input data set, a best-fit parameter data set and a bad-fit data set in figure \ref{fig:gofr1}, to illustrate the sensitivity of $g(r)$ given the model parameters. 
-Note that we are not able to accurately reproduce the structures of the SiO$_2$ $g(r)$ with our current model, but hope that we during future work will be able to produce another model that provides a better fit with this measure.
+With the model described in equation \ref{eq:noisemodel1}, we perform a full likelihood
+analysis of the data set with the three free parameters $\tau$, $\nu$ and $\psi$ using $g(r)$ as the measure. 
+The resulting marginalised posteriors are depicted in figure \ref{fig:porous_results1},
+where the best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$ is extracted from the full 3D likelihood. 
+We also display $g(r)$ for the input data set, a best-fit parameter data set and a
+bad-fit data set in figure \ref{fig:gofr1}, to illustrate the sensitivity of $g(r)$ given the model parameters. 
+Note that we are not able to accurately reproduce the structures of the SiO$_2$ $g(r)$ with
+our current model, but hope that we during future work will be able to produce another model
+that provides a better fit with this measure. 
 \todo{skriv om hvorfor vi ikke plukker opp alle features?}
@@
 Note that the surface atoms are not in chemical equilibrium since some neighbours of a surface atom might have been removed. 
-In order to use the generated geometry in molecular dynamics, one would need either noble gas atoms which may be stable, or to apply passivation step.
+In order to use the generated geometry in molecular dynamics, one would need either noble gas
+atoms which may be stable, or to apply passivation step.
 
@@
 \end{figure}
-Finally, we depict a comparison of the actual simulated input SiO$_2$ data set together with a representation of the best-fit parameters in figure \ref{fig:porous_vs_model}   
+Finally, we depict a comparison of the actual simulated input SiO$_2$ data set together with a
+representation of the best-fit parameters in figure \ref{fig:porous_vs_model}   
 \begin{figure}
 \includegraphics[width=.45\textwidth]{comparison.png}
-\caption{Right side: A realisation of the best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$.
-Left side: Simulated SiO$_2$ input data set used for parameter estimation.
+\caption{Right side: A realisation of the best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$.Left side: Simulated SiO$_2$ input data set used for parameter estimation.
 Although the pore geometry seems to be similar, we see some artifacts, such as single loose atoms in the generated geometry.}
@@
 
-The porosity of the input SiO$_2$ data set is $75\%$, while the best-fit data set has porosity $77\% \pm 3 \%$, depending on the random seed.
-In addition, the surface area of the original data set was $\sim \SI{29000}{\angstrom^2}$, while the best-fit data shows similar values with surface are ranging between $\sim \SI{28000}{\angstrom^2}$ and $\sim \SI{30000}{\angstrom^2}$.
+The porosity of the input SiO$_2$ data set is $75\%$, while the best-fit data set has
+porosity $77\% \pm 3 \%$, depending on the random seed. In addition, the surface area
+of the original data set was $\sim \SI{29000}{\angstrom^2}$, while the best-fit data
+shows similar values with surface are ranging between $\sim \SI{28000}{\angstrom^2}$ and $\sim \SI{30000}{\angstrom^2}$.
 \todo{legg ved 3D plot av likelihood for aa vise at den er kjekk og fin}
@@
 \subsection{Discussion}
-It is worth to note that since we are chiefly interested in obtaining a set of parameters that reproduces a statistical correct representation of the input data, we are relatively free to choose parameters as desired within the posterior distribution width of the parameter.
-In the previous analysis, we selected a best fit threshold value of $\tau=-0.2$, but from the width of the likelihood it is clear that nearby parameters would also be suitable. 
-However, the surface area is strongly correlated with this value, so even slight deviations (e.g. $\Delta \tau = 0.05$) yields a relatively large change in surface area. 
-In addition, we do not claim that this model completely reproduces the pore structure observed in nanoporous simulations, as is evident from the RDF shown in figure \ref{fig:gofr1}, but hope to discover a better model that will produce a $g(r)$ that is more in accordance with the data.
+It is worth to note that since we are chiefly interested in obtaining a set of
+parameters that reproduces a statistical correct representation of the input data,
+we are relatively free to choose parameters as desired within the posterior distribution width of the parameter.
+In the previous analysis, we selected a best fit threshold value of $\tau=-0.2$,
+but from the width of the likelihood it is clear that nearby parameters would also be suitable. 
+However, the surface area is strongly correlated with this value, so even slight
+deviations (e.g. $\Delta \tau = 0.05$) yields a relatively large change in surface area. 
+In addition, we do not claim that this model completely reproduces the pore structure
+observed in nanoporous simulations, as is evident from the RDF shown in figure \ref{fig:gofr1},
+but hope to discover a better model that will produce a $g(r)$ that is more in accordance with the data.
 
@@
 In terms of the data, the variation is attributed as the initial velocities of the particles in the LAMMPS simulation. 
-For the model, it corresponds to the random seed. We found that varying these initial conditions have negligible effect on the calculation of $g(r)$.
+For the model, it corresponds to the random seed. We found that varying these
+initial conditions have negligible effect on the calculation of $g(r)$.
 
@@
 
-In this paper, we have proposed a novel method for generating a nanoporous media using procedural methods in the form of simplex noise.
-Procedural methods are not physical, but rather statistical models, and this gives them some advatages over prevalent methods: they are computationally efficient, have low memory footprints and are only limited in terms of detail by the model in question.
+In this paper, we have proposed a novel method for generating a nanoporous media
+using procedural methods in the form of simplex noise.
+Procedural methods are not physical, but rather statistical models, and this gives
+them some advatages over prevalent methods: they are computationally efficient, have low memory footprints and are only limited in terms of detail by the model in question. 
 In particular, it is possible to create simulations of arbitrary scales and resolution, indeed with any level of detail.
 
-\todo{AMS: This is a bit weak currently. State first what you have succeeded in doing. Then state the possible weaknesses of the model.}
-We have developed a framework to generate nanoporous media using molecular dynamics simulations and to estimate model parameters given such data. 
-We have shown that it is possible through a full likelihood analysis to successfully reproduce the input parameters of procedurally simulated data. 
-In addition, we performed an analysis on a simulated $\SI{159}{\angstrom^{3}}$ data set of SiO$_2$, and explain that while porosity and surface area correspond well between the model and data, analysis of $g(r)$ shows that the current model reproduces the main features of the material, but that there are some deviations.
-We find the correspondence strong, which demonstrates the proof-of-principle of the use of procedural noise methods, but we realize that even better correspondence may be achieved by including other summations in the procedural noise modes.
+\todo{AMS: This is a bit weak currently. State first what you have succeeded in doing.
+Then state the possible weaknesses of the model.}
+We have developed a framework to generate nanoporous media using molecular dynamics simulations and
+to estimate model parameters given such data. 
+We have shown that it is possible through a full likelihood analysis to successfully reproduce the
+input parameters of procedurally simulated data. 
+In addition, we performed an analysis on a simulated $\SI{159}{\angstrom^{3}}$ data set of SiO$_2$,
+and explain that while porosity and surface area correspond well between the model and data,
+analysis of $g(r)$ shows that the current model reproduces the main features of the material, but that there are some deviations. 
+We find the correspondence strong, which demonstrates the proof-of-principle of the use of
+procedural noise methods, but we realize that even better correspondence may be achieved by
+including other summations in the procedural noise modes.
 
 \subsection{Future work}
-We have only considered the simplest kind of procedural noise: a simple sum over modes with threshold cut-off $\tau$, scale fall-off persistence $\psi$ and initial scale $\nu$. 
-We have also seen that, even through both porosity and surface area accurately reproduced, there are features in the RDF that are not fully captured with the current model. 
-This is evident from the images of the simulated structure seen in figure \ref{fig:porous_vs_model}, where the connecting nearly one-atom thin walls are difficult to reproduce with the current model. 
-However, as procedural noise exert properties closely related to that of fractals, it is possible to extend the number of parameters and define new models with very different properties. 
-Some examples of these models are depicted in figure \ref{fig:future_models}, where the left- and centre images show a new model with two new parameters (offset and gain), while the rightmost panel shows a simulation where the scale $\nu$ is itself being perturbed by procedural noise, yielding strong circular asymmetric patterns.
+We have only considered the simplest kind of procedural noise: a simple sum over modes with
+threshold cut-off $\tau$, scale fall-off persistence $\psi$ and initial scale $\nu$. 
+We have also seen that, even through both porosity and surface area accurately reproduced,
+there are features in the RDF that are not fully captured with the current model. 
+This is evident from the images of the simulated structure seen in figure \ref{fig:porous_vs_model},
+where the connecting nearly one-atom thin walls are difficult to reproduce with the current model. 
+However, as procedural noise exert properties closely related to that of fractals, it is possible
+to extend the number of parameters and define new models with very different properties. 
+Some examples of these models are depicted in figure \ref{fig:future_models}, where the left- and
+centre images show a new model with two new parameters (offset and gain), while the rightmost panel shows a simulation where the scale $\nu$ is itself being perturbed by procedural noise, yielding strong circular asymmetric patterns.
 
 Another immediate application would be to estimate model parameters given experimental observations of porous media. 
-This would enable us to estimate new parameters such as instrumental noise, deviations from symmetry (asymmetric properties), and pureness of samples. 
-In addition, by analysing just a few samples and developing a model that produces similar structures as in the specimens, we are able to produce an arbitrary amount of statistically correct samples, free from instrumental noise and artefacts.
+This would enable us to estimate new parameters such as instrumental noise, deviations from symmetry
+(asymmetric properties), and pureness of samples. 
+In addition, by analysing just a few samples and developing a model that produces similar structures
+as in the specimens, we are able to produce an arbitrary amount of statistically correct samples,
+free from instrumental noise and artefacts.
 
-A final key analysis would be to compare properties such as permeability through finite-element simulations of gas/liquid flow through our simulated structures, in addition to exerting shear forces, and comparing with actual simulated data from e.g.
-LAMMPS.
-Since the noise function is a continuous boolean function that only checks whether a position is within a wall or void given a threshold, we are not limited to any size of the simulated system, and simulated particles traversing the media has arbitrary range.
+A final key analysis would be to compare properties such as permeability through finite-element
+simulations of gas/liquid flow through our simulated structures, in addition to exerting shear forces,
+and comparing with actual simulated data from e.g. LAMMPS. 
+Since the noise function is a continuous boolean function that only checks whether a position is
+within a wall or void given a threshold, we are not limited to any size of the simulated system, and simulated particles traversing the media has arbitrary range.
 

### 2018-05-29 ms.tex
More fixes

@@
 to produce a wide range of nature-like patterns. A generic
-$N$-dimensional procedural function $\Phi(\vec x): \mathbb R^N \to
+$N$-dimensional procedural function $\Phi(\vec r): \mathbb R^N \to
 \mathbb R$ could be expressed as
@@
 \label{eq:procedural}
- \Phi(\vec x) = \Theta \Big(\sum_k^\Omega \kappa (k) P(  f_s (\vec x,k) )) \Big),
+ \Phi(\vec r) = \Theta \Big(\sum_k^\Omega \kappa (k) P(  f_s (\vec r,k) )) \Big),
  \end{equation}
-where $f_s(\vec x,k)$ describes the frequency scale modifier (e.g. $f_s(\vec x,k) =
-k\vec x$), $\Omega$ is the amount of octaves (modes), $\kappa(k)$ is the octave amplitude (e.g. $\kappa(k) = 1/k$)
+where $f_s(\vec r,k)$ describes the frequency scale modifier (e.g. $f_s(\vec r,k) =
+k\vec r$), $\Omega$ is the amount of octaves (modes), $\kappa(k)$ is the octave amplitude (e.g. $\kappa(k) = 1/k$)
 and $\Theta(x)$ an overall modifier (e.g. $\Theta(x) = x$ or
@@
 \label{eq:perlinlinear}
- \Phi(\vec x) = \sum_k^\Omega \frac{1}{k} P( 2k\vec x).
+ \Phi(\vec r) = \sum_k^\Omega \frac{1}{k} P( 2k\vec r).
 \end{equation}
@@
 \begin{equation}
-  M_1(\bar \theta) = M_1(\vec x, \psi, \Omega, \nu) = \sum_{\omega=0}^{\Omega} \psi^{\omega -1}   N\big((\vec x + \vec\sigma \cdot \omega)\nu \cdot 2^{\omega-1} \big),
+  M_1(\vec r |\bar \theta) = M_1(\vec r | \psi, \Omega, \nu) = \sum_{\omega=0}^{\Omega} \psi^{\omega -1}   N\big((\vec r + \vec\sigma \cdot \omega)\nu \cdot 2^{\omega-1} \big),
 \label{eq:noisemodel1}
@@
 the set of model parameters. For each mode, the frequency is thus multiplied, while
-the overall amplitude is damped by a given persistence. In the end, we simulate a
-large set of bulk particles without any walls, and use the noise function to remove
-particles that are not within a procedural void given by a threshold $\tau$. Three
-examples of this model are depicted in figure \ref{fig:model_example}, where we have
-used $\Omega=6$ octaves and the scale is $\nu=0.01$ and cut-off $\tau=0$. The left
+the overall amplitude is damped by a given persistence.
+
+We then get, for a given position $\vec r$, a scalar value from the noise function.
+Assume now that we have a system in molecular dynamics where all atomic positions are given.
+A porous material is then generated by introducing a threshold $\tau$ which is used to decide whether or not an atom should be removed from the system.
+A boolean function $\text{Delete}(\vec r)$ is defined as
+
+\begin{align}
+  \text{Delete}(\vec r) = 
+  \begin{cases} 
+    \text{True} & \text{if} \quad M_1(\vec r |\bar \theta) < \tau \\
+    \text{False} & \text{if} \quad M_1(\vec r |\bar \theta) \geq \tau,
+  \end{cases}
+\end{align}
+where we simply evaluate this function for every atom in the system.
+The threshold $\tau$ is now added as a parameter to the model.
+Three examples of this model are depicted in figure \ref{fig:model_example}, where we have
+used $\Omega=6$ octaves and scale $\nu=0.01$ with threshold $\tau=0$. The left
 image depicts low persistence ($\psi = 0.0$), where only the first octave dominates
@@
 a higher persistence $(\psi = 0.5)$, where smaller scales become more dominating.
-In the rightmost picture, the effect of the cut-off threshold $\tau = -0.3$ is shown to emphasize threshold cut-off.
+In the rightmost picture, the effect of the threshold $\tau = -0.3$ is shown to emphasize threshold cut-off.
 
@@
 relevant for the physical study of interest. 
-For instance, transport in porous media heavily depends on the pore size
-distribution and porosity, whereas chemical reactons are sensitive to surface area \cite{coussy2011mechanics}.
+For instance, transport in porous media heavily depends on the pore size distribution and porosity\cite{kozeny1927uber, carman1937fluid},
+whereas chemical reactons are sensitive to surface area\cite{coussy2011mechanics}.
 In this paper, we choose to focus on the radial distribution function (RDF)
-$g(r)$ as defined and implemented in LAMMPS \cite{plimpton1995fast}.
+$g(r)$ as defined and implemented in LAMMPS\cite{plimpton1995fast}.
 $g(r)$ describes how density varies as a function of distance from a 
@@
 
-The method used for calculating porosity of a media is defined as in \cite{gelb1998characterization},
-and we obtain the surface area of a porous media by using m arching cubes triangulation to sum the
-surface area of the generated triangles.
+The method used for calculating porosity of a porous media is the same as in \cite{gelb1998characterization}.
+Surface area is obtained by creating a voxelized grid of the system with values being the distance from each voxel center to the nearest atom.
+This is now a scalar field and we can apply marching cubes to triangulate an isosurface of the distance field.
+This gives a trianglated representation of the porous media defined by a set of atomic positions.
 As discussed in \cite{gelb1998characterization}, the definition of surface area is ill-defined in
-atomic simulations, because a clear definition of pore space is needed. 
-In the marching cubes triangulation, we create a voxelized grid, a scalar field with values being
-the distance from the voxel centre to the nearest atom. 
-The isovalue threshold defining the surface is chosen to be \SI{2}{\angstrom}, large enough to
-contribute to a dense pore matrix, but small enough to have most of the available pore space defined as such. 
-The important note here is that the same definition of both surface and porosity is being used
-in all systems that are compared, physical simulations, procedural methods or data.
+atomistic simulations, because a clear definition of pore space is difficult.
+We have chosen to define the pore space as the set of points being more than \SI{2}{\angstrom} to any atom.
+This is large enough to contribute to a dense pore matrix, but small enough to have most of the available pore space defined as such. 
+The important note here is that the same definition of both surface and porosity is being used in all systems that are compared, physical simulations, procedural methods or data.
 

### 2018-05-29 ms.tex
Added improved 1d noise figure, wrote more about model

@@
 \begin{figure}
-\includegraphics[width=.5\textwidth]{1d_perlin.png}
-\caption{1D simplex noise with 1 (blue), 3 (green) and 30 (red) octaves. }
+\includegraphics[width=.5\textwidth]{1d_noise.pdf}
+\caption{1-dimensional simplex noise with 1 (yellow), 3 (red) and 30 (blue) octaves. Adding more octaves adds high frequency features, but it still follows the 1 octave base value.}
 \label{fig:1dperlin}
@@
 In this paper, we are interested in developing a model using simplex noise to reproduce
-nanoporous structures that are statistically similar to physical simulations of porous media.
+nanoporous structures that are statistically similar to physical simulations generated by molecular dynamics.
+A series of statistical similar nanoporous materials are generated using molecular dynamics from which the model will be fitted to.
 For a given data set and a model, we wish to estimate the parameters that best represent the data. 
-This is performed through a full likelihood analysis in order to obtain the posterior
-probability distribution of the parameters. 
-The model measure of the analysis is the radial distribution function $g(r)$, for
-which the parameters will be adjusted to maximise.
-In addition, we also investigate the porosity and surface area of the system. 
-The simulated system is confined to a cube of size $\SI{159} {\angstrom}^3$, but since
-the procedural noise function is continuous on all scales, the box could be of arbitrary size.
+This is performed through a full likelihood analysis in order to obtain the posterior probability distribution of the parameters. 
+The model measure of the analysis is the radial distribution function (or pair correlation function) $g(r)$, a quantity that measures how the density varies as a function of the distance from an atom.
+For low $r$, there is typically a peak at the equilibrium distance between nearest neighbor atoms in crystals or molecules, and several more peaks are often well defined as next nearest neighbors usually obey some symmetries as well.
+The fitting procedure consists of measuring $g(r)$ for both the data and the model, and deviations contributes to the error $\chi^2$ (defined below).
+In addition, we also investigate the porosity and surface area of the system after fitting since these are important physical quantities in such systems.
 
@@
 \end{figure*}
-We construct a generic sum over procedural noise octaves with the expression: 
-\begin{equation}
-  M_1(\vec r |\bar \theta) = M_1(\vec r | \psi, \Omega, \nu) = \sum_{\omega=0}^{\Omega} \psi^{\omega -1}   N\big((\vec r + \vec\sigma \cdot \omega)\nu \cdot 2^{\omega-1} \big),
+
+Given a set of parameters $\bar \theta$, we construct a generic sum over procedural noise octaves with the expression:
+\begin{align}
+  M_1(\vec r |\bar \theta) &= M_1(\vec r | \psi, \Omega, \nu)\nonumber\\
+  &= \sum_{\omega=1}^{\Omega} \psi^{\omega-1}   N\big((\vec r + \vec\sigma \cdot \omega)\nu \cdot 2^{\omega-1} \big),
 \label{eq:noisemodel1}
-\end{equation}
-where $\Omega$ is the number of octaves, $\psi$ the persistence (scale falloff),
-$N$ a generic $N$-dimensional noise function, $\vec \sigma$ a predefined random
+\end{align}
+where $\Omega$ is the number of octaves, $\psi$ the persistence (the amplitude at different scales),
+$N$ a generic $N$-dimensional noise function (in this paper Perlin noise or simplex noise), $\vec \sigma$ a predefined random
 vector that removes anisotropic artefacts and $\nu$ the frequency modifier representing
-the initial largest scales in the model. For simplicity, $\bar \theta$ here denotes
-the set of model parameters. For each mode, the frequency is thus multiplied, while
-the overall amplitude is damped by a given persistence.
+the initial largest scales in the model. For each octave, the frequency is doubled, while the overall amplitude is damped by a certain persistence $\psi$.
 
-We then get, for a given position $\vec r$, a scalar value from the noise function.
-Assume now that we have a system in molecular dynamics where all atomic positions are given.
+We then get, for a given position in physical space $\vec r$, a scalar value from the noise function.
+Assume now that we have a bulk system in molecular dynamics where all atomic positions are given.
 A porous material is then generated by introducing a threshold $\tau$ which is used to decide whether or not an atom should be removed from the system.
@@
 The threshold $\tau$ is now added as a parameter to the model.
+
 Three examples of this model are depicted in figure \ref{fig:model_example}, where we have
@@
 In the rightmost picture, the effect of the threshold $\tau = -0.3$ is shown to emphasize threshold cut-off.
+The bulk system is a cubic system of amorphous SiO$_2$ with size \SI{159}{\angstrom}.
 
@@
 Surface area is obtained by creating a voxelized grid of the system with values being the distance from each voxel center to the nearest atom.
-This is now a scalar field and we can apply marching cubes to triangulate an isosurface of the distance field.
+This is now a scalar field and we can apply marching cubes\cite{lorensen1987marching} to triangulate an isosurface of the distance field.
 This gives a trianglated representation of the porous media defined by a set of atomic positions.

### 2018-05-29 ms.tex
More fixes

@@
 
-\subsection{Model Measure}
+\subsection{Model measure}
 In order to perform data analysis on the model parameters, we developed a code that
 uses Monte Carlo Markov chains to produce samples for building the posterior probability distribution for the parameters. 
-Before doing so, we need a model measure - a quantitative method that characterizes
-the internal structure of the porous media. This quantity should be computed for both
-the original data set and the one produced by our noise model.
+As mentioned previously, we have used the radial distribution function $g(r)$ as a measure to compare the model to data.
 
@@
 For instance, transport in porous media heavily depends on the pore size distribution and porosity\cite{kozeny1927uber, carman1937fluid},
-whereas chemical reactons are sensitive to surface area\cite{coussy2011mechanics}.
-In this paper, we choose to focus on the radial distribution function (RDF)
-$g(r)$ as defined and implemented in LAMMPS\cite{plimpton1995fast}.
-$g(r)$ describes how density varies as a function of distance from a 
-reference particle, picking up both the intramolecular and intermolecular structure. 
-We have chosen to ignore the short range contributions of $g(r)$ since
-the pore geometry does not depend on the intra-molecular structure of
-SiO$_2$, but only the density variations at larger distances. 
+whereas chemical reactons are sensitive to surface area\cite{coussy2011mechanics}. We have used $g(r)$ as defined and implemented in LAMMPS\cite{plimpton1995fast}.
+In our cost function (the $\chi^2$ measure), we have ignored the low $r$ values of $g(r)$ since the pore geometry does not depend on the intra-molecular structure, but only the density variations at larger distances.
+
 Hereafter, $g_d(r)$ denotes the RDF calculated from simulated or experimental data,
@@
 
+As an additional validation, we also compute the porosity and surface area of the model since these are important quantities for fluid transport.
 The method used for calculating porosity of a porous media is the same as in \cite{gelb1998characterization}.
@@
 The important note here is that the same definition of both surface and porosity is being used in all systems that are compared, physical simulations, procedural methods or data.
-
-\todo{skriv litt mer om hvorfor vi velger g(r)}
 

### 2018-05-30 ms.tex
More fixes

@@
 \includegraphics[width=.5\textwidth]{1d_noise.pdf}
-\caption{1-dimensional simplex noise with 1 (yellow), 3 (red) and 30 (blue) octaves. Adding more octaves adds high frequency features, but it still follows the 1 octave base value.}
+\caption{1-dimensional simplex noise with 1 (yellow), 3 (red) and 30 (blue) octaves. Adding more octaves adds high frequency features, but it still follows the 1 octave base signal.}
 \label{fig:1dperlin}
@@
   \end{cases}
+  \label{eq:delete_threshold}
 \end{align}
@@
 whereas chemical reactons are sensitive to surface area\cite{coussy2011mechanics}. We have used $g(r)$ as defined and implemented in LAMMPS\cite{plimpton1995fast}.
-In our cost function (the $\chi^2$ measure), we have ignored the low $r$ values of $g(r)$ since the pore geometry does not depend on the intra-molecular structure, but only the density variations at larger distances.
+In our cost function (the $\chi^2$ measure), we have ignored the values of $g(r)$ for $r<$\SI{3}{\angstrom} since the pore geometry does not depend on the intra-molecular structure, but only the density variations at larger distances.
+The cutoff value of \SI{3}{\angstrom} was chosen since the peak of $g(r)$ is at \SI{2.9}{\angstrom} for Si-O\cite{vashishta1990interaction}, and has .
 
@@
 This is now a scalar field and we can apply marching cubes\cite{lorensen1987marching} to triangulate an isosurface of the distance field.
-This gives a trianglated representation of the porous media defined by a set of atomic positions.
+This gives a triangulated representation of the porous media defined by a set of atomic positions.
 As discussed in \cite{gelb1998characterization}, the definition of surface area is ill-defined in
@@
 quantity before some evidence is taken into account). $P(d)$ is called the evidence,
-and is the probability of the observations (here assumed to be 1). Prior parameters
-are restricted to freely roam within a minimum and maximum value, so we can therefore
-assume that the posterior distribution is proportional to the likelihood, or
-$P(d | \bar \theta) \propto \mathcal L$. In other words, to calculate the posterior
-probability distribution and obtaining the best fit model parameters, we need only
-to estimate the likelihood function, which is connected to the $\chi^2$ by 
+and is the probability of the observations (here assumed to be 1). 
+To calculate the posterior probability distribution and obtain the best fit model parameters, we need only
+to estimate the likelihood function, which is connected to the $\chi^2$\todo{Why do we have an exponential dependence of $\chi^2$ here?} by 
 \begin{equation}
@@
 the model measure $g(r)$, we calculate the $\chi^2$ between the two one-dimensional
-RDFs. Pearson's $\chi^2$ is defined as 
+RDFs. We calculate $g(r)$ using a histogram over spherical shells of distance $r_i$ to the center for bin $i$. Pearson's $\chi^2$ is then defined as 
 \begin{equation}
-  \chi^2 = \sum_r \frac{ \Big(g_d(r) - g_{M(\bar \theta)}(r) \Big)^2}{g_{M(\bar \theta)}(r)^2},
+  \chi^2 = \sum_{i} \frac{ \Big(g_d(r_i) - g_{M(\bar \theta)}(r_i) \Big)^2}{g_{M(\bar \theta)}(r_i)^2},
 \end{equation}
-so for low values of $\chi^2$, the better the fit between data and model, up to
-$g(r)$. For a single-parameter model, we could simply just iterate over the lone
-parameter to obtain the minimum $\chi^2$ (where the best-fit parameter value resides),
-however, with increased number of parameters comes increased computational cost.
-If we divide the parameter space into $N$ bins, the number of $\chi^2$ computations
-scales as $N^d$, where $d$ is the number of model parameters. To be able to build
-the full landscape of the posterior probability distribution, we need to use the Monte Carlo Markov chain (MCMC) method .
+which means that for low values of $\chi^2$, the better the fit between data and model, up to $g(r)$.
+
+The difficult task is now to find which set of parameters $\bar\theta$ that maximizes the likelihood $\mathcal L$.
+For a single-parameter model with parameter $\psi$, we could simply just discretize the interval $[\psi_\text{min}, \psi_\text{max}]$ into $N$ values,
+and loop over all parameter to obtain the minimum $\chi^2$.
+However, each parameter is an extra dimension in the parameter space, with increases the computational cost substantially.
+If each parameter is discretized into into $N$ values, the full number of $\chi^2$ computations
+scales as $N^d$, where $d$ is dimension of the parameter space - i.e. the number of model parameters. To be able to build
+the full landscape of the posterior probability distribution, we have chosen to use the Monte Carlo Markov chain (MCMC) method.
 
@@
 parameters randomly from a $d$-dimensional sphere around the previous parameter
-composition. The model also selects a new random seed. We calculate the proposed
+composition. We calculate the proposed
 likelihood $\mathcal L_{n+1}$ at this new step and perform the MCMC test: accept
@@
 such as discontinuous patches or an infinite amount of maxima (as is the case for the
-function $f(x)=sin(\frac{1}{x}$). Luckily, as we shall see, the likelihood surface for
+function $f(x)=\sin(\frac{1}{x}$)). Luckily, as we shall see, the likelihood surface for
 our models tend to turn out with a single, well-defined maximum.
@@
 if we successfully reproduce the original input parameters. We then continue with a
-full analysis of simulated SiO$_2$ data, and see whether the resulting best-fit
-parameter data set is consistent with the SiO$_2$ data set.
-
+full analysis of physical system simulated with molecular dynamics, and seek parameters that reproduces $g(r)$ of this system.
+\todo{Figure 5 is the same as left in fig 9.}
 \subsection{Model verification: Parameter estimation of mock data}
@@
 In this example, we only use standard simplex noise with three free parameters:
-the persistence $\psi$, threshold $\tau$ and scale $\nu$. We simulate a data set
-of size $\SI{159} {\angstrom}^3$ with $\sim 255 000$ atoms, and perform the cut-off
-given by input parameters $\Omega=3$, $\Psi = 0.6$, $\tau=-0.2$ and $\nu=0.03$.
+the persistence $\psi$, threshold $\tau$ and scale $\nu$.
+
+The physical system is a periodic box with amorphous SiO$_2$ of size $\SI{159} {\angstrom}^3$ (255552 atoms), created with a heat-and-quench method.
+A nanoporous system is created by removing atoms from the system accordin to equation \eqref{eq:delete_threshold} with parameters $\Omega=3$, $\Psi = 0.6$, $\nu=0.03$ and $\tau=-0.2$.
 The resulting 3D structure is depicted in figure \ref{fig:mockdata}.
@@
 parallelized code written in C++ running on a 48 core cluster. About 48 hours
-was needed to produce about 300 000 samples. The results from this analysis
+was needed to produce about 300000 samples. The results from this analysis
 can be seen in figure \ref{fig:mockdataresults_pts}, where the upper panel
@@
 $\tau=-0.2$(threshold), $\psi=0.6$(persistence) and $\nu=0.03$ are all reproduced
-well within $1 \sigma$. Note that there seems to be no problems with the posterior
+well within $1 \ \sigma$ ($\Omega$ was kept fixed equal to 3). Note that there seems to be no problems with the posterior
 distribution in terms of local minima or multiple maxima. In addition, the results
@@
 \end{figure*}
-Having verified that the likelihood code yields correct results when analysing mock data, we turn to 
+Having verified that the likelihood code yields satisfying results when analysing mock data, we turn to 
 performing a parameter fitting analysis on realistic simulated nanoporous media.
-Using LAMMPS \cite{plimpton1995fast} and the Vashishta-potential \cite{vashishta1990interaction},
-we simulate system consisting of $14^3$ betacristobalite unit cells of SiO$_2$, corresponding
+Using LAMMPS\cite{plimpton1995fast} and the Vashishta-potential\cite{vashishta1990interaction},
+we simulate system consisting of $14^3$ stretched\footnote{The nanoporous material will be amorphous, so we have stretched the initial crystal to match the density of amorphous silica.} betacristobalite unit cells of SiO$_2$, corresponding
 to system size of approximately \SI{100}{\angstrom}. 
-The temperature is then increased to \SI{4500}{\kelvin} using NVT integration. 
-After an initial period, we expand the box to \SI{159}{\angstrom} to achieve a porosity $\phi=0.75$. 
-Then, the system is quenched by setting the temperature to \SI{300}{\kelvin},
+The system is then melted at \SI{4500}{\kelvin} using the Nose-Hoover thermostat.
+After a few nanoseconds, the system box is slowly expanded to \SI{159}{\angstrom} to achieve a porosity of $\phi=0.75$ while keeping the thermostat on to keep the temperature constant.
+Then, the system is quenched by setting the target thermostat temperature to \SI{300}{\kelvin},
 resulting in the nanoporous media on the right side in figure \ref{fig:porous_vs_model}. 
-By choosing a different seed for the initial velocities, it is possible to produce
-statistically equal samples but with random realizations.
+By repeating this for different initial velocities, we can produce statistically similar samples with random realizations.
 
@@
 The resulting marginalised posteriors are depicted in figure \ref{fig:porous_results1},
-where the best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$ is extracted from the full 3D likelihood. 
+where the best-fit parameters $\tau=-0.2$, $\psi=0.9$ and $\nu=0.03$ is extracted from the full 3D likelihood. 
 We also display $g(r)$ for the input data set, a best-fit parameter data set and a
 bad-fit data set in figure \ref{fig:gofr1}, to illustrate the sensitivity of $g(r)$ given the model parameters.
-Note that we are not able to accurately reproduce the structures of the SiO$_2$ $g(r)$ with
-our current model, but hope that we during future work will be able to produce another model
-that provides a better fit with this measure. 
+$\chi^2$ is only calculated for $r\in (3,15)$ since the nanoporous structure signal is found at larger $r$.
+The main features of $g(r)$ from the data is picked up by our best-fit model and the bad fit (which has very high porosity) has slower convergence.
+We see that we don't accurately reproduce all structures of the SiO$_2$ $g(r)$ with
+our current model. For instance, the long tail is consistantly slightly lower in our model, and some of the bumps have details that differ from what our model gives.
+
 \todo{skriv om hvorfor vi ikke plukker opp alle features?}
 
-Note that the surface atoms are not in chemical equilibrium since some neighbours of a surface atom might have been removed. 
+Also note that the surface atoms are not in chemical equilibrium since some neighbours of a surface atom might have been removed. 
 In order to use the generated geometry in molecular dynamics, one would need either noble gas

### 2018-05-30 ms.tex
Fixed more, and figure

@@
 \includegraphics[width=.45\textwidth]{model_test.png}
-\caption{Simulated mock data used for the initial model testing analysis. Input parameters are $\nu=0.03$,$\tau=-0.02$ and $\psi=0.6$.}
+\caption{Simulated mock data used for the initial model testing analysis. Input parameters are $\nu=0.03$,$\tau=-0.2$ and $\psi=0.6$.}
 \label{fig:mockdata}
@@
 The upper panel displays the marginalised 1D posteriors for each parameter, while the lower panel depicts the marginalised 2D posteriors.
-Best fit parameters from the maximum 3D likelihood suggest $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$.
+Best fit parameters from the maximum 3D likelihood suggest $\tau=-0.2$, $\psi=0.9$ and $\nu=0.03$.
 }
@@
 $\chi^2$ is only calculated for $r\in (3,15)$ since the nanoporous structure signal is found at larger $r$.
-The main features of $g(r)$ from the data is picked up by our best-fit model and the bad fit (which has very high porosity) has slower convergence.
-We see that we don't accurately reproduce all structures of the SiO$_2$ $g(r)$ with
-our current model. For instance, the long tail is consistantly slightly lower in our model, and some of the bumps have details that differ from what our model gives.
+The main features of $g(r)$ from the data is reproduced by our best-fit model, whereas the bad fit (which has higher porosity) clearly differs.
+It converges slowly due to the large pores in the system. Even if all $g(r)$ are normalized to equal areas, the bad fit shows different scaling at large $r$.
 
-\todo{skriv om hvorfor vi ikke plukker opp alle features?}
+Our current best fit model doesn't reproduce all features of the SiO$_2$ $g(r)$, especially the two weak bumps at \SI{6.5}{\angstrom} and \SI{7.5}{\angstrom} and a constantly slightly lower value for large $r$.
+Another important detail here is that the surface atoms are not in chemical equilibrium since some neighbours of a surface atom might lose some of its covalently bonded neighbors have been removed. 
+In order to use the generated geometry in molecular dynamics, one would need either noble gas atoms which may be stable, or to apply passivation step.
 
-Also note that the surface atoms are not in chemical equilibrium since some neighbours of a surface atom might have been removed. 
-In order to use the generated geometry in molecular dynamics, one would need either noble gas
-atoms which may be stable, or to apply passivation step.
+\todo{skriv om hvorfor vi ikke plukker opp alle features? Men i diskusjon!}
 
@@
 \includegraphics[width=0.5\textwidth]{gofr_figure.pdf}
-\caption{$g(r)$ for three data sets: the simulated SiO$_2$ input data set (red), the best-fit parameter model (blue) and a random low-fit parameter model (green).}
+\caption{$g(r)$ for three data sets: the simulated SiO$_2$ input data set (red), the best-fit parameter model (blue) and a random low-fit parameter model (yellow).
+The red shaded area for low $r$ represents the part of $g(r)$ we have not used to calculate $\chi^2$ since it only captures nearest neighbor features, not the pore structure.
+The main difference between the bad fit and the others is the scaling which is due to slow convergence since the bad fit dataset has very large pores.
+There are two weak bumps at \SI{6.5}{\angstrom} and \SI{7.5}{\angstrom} which we don't pick up with our model, but the general behaviour is reproduced by our model both on short and long length scales.}
 \label{fig:gofr1}
 \end{figure}
+
 Finally, we depict a comparison of the actual simulated input SiO$_2$ data set together with a
@@
 \includegraphics[width=.45\textwidth]{comparison.png}
-\caption{Right side: A realisation of the best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$.Left side: Simulated SiO$_2$ input data set used for parameter estimation.
-Although the pore geometry seems to be similar, we see some artifacts, such as single loose atoms in the generated geometry.}
+\caption{Comparison of our best fit model (left) with parameters $\tau=-0.2$, $\psi=0.9$ and $\nu=0.03$ and the simulated SiO$_2$ input data (right).
+Although the pore geometry seems to be similar, we see some artifacts, such as single loose atoms in the generated geometry.
+To obtain a structure that can be used further, one would have to passivate the system by adding atoms so each atom has the desired coordination number.
+Both the porosity and the surface area of the model and the data are found to be similar.}
 \label{fig:porous_vs_model}
@@
 porosity $77\% \pm 3 \%$, depending on the random seed. In addition, the surface area
-of the original data set was $\sim \SI{29000}{\angstrom^2}$, while the best-fit data
+of the original data set was found to be $\sim \SI{29000}{\angstrom^2}$, while the best-fit data
 shows similar values with surface are ranging between $\sim \SI{28000}{\angstrom^2}$ and $\sim \SI{30000}{\angstrom^2}$.
-\todo{legg ved 3D plot av likelihood for aa vise at den er kjekk og fin}
 
@@
 It is worth to note that since we are chiefly interested in obtaining a set of
-parameters that reproduces a statistical correct representation of the input data,
+parameters that reproduces a statistical correct representation of the input geometry,
 we are relatively free to choose parameters as desired within the posterior distribution width of the parameter.
@@
 but from the width of the likelihood it is clear that nearby parameters would also be suitable. 
-However, the surface area is strongly correlated with this value, so even slight
+However, we found that the surface area is strongly correlated with this value, so even slight
 deviations (e.g. $\Delta \tau = 0.05$) yields a relatively large change in surface area.
-In addition, we do not claim that this model completely reproduces the pore structure
-observed in nanoporous simulations, as is evident from the RDF shown in figure \ref{fig:gofr1},
-but hope to discover a better model that will produce a $g(r)$ that is more in accordance with the data.
+Our simple measure was only using $g(r)$, and did not use the surface area to find optimal parameters.
+The porosity $\phi$ is probably indirectly correlated with the tail of $g(r)$, so introducing $\phi$ in the $\chi^2$ may not yield improved results.
+Extending $\chi^2$ to include more effects such as the surface area, pore connectivity or even physical simulations such as diffusion constants from random walks is not difficult to do.
+Adding more measures does not necessarily help since our model may not have enough degrees of freedom on enough length scales.
+
+We do not claim that our simple model fully reproduces the pore structure obtain from the molecular dynamics simulations, as is evident from $g(r)$ shown in figure \ref{fig:gofr1},
+but we hope to find better models.
 
 \subsection{$g(r)$ variation}
-We also calculated the variation in the $g(r)$ given initial conditions on both our model and simulated data. 
-In terms of the data, the variation is attributed as the initial velocities of the particles in the LAMMPS simulation. 
-For the model, it corresponds to the random seed. We found that varying these
-initial conditions have negligible effect on the calculation of $g(r)$.
+Realizations from different random seeds yield different pore networks.
+In the simulated systems, we can obtain this by using different initial conditions for the velocities.
+Our procedural noise model can be randomized by applying random permutations of the gradient vectors as described in \cite{perlin1985image}.
+By generating multiple statistically similar data sets with the two methods (molecular dynamics simulation and from our noise model), we can test the robustness of using $g(r)$ as our model measure.
+We found that varying these initial conditions have negligible effect on the calculation of $g(r)$.
 

### 2018-05-30 ms.tex
Fixed better conclusion

@@
 
-\todo{skriv om hvorfor vi ikke plukker opp alle features? Men i diskusjon!}
-
 \begin{figure}
@@
 
+\todo{skriv om hvorfor vi ikke plukker opp alle features? Men i diskusjon!}
+
 We do not claim that our simple model fully reproduces the pore structure obtain from the molecular dynamics simulations, as is evident from $g(r)$ shown in figure \ref{fig:gofr1},
@@
 \section{Conclusion}
-\todo{konklusjon maa renskrives}
-
-In this paper, we have proposed a novel method for generating a nanoporous media
-using procedural methods in the form of simplex noise.
-Procedural methods are not physical, but rather statistical models, and this gives
-them some advatages over prevalent methods: they are computationally efficient, have low memory footprints and are only limited in terms of detail by the model in question. 
+In this paper, we have proposed a novel method for generating porous materials using procedural methods in the form of simplex noise.
+Procedural methods are not physical, but rather statistical models, which gives them some advantages over prevalent methods: they are computationally efficient, have low memory footprints and are only limited in terms of detail by the model in question.
 In particular, it is possible to create simulations of arbitrary scales and resolution, indeed with any level of detail.
 
-\todo{AMS: This is a bit weak currently. State first what you have succeeded in doing.
-Then state the possible weaknesses of the model.}
-We have developed a framework to generate nanoporous media using molecular dynamics simulations and
-to estimate model parameters given such data. 
-We have shown that it is possible through a full likelihood analysis to successfully reproduce the
-input parameters of procedurally simulated data. 
-In addition, we performed an analysis on a simulated $\SI{159}{\angstrom^{3}}$ data set of SiO$_2$,
-and explain that while porosity and surface area correspond well between the model and data,
-analysis of $g(r)$ shows that the current model reproduces the main features of the material, but that there are some deviations. 
+We have generated a series of statistically similar nanoporous materials with 75 \% porosity using molecular dynamics simulations and a heat, expand and quench method explained above.
+These systems have complicated pore networks, and are used as data input to our procedural noise model.
+The procedural noise model we have used is similar to what can be found in \cite{groeneboom:2014}, and has degrees of freedom on both short and long length scales.
+
+A full MCMC framework was developed.
+It takes files with atomic positions as input, deletes atoms according to equation \eqref{eq:delete_threshold} and compares each realization with the input data using different model measures, such as $g(r)$, surface area or porosity.
+In our work, we used $g(r)$ to compare a proposed parameter set to the input data.
+Our framework is validated by reproducing known parameters on a mock data set with our MCMC code before we run the code on the simulated system from molecular dynamics.
+The model reproduces the main features of $g(r)$, but both surface area and porosity were also found to be in good agreement between our best fit model and the data even though they were not used during the fitting process.
+
+However, by visually comparing the two systems in figure \ref{fig:porous_vs_model}, we see clear differences.
+For instance, covalent bonds are certainly broken, but this is expected since our model just removes atoms with no chemical care.
+In addition, the calculated $g(r)$ of our best fit model is lacking some features that exists in the simulated data, such as two weak bumps and a slight lower-valued tail.
+
 We find the correspondence strong, which demonstrates the proof-of-principle of the use of
-procedural noise methods, but we realize that even better correspondence may be achieved by
-including other summations in the procedural noise modes.
+procedural noise methods, but we realize that even better correspondence may be achieved by including other summations in the procedural noise modes.
 
@@
 threshold cut-off $\tau$, scale fall-off persistence $\psi$ and initial scale $\nu$. 
-We have also seen that, even through both porosity and surface area accurately reproduced,
-there are features in the RDF that are not fully captured with the current model. 
+We have also seen that, even though both porosity and surface area accurately reproduced,
+there are features in the radial distribution function that are not fully captured with the current model. 
 This is evident from the images of the simulated structure seen in figure \ref{fig:porous_vs_model},

### 2018-11-04 ms.tex
stuff

@@
 to extend the number of parameters and define new models with very different properties.
-Some examples of these models are depicted in figure \ref{fig:future_models}, where the left- and
-centre images show a new model with two new parameters (offset and gain), while the rightmost panel shows a simulation where the scale $\nu$ is itself being perturbed by procedural noise, yielding strong circular asymmetric patterns.
 
@@
 
-%Minkowski functionals and topology, genus etc
-
-%Swiss noise
-
-\begin{figure*}
-\includegraphics[width=1\textwidth]{future_models.png}
-\caption{Example of future models.
-Left panel: A multi-fractal model with thin surfaces and large pores.
-Centre panel: a multi fractal with cave-like structures.
-Right panel: A model with strong asymmetric noise.
-}
-\label{fig:future_models}
-\end{figure*}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2019-01-03 ms.tex
Fixes

@@
 \usepackage{color}
-\definecolor{Blue}{rgb}{0.1,0.1,1.0} 
-\definecolor{Magenta}{rgb}{1.0,0.1,0.5} 
-\definecolor{LRed}{rgb}{0.8,0.0,0.0}
-
- \newcommand{\todo}[1]{ {\color{Magenta} TODO: \color{Blue} \textbf{#1} }}
 
@@
 \title{Structural replication of nanoporous media using procedural noise}
-%\author{A. Hafreager, N. Groeneboom, A. Malthe-S\oe renssen$^1$}
 \email{anders.hafreager@fys.uio.no}
@@
 input data to create complex, sometimes infinite repeating patterns.
-A firm example would be fractals, such as the Julia/Mandelbrot [ref Mandelbrot] set, where one
+A firm example would be fractals, such as the Mandelbrot set, where one
 tests whether the complex quadratic recurrence equation $z_{n+1} = z_n^2 +C$ for $z \in \mathbb C$
@@
 In fact, this is the very essence of procedural noise, which we are focusing on in our model,
-and the first person to create such a fast algorithm for N-dimensional spline interpolation
+and the first person to create such a fast algorithm for $N$-dimensional spline interpolation
 between pseudo-random points was Ken Perlin\cite{perlin1985image}.
@@
 \begin{figure}
-\includegraphics[width=.5\textwidth]{noise.eps}
+\includegraphics[width=.5\textwidth]{fig1.eps}
 \caption{Perlin noise (left) versus simplex noise (right). Note the
@@
 \begin{figure}
-\includegraphics[width=.5\textwidth]{noise_comparison.png}
+\includegraphics[width=.5\textwidth]{fig2.png}
 %\mbox{\epsfig{file=perlinstats.eps, width=\linewidth}}
@@
 \begin{figure}
-\includegraphics[width=.5\textwidth]{1d_noise.pdf}
+\includegraphics[width=.5\textwidth]{fig3.pdf}
 \caption{1-dimensional simplex noise with 1 (yellow), 3 (red) and 30 (blue) octaves. Adding more octaves adds high frequency features, but it still follows the 1 octave base signal.}
@@
 \begin{figure*}
-\includegraphics[width=.95\textwidth]{model_examples.png}
+\includegraphics[width=.95\textwidth]{fig4.png}
 \caption{Examples of 3D simplex noise structures in nanoporous media. Left image:
@@
 To calculate the posterior probability distribution and obtain the best fit model parameters, we need only
-to estimate the likelihood function, which is connected to the $\chi^2$\todo{Why do we have an exponential dependence of $\chi^2$ here?} by 
+to estimate the likelihood function, which is connected to the $\chi^2$ by 
 \begin{equation}
@@
 full analysis of physical system simulated with molecular dynamics, and seek parameters that reproduces $g(r)$ of this system.
-\todo{Figure 5 is the same as left in fig 9.}
 \subsection{Model verification: Parameter estimation of mock data}
 \begin{figure}[htb!]
-\includegraphics[width=.45\textwidth]{model_test.png}
+\includegraphics[width=.45\textwidth]{fig5.png}
 \caption{Simulated mock data used for the initial model testing analysis. Input parameters are $\nu=0.03$,$\tau=-0.2$ and $\psi=0.6$.}
@@
 \begin{figure*}
-\includegraphics[width=.99\textwidth]{mock_data_results_pts.png}
+\includegraphics[width=.99\textwidth]{fig6.png}
 \caption{Results from the model validation simulation.The upper panel displays the marginalised 1D posteriors for each parameter, while the lower panel depicts the marginalised 2D posteriors.
@@
 \begin{figure*}
-\includegraphics[width=.99\textwidth]{results_porous_full.png}
+\includegraphics[width=.99\textwidth]{fig7.png}
 \caption{
@@
 \begin{figure}
-\includegraphics[width=0.5\textwidth]{gofr_figure.pdf}
+\includegraphics[width=0.5\textwidth]{fig8.pdf}
 \caption{$g(r)$ for three data sets: the simulated SiO$_2$ input data set (red), the best-fit parameter model (blue) and a random low-fit parameter model (yellow).
@@
 \begin{figure}
-\includegraphics[width=.45\textwidth]{comparison.png}
+\includegraphics[width=.45\textwidth]{fig9.png}
 \caption{Comparison of our best fit model (left) with parameters $\tau=-0.2$, $\psi=0.9$ and $\nu=0.03$ and the simulated SiO$_2$ input data (right).
@@
 Adding more measures does not necessarily help since our model may not have enough degrees of freedom on enough length scales.
-
-\todo{skriv om hvorfor vi ikke plukker opp alle features? Men i diskusjon!}
 
